### 2023-11-01 docs/Proposal.tex
building repo with ningrui

\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{fancyhdr}
\pagenumbering{arabic}
\linespread{1.5}
\setlength{\parskip}{12pt}
\pagestyle{fancy}
\fancyhf{}

\lhead{ID: 400323416}
\chead{BIO 4C12 Research Proposal}
\rhead{Ningrui Xie}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}

\begin{document}
\begin{spacing}{1.0}
\begin{center}
    \Large
    \textbf{Mathematical Approaches for Simulating Epidemic Progression: Addressing Limitations of the Linear Chain Trick in ODE Models}

    \large
    \vspace{0.4cm}
    Ningrui Xie \\ Supervisor: Dr. Jonathan Dushoff
\end{center}
\end{spacing}

\section*{Abstract}
Mathematical models play a pivotal role in epidemiology by offering insights into the dynamics of infectious disease spread and enabling predictions of future trends. Yet the commonly used SIR model relies on the unrealistic assumption of exponentially distributed infectious stage durations. The strategy used by modelers to solve this problem is to subdivide the infectious stage into an integer number of substages, each having exponentially distributed substage durations with the same recovery rate (SI$^n$R model). This technique, often referred to as the Linear Chain Trick (LCT), enables the modeling of the entire infectious period durations as Erlang-distributed, aligning more closely with biological realism. However, the LCT approach has limitations, including cumbersome shape parameter determination process and discrete parameter values. In this study, we propose a novel model that assumes a geometrically distributed recovery rate for each infectious substage. We aim to assess its performance, aligning with the established SI$^n$R model, determining optimal parameters, and evaluating its fitness with real-world data. This approach holds the potential to enhance model flexibility and to develop user-friendly tools for effective epidemic modeling and prediction.
\pagebreak

\section*{Introduction}
In the realm of epidemiology, mathematical models serve as a powerful tool for understanding and effectively managing the spread of infectious disease. They enable us to identify critical factors influencing transmission and predict future trends.
\par
Among the prominent models employed in epidemiology studies, the SIR compartmental model stands as a prevalent choice \cite{McKendrick1927}\cite{Anderson1991}. This model categorizes the population into three compartments representing Susceptible, Infectious, and Recovered individuals. Central to this model is the assumption that the rate at which individuals exit the infectious compartment follows an exponential distribution with a mean time of $\frac{1}{\gamma}$, where $\gamma$ represent the individual recovery rate \cite{Hethcote2000}. However, this assumption, although mathematically convenient, deviates from the biological reality, as it fails to account for the duration of time individuals already spend in the infectious stage. A more biologically plausible approach would involve a probability distribution where the likelihood of exiting the infectious stage begins small but gradually increases as time approximates the mean duration of infection $\frac{1}{\gamma}$, eventually approaching zero as time extends infinitely. In other words, the probability distribution of the duration required to exit the infectious stage should exhibit a pronounced central tendency \cite{Sartwell1950}\cite{Simpson1952}\cite{Bailey1954}\cite{Gough1977}.
\par
To address this biological incongruity and to offer a more practical way of constructing ODE models, the Erlang distribution \cite{Anderson1980}\cite{Lloyd2001}\cite{Ma2006}, often described as the gamma distribution with integer shape parameters, becomes an instrumental choice for defining the duration of the infectious stage \cite{Olga2013}\cite{Nguyen2007}. This distribution allows us to employ the concept of subdividing the single infectious stage into $n$ substages, each following an exponential distribution with a recovery rate of $n \gamma$ ($\gamma_i = n \gamma$). The subdivision is made feasible by the property of the Erlang distribution, which is equivalent to a sequence of independent and identically distributed exponential distributions \cite{Tummal2011}. We represent the Erlang-distributed SIR model as SI$^n$R. The probability density of Erlang distribution with shape parameter $n$ and scale parameter $n \gamma$ have the following form: 
\begin{equation*}
    f(x; n,n\gamma) = \frac{(n\gamma)^n}{(n-1)!} x^{n-1} e^{-n\gamma x} \quad x>0, n \in \mathbb{N}
\end{equation*}

Note that the selection of the substage recovery rate as $n\gamma$ was to ensure that the subdivision of the infectious stage does not impact the mean recovery period ($\frac{1}{\gamma}$), as
\begin{align*}
    \frac{1}{\gamma} = \sum_{i=1}^{n} \frac{1}{\gamma_i} = \sum_{i=1}^{n} \frac{1}{n\gamma}
\end{align*}

This approach, often referred to as the Linear Chain Trick (LCT)\cite{MacDonald1978}, is a powerful technique used to transform continuous-time stochastic state transition models, where an individual's time spent in a specific state follows an Erlang distribution, into mean-field ODE models \cite{Kirosingh2019}. Importantly, the predictions obtained from the SI$^n$R model that incorporates the LCT for modeling infectious stages exhibit significant divergence from those of the original SIR model, even when the mean infectous stage duration remains the same. This disparity serves to rectify the tendency of overoptimistic predictions that suggest minimal control measures are sufficient to mitigate an epidemic in the SIR model\cite{Wearing2005}.
\par
Nonetheless, the utilization of the Linear Chain Trick to generate Erlang-distributed stage durations comes with certain limitations. Firstly, the process of determining the appropriate number of substages (shape parameter, $n$) is cumbersome and time-intensive. Previous research employing this technique necessitated data fitting for each model with varying substage counts \cite{Wearing2005}\cite{Hurtado2021}. Prior studies have explored methods for estimating the number of states in the linear chain through techniques like profile likelihood \cite{Raue2009} and model reduction \cite{Maiwald2016}. However, the inherent rigidity in the model's parameters remains a substantial limitation. As the infectious stage in the SI$^n$R model must be divided into an integer number of substages for incorporation into the ordinary differential equations (ODE), the parameters of interest, specifically the recovery rate of each substage ($n\gamma$) and the number of substages ($n$), are inherently governed by discrete values. These constraints substantially restrict the model's adaptability and performance in practical applications.
\par
In this study, we aim to address these limitations by introducing an innovative approach, which assumes a geometrically distributed recovery rate ($\gamma_i = ar^{i-1}$) from each infection substage rather than a constant rate of $n\gamma$. This transformative shift in perspective eliminates the need for using integers as the shape parameter to control the distribution, allowing us to maintain a constant value for $n$ while leveraging $a$ and $r$ to modify its shape, thereby transitioning our parameters of interest from discrete $n$ and $n\gamma$ to continuous $a$ and $r$. We refer this novel model as the SIgR model. In the initial phase of our study, we will test the performance of the proposed model by assessing its ability to align with the established SI$^n$R model. This will involve devise a method for determining the optimal values of $a$ and $r$ that most accurately match with the mean ($M$) and quadratic coefficient of variation ($\kappa$) of the SI$^n$R model across various substages, and evaluating the shape of the probability distribution for the duration of the infectious stage. We will subsequently conduct an empirical validation of the novel model through the process of data fitting. This evaluation will encompass aspects such as the model's alignment with real-world data and its predictive capabilities in projecting future trends. We will carry out this assessment alongside the established SI$^n$R model using the identical dataset to analyze the effectiveness of the SIgR model. We anticipate that this approach will enhance the model's flexibility and utility, thus paving the way for the development of user-friendly software tools that empower modelers to efficiently select and employ the most suitable models for describing and predicting future epidemic scenarios.
\pagebreak

\section*{Research Design}
We begin our study by examing the distribution of the stage durations within an SI$^n$R model to determine whether it conforms to an Erlang distribution when dividing the infectious stage into a sequence of n substages, each exponentially distributed with a mean recovery rate of $n\gamma$. To achieve this, a simplified I$^n$R model has been created, assuming a fixed number of infected individuals at time 0, with no new infections occurring throughout the study period, and confers permanent immunity upon recovery:
\begin{align*}
    \frac{dI_1}{dt} &= -n \gamma I_1 \\
    \frac{dI_2}{dt} &= n \gamma I_1 - n \gamma I_2 \\
    &\vdots \\
    \frac{dI_n}{dt} &= n \gamma I_{n-1} - n \gamma I_n \\
    \frac{dR}{dt} &= n \gamma I_n 
\end{align*}
We employed the R package ‘deSlove’ to solve the I$^n$R model, enabling us to examine how the number of infected and recovered individuals changes over time. By setting the initial number of infected individuals to 1, we derived the cumulative distribution function (CDF) for both infected and recovered individuals. Taking the derivative of the CDF functions obtained form the model would give us the probability density function (PDF), which facilitates a comparison with the Erlang distribution, characterized by shape parameter (number of substages) $n$ and scale parameter (recovery rate of each substage) $n\gamma$. The outcomes of this analysis will serve to empirically validate the notion that the sum of an integer number of independent exponentially distributed random variables follows an Erlang distribution \cite{Tummal2011}. After achieving concordance between the Erlang distribution and the infectious stage duration distribution defined by the I$^n$R model, we will use these two interchangeably.
\par
Established epidemic models featuring Erlang-distributed infection stage durations use a constant recovery rate $n\gamma$ within each substage. In our next phase, we intend to replace this constant rate with a geometrically distributed sequence $\gamma_i = ar^{i-1}$. We begin by assessing if the modified model retains the conventional Erlang-distributed infectious stage duration shape of the while preserving $M$ and $\kappa$. For models involving geometric infection substage durations (SIgR), the parameters of interest shift from $n$ and $n\gamma$ to $a$ and $r$.
\par
Our initial approach involves fixing the number of substages for the geometric distribution ($n_{PE}$) at 12 and determining the optimal $r$ value to approximate the shape of the Erlang distribution with various infection stages ($n_E$). However, the presence of the power $n$ introduces complexity when attempting to explicitly determine $r$ and $a$ for achieving concordance in $M$ and $\kappa$ between the infectious stage defined by SIgR and the Erlang distribution, shown as follow: 
\begin{align*}
    &M: \quad \underbrace{\frac{1}{\gamma}}_{\text{SI$^n$R}} = \underbrace{\frac{\frac{1}{a} (\frac{1}{r^{n_{PE}}}-1)}{\frac{1}{r}-1}}_{\text{SIgR}} \\
    &\kappa: \quad \underbrace{\frac{1}{n_E}}_{\text{SI$^n$R}} = \underbrace{\frac{\frac{\frac{1}{a^2} (\frac{1}{r^{2n_{PE}}}-1)}{\frac{1}{r^2}-1}}{M^2}}_{\text{SIgR}}
\end{align*}
Therefore, we plan to address $M$ and $\kappa$ separately. The approach involves fixing $M$ and determining the range of $r$ that satisfies the equalization of $\kappa$. To maintain the average recovery time as $\frac{1}{\gamma}$, certain constraints must be met: 
\begin{equation*}
    a = \frac{\gamma (\frac{1}{r^{n_{PE}}}-1)}{\frac{1}{r}-1}
\end{equation*}
Determining the range of values for the variable $r$ will faciliate the process of identifying the specific $r$ value, within that range, at which $K$ equals $\frac{1}{n_E}$.

\par
Upon successfully identifying $a$ and $r$ values that match $M$ and $\kappa$ of the infection stage durations distribution between the two models, we will take our approach further. Given a predetermined number of substages, a desired $\kappa$ value, and mean, we aim to numerically compute the $r$ value that matches $\kappa$. This approach will result in a novel methodology for characterizing the stage distribution, which offers increased flexibility compared to the conventional approach using a integer number of substage with identical rate, as the parameters $a$ and $r$ in the new model will be continuous rather than discrete, providing more refined control over the shape of the distribution.
\par
Following the initial theoretical evaluation of the viability of the SIgR model, our research will progress to the empirical phase, where data fitting will be undertaken to evaluate the model's efficacy. The dataset under consideration encompass disease outbreak data, vaccination records, and case-specific information. The fitting process entails the adjustment of model parameters to achieve the best fit. The assessment of model fitness will be executed through methodologies such as maximum likelihood estimation and Bayesian inference. Subsequently, the goodness-of-fit of each model will be calculate to compare their respective performance and suitability for the given dataset.

\section*{Concluding Remarks}
Our proposed projects and experiments will comprehensively assess the effectiveness of our novel model, which replaces the established constant substage recovery rate with a geometrically distributed one. The outcomes of these experiments promise to faciliate the selection process for modelers when characterizing diseases by reduce the data fitting processes and transition the model's parameter focus from discrete to continuous values, consequently augmenting its adaptability and practicality. This research holds the potential to guide more effective public health responses and decision-making during future epidemics.
\pagebreak

\bibliographystyle{unsrt}
\bibliography{reference} 
\end{document}

### 2023-11-01 Code/functions.R

@@ -10,7 +10,10 @@ r2kappaF <- function(r, n, offset=0){
 }
 
 kappa2r <- function(kappa, n){
-	u <- uniroot(r2kappa, interval=c(1, 10), n=n, offset=kappa)
+	rmax <- 7 ## BAD CODE, XNR please fix
+	if(kappa>=1) return(NA)
+	if(kappa<1/n) return(NA)
+	u <- uniroot(r2kappa, interval=c(1, rmax), n=n, offset=kappa)
 	return(u$root)
 }
 
@@ -23,6 +26,6 @@ r2kappa(2, 10)
 r2kappaF(1, 5)
 r2kappaF(2, 10)
 
-k <- kappa2r(0.4, 10)
-print(k)
-r2kappa(k, 10)
+kappa2r(1, 10)
+kappa2r(0.1, 10)
+kappa2r(0.4, 2)

### 2023-11-01 docs/Proposal.tex
modified:   Proposal.tex

@@
 \section*{Abstract}
+Hello!
 Mathematical models play a pivotal role in epidemiology by offering insights into the dynamics of infectious disease spread and enabling predictions of future trends.

### 2023-11-02 docs/Proposal.tex
Autocommit
goodbye, hello, world
Starting to comment

@@
 \documentclass[12pt]{article}
+\usepackage[margin=1in]{geometry}
+
 \usepackage{amsmath, amssymb, amsfonts}
-\usepackage[margin=1in]{geometry}
 \usepackage{setspace}
@@
 \usepackage{fancyhdr}
+\usepackage{color}
+
 \pagenumbering{arabic}
@@
 \fancyhf{}
+
+\newcommand{\comment}{}
+\renewcommand{\comment}{\nocomment}
+\renewcommand{\comment}{\showcomment}
+
+\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
+\newcommand{\nocomment}[3]{}
+
+\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
 
@@
 \section*{Abstract}
-Hello!
+
 Mathematical models play a pivotal role in epidemiology by offering insights into the dynamics of infectious disease spread and enabling predictions of future trends.
 Yet the commonly used SIR model relies on the unrealistic assumption of exponentially distributed infectious stage durations.
-The strategy used by modelers to solve this problem is to subdivide the infectious stage into an integer number of substages, each having exponentially distributed substage durations with the same recovery rate (SI$^n$R model).
+One strategy used by modelers to solve this problem is to subdivide the infectious stage into an integer number of substages, each having exponentially distributed substage durations with the same recovery rate (SI$^n$R model).
 This technique, often referred to as the Linear Chain Trick (LCT), enables the modeling of the entire infectious period durations as Erlang-distributed, aligning more closely with biological realism.
@@
 This approach holds the potential to enhance model flexibility and to develop user-friendly tools for effective epidemic modeling and prediction.
-\pagebreak
+%% \pagebreak
 
@@
 They enable us to identify critical factors influencing transmission and predict future trends.
-\par
+
 Among the prominent models employed in epidemiology studies, the SIR compartmental model stands as a prevalent choice \cite{McKendrick1927}\cite{Anderson1991}.
@@
 In other words, the probability distribution of the duration required to exit the infectious stage should exhibit a pronounced central tendency \cite{Sartwell1950}\cite{Simpson1952}\cite{Bailey1954}\cite{Gough1977}.
-\par
+
 To address this biological incongruity and to offer a more practical way of constructing ODE models, the Erlang distribution \cite{Anderson1980}\cite{Lloyd2001}\cite{Ma2006}, often described as the gamma distribution with integer shape parameters, becomes an instrumental choice for defining the duration of the infectious stage \cite{Olga2013}\cite{Nguyen2007}.
+\jd{Why do you say “often described”? It's just a definition. Try to keep your language clean and clear.}
 This distribution allows us to employ the concept of subdividing the single infectious stage into $n$ substages, each following an exponential distribution with a recovery rate of $n \gamma$ ($\gamma_i = n \gamma$).
 The subdivision is made feasible by the property of the Erlang distribution, which is equivalent to a sequence of independent and identically distributed exponential distributions \cite{Tummal2011}.
+\jd{What is meant by “made feasible by”? We could literally do it anyway. It “leads to an Erlang”, I guess would be better language.}
 We represent the Erlang-distributed SIR model as SI$^n$R. 
@@
 
-Note that the selection of the substage recovery rate as $n\gamma$ was to ensure that the subdivision of the infectious stage does not impact the mean recovery period ($\frac{1}{\gamma}$), as
+Note that using $n\gamma$ as the substage recovery rate preserves the mean recovery period ($\frac{1}{\gamma}$):
 \begin{align*}
@@
 This approach, often referred to as the Linear Chain Trick (LCT)\cite{MacDonald1978}, is a powerful technique used to transform continuous-time stochastic state transition models, where an individual's time spent in a specific state follows an Erlang distribution, into mean-field ODE models \cite{Kirosingh2019}.
-Importantly, the predictions obtained from the SI$^n$R model that incorporates the LCT for modeling infectious stages exhibit significant divergence from those of the original SIR model, even when the mean infectous stage duration remains the same.
+Importantly, predictions obtained from the SI$^n$R model can exhibit significant divergence from those of the original SIR model, even when the mean infectous stage duration remains the same.
 This disparity serves to rectify the tendency of overoptimistic predictions that suggest minimal control measures are sufficient to mitigate an epidemic in the SIR model\cite{Wearing2005}.
-\par
+\jd{Try to use simpler language -- “disparity serves to rectify”?}
+
 Nonetheless, the utilization of the Linear Chain Trick to generate Erlang-distributed stage durations comes with certain limitations.
@@
 These constraints substantially restrict the model's adaptability and performance in practical applications.
-\par
-In this study, we aim to address these limitations by introducing an innovative approach, which assumes a geometrically distributed recovery rate ($\gamma_i = ar^{i-1}$) from each infection substage rather than a constant rate of $n\gamma$.
-This transformative shift in perspective eliminates the need for using integers as the shape parameter to control the distribution, allowing us to maintain a constant value for $n$ while leveraging $a$ and $r$ to modify its shape, thereby transitioning our parameters of interest from discrete $n$ and $n\gamma$ to continuous $a$ and $r$.
-We refer this novel model as the SIgR model.
-In the initial phase of our study, we will test the performance of the proposed model by assessing its ability to align with the established SI$^n$R model.
-This will involve devise a method for determining the optimal values of $a$ and $r$ that most accurately match with the mean ($M$) and quadratic coefficient of variation ($\kappa$) of the SI$^n$R model across various substages, and evaluating the shape of the probability distribution for the duration of the infectious stage.
-We will subsequently conduct an empirical validation of the novel model through the process of data fitting.
+
+%% Where does \par come from? It's normal to just use blank lines I think in regular text.
+In this study, we aim to address these limitations by introducing a geometrically distributed recovery rate ($\gamma_i = ar^{i-1}$) from each infection substage rather than a constant rate of $n\gamma$.
+This shift in perspective eliminates the need for using integers as the shape parameter to control the distribution, allowing us to maintain a constant value for $n$ while leveraging $a$ and $r$ to modify its shape, thereby transitioning our parameters of interest from discrete $n$ and $n\gamma$ to continuous $a$ and $r$.
+We refer this novel model as the SIgR model. In the initial phase of our study, we will test the performance of the proposed model by assessing its ability to align with the established SI$^n$R model.
+This will involve devising a method to choose values of $a$ and $r$ to accurately match with the mean ($M$) and quadratic coefficient of variation ($\kappa$) of the SI$^n$R model across various substages, and evaluating the shape of the probability distribution for the duration of the infectious stage.
+We will subsequently conduct an empirical validation of our model through the process of data fitting.
 This evaluation will encompass aspects such as the model's alignment with real-world data and its predictive capabilities in projecting future trends.
@@
 We anticipate that this approach will enhance the model's flexibility and utility, thus paving the way for the development of user-friendly software tools that empower modelers to efficiently select and employ the most suitable models for describing and predicting future epidemic scenarios.
-\pagebreak
+%% \pagebreak
 
@@
 \end{align*}
-We employed the R package ‘deSlove’ to solve the I$^n$R model, enabling us to examine how the number of infected and recovered individuals changes over time.
+We employed the R package ‘deSolve’ to solve the I$^n$R model, enabling us to examine how the number of infected and recovered individuals changes over time.
 By setting the initial number of infected individuals to 1, we derived the cumulative distribution function (CDF) for both infected and recovered individuals.
@@
 After achieving concordance between the Erlang distribution and the infectious stage duration distribution defined by the I$^n$R model, we will use these two interchangeably.
-\par
+
 Established epidemic models featuring Erlang-distributed infection stage durations use a constant recovery rate $n\gamma$ within each substage.
@@
 For models involving geometric infection substage durations (SIgR), the parameters of interest shift from $n$ and $n\gamma$ to $a$ and $r$.
-\par
+
 Our initial approach involves fixing the number of substages for the geometric distribution ($n_{PE}$) at 12 and determining the optimal $r$ value to approximate the shape of the Erlang distribution with various infection stages ($n_E$).
@@
 
-\par
 Upon successfully identifying $a$ and $r$ values that match $M$ and $\kappa$ of the infection stage durations distribution between the two models, we will take our approach further.
@@
 This approach will result in a novel methodology for characterizing the stage distribution, which offers increased flexibility compared to the conventional approach using a integer number of substage with identical rate, as the parameters $a$ and $r$ in the new model will be continuous rather than discrete, providing more refined control over the shape of the distribution.
-\par
+
 Following the initial theoretical evaluation of the viability of the SIgR model, our research will progress to the empirical phase, where data fitting will be undertaken to evaluate the model's efficacy.
@@
 This research holds the potential to guide more effective public health responses and decision-making during future epidemics.
-\pagebreak
+%% \pagebreak
 

### 2024-01-12 docs/Midyear/Mid_year_report.tex
Add Midyear report

\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{subcaption}

\pagenumbering{arabic}
\linespread{1.5}
\setlength{\parskip}{12pt}
\pagestyle{fancy}
\fancyhf{}

\lhead{Ningrui Xie}
\chead{BIO 4C12 Midyear Report}
\rhead{ID: 400323416}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}

\begin{document}
\title{Mathematical Approaches for Simulating Epidemic Progression: Addressing Limitations of the Linear Chain Trick in ODE Models}

\author{Submitted to\\ Dr. Jonathan Dushoff 
\\McMaster University\\Hamiltion, Ontario, Canada L8S 4K1}
\date {January 1, 2024}
\maketitle

\centerline{Reported by}
\centerline{\textbf{Ningrui Xie}}
\centerline{\textbf{400323416}}

\newpage
\section{Abstract}
The SIR framework, because of its simplicity and ability to describe dynamic changes, is widely used in infectious disease modeling. One issue with the traditional SIR model is its implicit assumption of an exponentially distributed infectious stage duration, which often diverges from real-world scenarios. A common solution is the application of the Linear Chain Trick (LCT), wherein the single infectious stage is divided into several substages, each following an exponential distribution. The incorporation of LCT results in an Erlang distributed stage duration. However, despite its widespread application, LCT presents challenges, including the complex determination of substage numbers and rigid parameter values. Here, we propose a novel approach, maintaining a fixed number of substages but employing a geometrically distributed substage transition rate. This method, while akin to LCT in its foundational principles, offers greater flexibility in defining the stage duration shape. In contrast to the recently introduced Generalized Linear Chain Trick (GLCT), which broadens the scope to encompass a wider range of phase-type distributions, our method provides a more straightforward and efficient means to achieve model flexibility. Although our discussion focuses on epidemic modeling, the potential applications of this approach extend to various forms of dynamic modeling, offering advantages in terms of computational efficiency, ease of parameter estimation, and adaptability to diverse modeling scenarios.

\section{Introduction}
Ordinary differential equations (ODEs) are a popular choice for mathematical modeling in biology due to their ability to accurately represent the dynamic processes of systems over time. ODEs provide a powerful framework for describing phenomena such as population dynamics \cite{salisbury2011mathematical}\cite{sego2021generation}, the spread of infectious diseases \cite{Anderson1991}\cite{Diekmann2000}\cite{Feng2016}, and cell proliferation (e.g., cancer) \cite{baker1998modelling}\cite{jarrett2018mathematical}, which are characterized by continuous state changes. The inherent flexibility and simplicity of ODEs make them not only easy to solve and understand but also easy to manipulate and accommodate to various situations.

In the field of epidemiology, ordinary differential equations (ODEs) serve as a powerful tool to characterize the progression of diseases and to make predictions. A commonly used model is the SIR compartmental model \cite{Anderson1991}\cite{kermack1927contribution}, which divides the population into three states: Susceptible, Infectious, and Recovered. In this model, a susceptible individual, upon contact with the infected population at a rate $\beta$, transitions immediately to the infectious state. An infected individual then moves to the recovered state at a recovery rate of $\gamma$. However, one major limitation of this model is its implicit assumption that the duration of the infectious stage follows an exponential distribution. This can be seen in a simplified equation modeling the dynamics of the infected population
\begin{align*}
    \frac{dI}{dt} = -\gamma I
\end{align*}
Solving this equation explicitly yields
\begin{align*}
    I(t) = I_0 e^{-\gamma t}
\end{align*}
which implies a stochastic state transition model where the time spent in the infectious stage $I$ is exponentially distributed with a mean stage duration of $\frac{1}{\gamma}$. This dwell time distribution suggests that an individual is most likely to recover immediately after infection, which is often not realistic. In reality, a delay between the successful infection of an individual and the recovery from the disease is inevitable, as shown in studies conducted by Greenhalgh and Day (2017) \cite{greenhalgh2017time} using measles outbreak data.

The constraint of the basic ODE framework is that it primarily offers a singular approach to assuming specific dwell time distributions: the first event time distribution of a nonhomogeneous Poisson process, including the exponential distribution as a special case \cite{hurtado2019generalizations}. This inflexibility in specifying probability distributions to describe the duration individuals spend in a given state is a nonnegligible shortcoming of the SIR model.

It is important to note that even when the mean is constant, the shape of the stage duration distributions (i.e., the timing of stage transitions) can significantly impact the dynamics and the outcomes of predictive models in applied settings \cite{krylova2013effects}\cite{keeling2002understanding}\cite{wearing2005appropriate}\cite{nguyen2008noise}. Consequently, it becomes crucial to consider the type of distribution characterizing an individual in a specific state when constructing these models. A standard approach for developing models that incorporate different stage duration distributions is through the use of integro-differential equations (IDEs) or integral equations (IEs) \cite{hurtado2019generalizations}\cite{kermack1927contribution}\cite{hethcote1980integral}. IDEs facilitate the extension of the SIR model to include arbitrary distributions for the duration of infectiousness, as discussed in various sources \cite{feng2000endemic}\cite{hethcote1980integral}\cite{ma2006generality}. This approach allows for the capture of the 'memory effect', where the current rate of change in each compartment of the model (Susceptible, Infectious, Recovered) is influenced by the entire history of the epidemic, rather than relying on a constant rate assumption.

However, while integro-differential equations enhance the model's flexibility, they are often challenging to formulate, analyze mathematically, and simulate, especially when compared to ordinary differential equations \cite{krylova2013effects}\cite{hurtado2019generalizations}\cite{burton2005volterra}. There have been efforts to simplify IDEs into systems of ODEs \cite{macdonald2008biological}\cite{ponosov2004thew}\cite{burton2005volterra}\cite{goltser2013reducing}\cite{diekmann2018finite}, with the most notable example being the reduction of IDEs that assume Erlang-distributed stage durations \cite{hurtado2019generalizations}. Unlike the exponential distribution, a more biologically plausible assumption might be a probability distribution where the likelihood of exiting the infectious stage starts small, increases as it nears the mean duration of infection, and eventually diminishes to zero as time extends indefinitely \cite{sartwell1966incubation}\cite{bailey1954statistical}. Due to its central tendency towards the mean value, the Erlang distribution, defined as a gamma distribution with an integer shape parameter, is a more realistic choice for depicting the duration of the infectious stage. The integration of the Erlang distribution into the ODE system is facilitated by the Linear Chain Trick (LCT) \cite{macdonald1978time}\cite{smith2011introduction}. In this approach, the single infected state is divided into $n$ substages, each following an exponential distribution with a rate of $n\gamma$. This technique is based on the principle that the sum of a series of independent exponentially distributed random variables follows a gamma distribution \cite{krylova2013effects}\cite{therrien2018probability}. When the LCT is applied to the SIR model, the number of substages $n$ determines the shape of the stage duration distribution. As $n$ increases, the variance decreases, resulting in a distribution that is more peaked and narrow around the mean value $\frac{1}{\gamma}$. As $n$ approaches infinity, the distribution converges to a delta function (a deterministic value, implying that all individuals who become infectious at time $t$ recover at exactly time $t + \frac{1}{\gamma}$), and the system evolves into a delay differential equation \cite{krylova2013effects}\cite{hethcote1980integral}.

The Linear Chain Trick (LCT) is extensively employed in modeling biological processes, not only in quantifying disease progression \cite{lloyd2001destabilization}\cite{lloyd2001realistic}, but also in analyzing virus \cite{lloyd2001dependence}\cite{kakizoe2015method} and population \cite{cushing2013integrodifferential} dynamics. Studies have demonstrated that switching the stage duration distribution from an exponential to an Erlang distribution leads to predictions that align more closely with real data points \cite{kakizoe2015method}. However, despite the increased realism they bring to modeling disease processes, employing LCT in ODE models presents several limitations. One primary challenge is determining the appropriate number of substages $n$ to divide, which is a cumbersome and time-consuming process. Multiple models with varying substages must be constructed and subjected to repeated data-fitting processes to identify the optimized parameter value. Prior research has investigated approaches for estimating the number of states in the linear chain, employing techniques such as profile likelihood \cite{raue2009structural} and model reduction \cite{maiwald2016driving}. Yet, a comparative study of these methods indicates that the typical quality of experimental data is insufficient for reliably estimating the chain length using these approaches \cite{hauber2020estimating}. Another inherent limitation of using LCT is the inflexibility of the model's parameters. To integrate $n$ independent exponentially distributed random variables into the ODE system, $n$ must be an integer. As the subdivision process is done without altering the overall mean of the stage, the rate of each substage $n\gamma$ is also constrained to discrete values. These limitations significantly restrict the adaptability and practical application performance of the model.

The recently proposed Generalized Linear Chain Trick (GLCT) enhances the Linear Chain Trick (LCT) by enabling models to incorporate a broader phase-type family of distributions \cite{hurtado2019generalizations}\cite{hurtado2021building}\cite{bladt2017phase}, which includes exponential, Erlang, hypoexponential, and Coxian distributions, thus allowing for greater flexibility and the ability to capture more complexity. The GLCT achieves this by creating a higher-dimensional Markov chain, enabling a more accurate representation of the complex nature of dwell time distributions. However, this process can be mathematically heavy and challenging to interpret. Additionally, this approach may lead to a increase in the number of variables or states within the model, potentially escalating computational demands. Therefore, finding a simpler and more efficient method to enhance the flexibility of the ODE model remains essential.

In this study, we focused on one particular direction: generalize the Erlang distribted stage duration currently achieved in the model to have more diverse shape through elminate the restriction from the shape parameter (K). The value of K in the SInR model also subjected to a set of discrete values.

extending the stage duration distribution inherent in the SIR model to approximate a gamma distribution. Gamma distributions offer a more versatile distribution shape compared to Erlang distributions, as they allow for non-integer shape parameters. The challenge in extending the Linear Chain Trick (LCT) to accommodate Gamma distributions is that their shape parameters cannot be directly decomposed into exponential phases. To overcome this limitation, as well as those previously mentioned, we propose a novel approach. Instead of using a variable number of substages $(n)$ with a constant rate $(\gamma_i = n\gamma)$ for all exponentially distributed substages, we fix $n$ to a specific integer value and use a series of rates that follow a geometric pattern. This series is represented by the formula $\gamma_i = ar^{i-1}$, governed by parameters $a$ and $r$. Consequently, the parameter of interest shifts from discrete values $(n$ and $ng)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability.

To illustrate this point, consider that the Gamma distribution can be parameterized in terms of its mean $(M)$ and quadratic coefficient of variation (also refer as $\kappa$), which determine the overall shape and position of the distribution. As mentioned above, in subdividing a single stage into n substages, each exponentially distributed substage must have a rate of $n\gamma$ to maintain the overall rate of that stage. While the mean of the Erlang distribution in the ODE system can be chosen freely, $\kappa$ is constrained to discrete values, shown as follows:
\begin{align*}
    Mean: \quad &\sum_{i=1}^{n_f} \frac{1}{n\gamma} = \frac{1}{\gamma}\\
    Variance: \quad &\sum_{i=1}^{n_f} \frac{1}{n^2\gamma^2} = \frac{1}{n\gamma^2} \\
    \kappa: \quad &\frac{\sigma^2}{M^2} = \frac{1}{n} \quad n \in \mathbb{N}
\end{align*}
Our approach, using a geometric series of rates, addresses this issue by allowing both the mean and $\kappa$ to take any value. A detailed explanation and the derivation process are provided in section $4.2$. Another advantage of this method is that it enables a fixed model structure where only the parameters change, eliminating the need to determine optimal substage numbers through the construction of multiple models, a process currently required in the ODE model that employs the LCT.

In this study, we primarily focused on the application of the LCT in the epidemic SIR model, which we refer to as the SI$^n$R model, where $n$ represents the number of substages divided. The objective of our study is to develop and evaluate the method of using a geometrically distributed substage rate as a replacement for the constant rate, aiming to enhance flexibility and efficiency. We have named the model that utilizes this method the SIgR model. We hypothesize that this approach will not only be computationally efficient but also provide a better fit to real-world data. In the following sections, we first aim to align the infectious stage duration distribution shapes generated by the SI$^n$R and SIgR models. Once we have confirmed that our proposed model can effectively substitute the SInR model, we proceed to allow the SIgR model to generate stage duration distributions with characteristics unattainable by the SI$^n$R model. We then invert the process: given a desired mean and $\kappa$ (which can be any real number within a specific range), we develop a computational system to determine the parameter values $(a$ and $r)$ that enable the model to produce a stage duration distribution with these properties. Subsequently, we will conduct empirical validation of our model through data fitting. This evaluation will cover aspects such as the model’s consistency with dataset and its predictive accuracy in forecasting future trends. We anticipate that this approach will enhance the model’s adaptability and practicality, thus paving the way for the development of user-friendly software tools that empower modelers to efficiently select and employ the most suitable models for describing and predicting future epidemic scenarios.

\section{Method}
Several assumptions have been made for the unforced SIR model (Susceptible, Infectious, and Recovered) that our study is based on: 
\begin{enumerate}
    \item \textbf{Closed Population}: A constant population size with equal birth and death rates and no disease-related mortalities.
    \item \textbf{Homogeneous mixing}: Individuals have an equal chance of coming into contact with any other individual in the population. 
    \item \textbf{Ignoring Latency}: Individuals move immediately from a susceptible state to an infected state, ignoring any incubation period associated with the disease.
    \item \textbf{Lifelong immunity}: Once an individual recovers, they do not become susceptible or infectious again.
\end{enumerate}
We also assume that the state variables have been scaled by the total population size $N$ so that $S + I + R = 1$. The SIR model can be expressed as a simple system of nonlinear ordinary differential equations:
\begin{align}
    \frac{dS}{dt} &= \mu - \beta SI - \mu S \\
    \frac{dI}{dt} &= \beta SI - \gamma I - \mu I \\
    \frac{dR}{dt} &= \gamma I - \mu R
\end{align}
where $\mu$ is the birth and natural mortality rate, $\beta$ is the disease transmission rate, and $\gamma$ is the recovery rate. The mean recovery time for an infected individual is $\frac{1}{\gamma}$.

The idea of dividing the infectious stage into $n$ sub-stages exploits the fact that the sum of a sequence of independent, exponentially distributed random variables follows an Erlang distribution (i.e. gamma distributed with integer shape parameter) [Therrien and Tummala 2011]. Thus, we begin by confirming the uniformity of the results generated by simulating the SI$^n$R model with the Erlang distribution, which has the following form:
\begin{equation*}
    f(x; n,n\gamma) = \frac{(n\gamma)^n}{(n-1)!} x^{n-1} e^{-n\gamma x} \quad x>0, n \in \mathbb{N}
\end{equation*}

By breaking a single stage into $n$ substages, each exponentially distributed with a rate of $n\gamma$ ($\gamma_i = n\gamma$), the SI$^n$R model can be represented by a system of ODEs:
\begin{align*}
    \frac{dS}{dt} &= \mu - \beta SI - \mu S \\
    \frac{dI_1}{dt} &= \beta SI - (n \gamma + \mu) I_1 \\
    \frac{dI_2}{dt} &= n\gamma I_1 - (n \gamma + \mu) I_2 \\
    &\vdots \\
    \frac{dI_n}{dt} &= n\gamma I_{n-1} - (n \gamma + \mu) I_n \\
    \frac{dR}{dt} &= n \gamma I_n - \mu R
\end{align*}

Note that this decomposition process will not change the mean stage duration, as the sum of the means of the $n$ exponentially distributed random variables still equals $\frac{1}{\gamma}$
\begin{align*}
    \frac{1}{\gamma} = \sum_{i=1}^{n} \frac{1}{n\gamma}
\end{align*}

For consistency, we focused exclusively on the dynamics of the recovered compartment. The simulation results generated by the SI$^n$R model represent the cumulative distribution of the recovered population. Therefore, we need to take the derivative to convert it into a probability density function, allowing for comparison with the Erlang distribution. After confirming that the stage duration distribution generated by the SI$^n$R model aligns with a gamma distribution with an integer shape parameter, we constructed the SIgR model. This model has a fixed number of substages $n_f$ and features substage rate that are geometrically distributed ($\gamma_i = ar^{i-1}$), shown as follow
\begin{align*}
    \frac{dS}{dt} &= \mu - \beta SI - \mu S \\
    \frac{dI_1}{dt} &= \beta SI - (a + \mu) I_1 \\
    \frac{dI_2}{dt} &= a I_1 - (ar + \mu) I_2 \\
    &\vdots \\
    \frac{dI_n}{dt} &= ar^{n-1} I_{n-1} - (ar^n + \mu) I_n \\
    \frac{dR}{dt} &= ar^n I_n - \mu R
\end{align*}

We aim to fix the model structure (number of substages) and alter the parameter $a$ and $r$ to determine if this approach is more effective than changing the number of substages. Our starting point involves keeping the mean stage duration $(M)$ constant while varying $r$ to explore the possibility of a match between the SI$^n$R and SIgR models. This match will be assessed both visually and by the shape parameter, which is the coefficient of variation squared $(\kappa = \sigma^2)$.

After determining the potential match, we have formulated the formulas for $M$ and $\kappa$ in terms of $a$ and $r$. Utilizing the geometric distribution of rates in the ODE model allows for independent changes in $M$ and $\kappa$. This enables us to achieve the desired value of $\kappa$ without altering the mean, and vice versa, a flexibility not present in the traditional SI$^n$R model. Our goal is to reverse the method: starting with given values of $M$ and $\kappa$, we aim to find specific values of $a$ and $r$ that will let the model generate the stage duration distribution that satisfies these properties. Due to the complexity introduced by the high powers in the formula, solving it explicitly is challenging. To address this, we have developed a system of functions in the R programming language, employing the uniroot() function to obtain a numerical solution for r. We then reintegrate this parameter value into the SIgR model and compare it with the SInR model, using the same $M$ and $\kappa$ values. Subsequently, the mean and the shape parameter of both simulation results are calculated numerically to confirm the validity of the match.

After confirming that the newly proposed model can effectively replace the SInR model by encompassing the same infectious stage duration distributions, we will proceed to empirically test the enhanced performance of the model through data fitting. This process will be divided into two phases and will be conducted in R using the \textbf{bbmle} package. As the starting point, we will employ a deterministic model and assume only observation noise. Our first step involves coding to generate a 'fake' dataset, treating the disease incidence time series from the SIgR model as the actual case, and introducing deviations (noises) into them using a mean reporting ratio and a negative binomial dispersion parameter. The model parameters will then be estimated using the Maximum Likelihood Estimation (MLE) method, applied to the fake dataset. The objective here is to accurately recover the original parameter values that were used to generate the disease incidence data. In the next phase of the study, we will fit the model using an real-world dataset. We plan to compare the fitting results obtained from both the SI$^n$R and SIgR models. This will allow us to quantitatively assess the efficacy of our proposed method in improving the model's accuracy in reflecting actual epidemiological scenarios.

\section{Results}
In Figure 1, we present the comparison plots between the infectious stage duration distribution generated by the SI$^n$R model with the Erlang distribution. In this model, the mean recovery time is set to be 10 days $(\gamma = \frac{1}{10})$ and the infectious stage is divided into $n$ substages $(n = 4, 8, 12, 18)$, with each substage having a rate of $\frac{n}{10}$. The corresponding shape and scale parameters for the gamma distribution are $n$ and $\frac{n}{10}$, respectively. The plot demonstrates a perfect match between the two distributions, despite being generated by different methods. In the subsequent sections, we will refer to these two distributions interchangeably.
\begin{figure}[h]
    \centering
    \includegraphics[width= 0.9\textwidth]{4.png}
    \caption{Probability Distribution Comparison Between Erlang and SI$^n$R Model}
\end{figure}

\subsection{Possible match between SI$^n$R and SIgR}
In the initial step following the successful construction of the SIgR model, which employs a fixed number of substages and a geometrically distributed substage rate, we want to evaluate the theoretical possibility of a match between the stage duration distribution produced by the SIgR model and that of the traditional SInR model. This match is defined in terms of two key parameters: the mean $(M)$ and the shape parameter $(\kappa)$.

The substage number of the SIgR model $(n_f)$ has been fixed at 12, and the mean stage duration for the entire infectious period is set to 10 days. We examine the effect of using 2, 4, and 8 infectious substages in the SI$^n$R model $(n = 2, 4, 8)$ on the values of $r$ required for the SIgR model to align formulaically with the shape parameter $(\kappa)$. In Figure 2, the plots on the left illustrate the changes in stage duration distribution resulting from varying $r$ within a range that includes the formulaic match for $\kappa$. The black curve represent duration distribution from the SI$^n$R model with the associated $n$. Plots on the right depict the variations in $\kappa$, calculated using the corresponding r values, where the red dashed line is the target value of $\kappa$. It is evident from the plots that, within the specified range of r, there is a concurrence in both the mean and shape parameter values between the SInR and SIgR model. These findings suggest that the results generated by the traditional mode are replicable using the model we propose.

\begin{figure}[h!]
    \centering

    \begin{subfigure}{0.6\linewidth}
       \includegraphics[width=\linewidth]{4.1(n=2).png}
    \end{subfigure}

    \begin{subfigure}{0.6\linewidth}
       \includegraphics[width=\linewidth]{4.1(n=4).png}
    \end{subfigure}

    \begin{subfigure}{0.6\linewidth}
        \includegraphics[width=\linewidth]{4.1(n=8).png}
     \end{subfigure}

    \caption{Adjusting $r$ to Evaluate Match While Preserving Mean Value}
    \label{fig:all_plots}
 \end{figure}

\subsection{Simple Equations Facilitating Independent and Flexible Changes in Mean and $\kappa$}
In the SIgR model, the calculation of the Mean $(M)$ and Variance $(\sigma^2)$ for the infectious stage duration distribution is accomplished by adding the corresponding properties of each substage, as follows:
\begin{align}
    Mean: \quad  &\sum_{i=1}^{n_f} \frac{1}{\gamma_i} = \frac{\frac{1}{a} (\frac{1}{r^n} - 1)}{\frac{1}{r} -1} \\
    Variance: \quad &\sum_{i=1}^{n_f} \left(\frac{1}{\gamma_i} \right)^2 = \frac{\frac{1}{a^2} (\frac{1}{r^{2n}} - 1)}{\frac{1}{r^2} -1}
\end{align}
The shape parameter $\kappa$, determined by $\frac{\sigma^2}{M}$, can thus be simplified to:
\begin{align}
    \kappa = \frac{(1+r^{n_f}) (1-r)}{(1+r) (1-r^{n_f})}
\end{align}
This simplification reveals that the value of $\kappa$ is exclusively dependent on the parameter $r$, thereby granting the model independence in property changes. This independence allows for the alignment of $\kappa$ without altering the Mean, achievable through adjustments to parameter $a$. Similarly, the Mean can be scaled to a desired value without impacting $\kappa$.

Importantly, it should be noted that the value of $\kappa$ is not affected by changes in the number of substages, as $n_f$ is fixed. Consequently, $\kappa$ can be any real number within the range $[\frac{1}{n_f}, 1)$, thereby significantly enhancing the model's flexibility. In contrast, in the SI$^n$R model, the value of $\kappa$ depends on how the substages are partitioned. Following the same steps of calculation, we obtain: $\kappa = \frac{1}{n}$. The shape parameter $\kappa$ in the SI$^n$R model is therefore restricted to a set of discrete values:
\begin{align*}
    \kappa \in \left\{1, \frac{1}{2}, \frac{1}{3}, \dots \right\}
\end{align*}
This limitation potentially reduces the capability of the SI$^n$R model to accurately reflect real-world scenarios.

\subsection{Identifying Parameters $a$ and $r$ to Align SI$^n$R and SIgR with Given Mean and $\kappa$}
Utilizing formulas (4) and (6), we have developed a computational system to reverse-engineer the process. The parameters $a$ and $r$, generated by this system, were reintegrated into the SIgR model. This enable us to compare the stage duration distribution with that of the SI$^n$R model, which possesses identical mean and shape parameters, as depicted in Figure 3. Although the formulaic $M$ and $\kappa$ are computed to be the same, we also conducted a numerical calculation of both properties, with results presented in Table 1. This analysis confirms that a match is achieved at both the formulaic and numerical levels. The visual discrepancies observed between the two distributions in Figure xx can be considered negligible, since both distributions serve as theoretical representations of real-world scenarios and do not require exact correspondence.
\begin{figure}[h]
    \centering
    \includegraphics[width= \textwidth]{4.3.1.png}
    \caption{Theoretical Alignment of Mean and Shape Parameters in SI$^n$R and SIgR Models with Calculated Parameter Values $a$ and $r$}
\end{figure}

\begin{table}[h!]
    \centering
    \begin{tabular}{ l | c | r }
      \hline
      Substages & Mean & $\kappa$ \\
      \hline
      n = 2 & 10.000000 & 0.4999971 \\
      n = 4 & 10.000000 & 0.2500002 \\
      n = 8 & 10.000 & 0.125 \\
      n = 12 & 10.000000 & 0.0999986 \\
      \hline
    \end{tabular}
    \caption{Numerically Calculated Parameter Values From the SIgR Model}
\end{table}

By incorporating the infectious stage duration distribution back into the entire SI$^n$R and SIgR models, we also present the dynamics of the infectious population density generated directly from the simulations. In Figure 4, using $n=2$ and $n=4$ for the SI$^n$R model and corresponding $a$ and $r$ values ( $r_2=2.9999604, a_2=0.1500007, r_4 = 1.6626998, a_4 = 0.2503359$) for the SIgR model, a close match is observed between the results generated by the two different methods. This indicates that the results of the SInR model are reproducible using the SIgR model, but with enhanced flexibility in parameter value changes.
\begin{figure}[h]
    \centering
    \includegraphics[width= \textwidth]{4.3.2.png}
    \caption{Simulation Comparison of the Population Dynamics Between SI$^n$R and SIgR Model with Equivalent Mean and Shape Parameter}
\end{figure}

\subsection{Enhanced Flexibility in Controlling Stage Duration Distribution Shapes}
Besides offering a simpler approach to constructing an epidemic model with efficacy comparable to the SInR model, the SIgR model further enhances our ability to manipulate the shape of the stage duration distribution. This is achieved by allowing the selection of any $\kappa$ value within the range $(\frac{1}{n_f}, 1)$. Figure 5 showcases the probability distribution plots generated by the SIgR model, given a desired mean $(M=10)$ and selected $\kappa$ value $(\kappa = 0.25, 0.45, 0.65, 0.85)$. This chosen $\kappa$ value (except 0.25) is unattainable in the SI$^n$R model, distinguishing the SIgR model's unique capabilities. The new model adeptly incorporates a wider array of distribution shapes into the SIR framework. It emerges as a versatile tool for capturing a more extensive range of dynamic patterns in disease spread and control.
\begin{figure}[h]
    \centering
    \includegraphics[width= \textwidth]{4.4.2.png}
    \caption{Generating Stage Duration Distributions with Custom Mean and Shape Parameters Using the SIgR Model}
\end{figure}

\section{Discussion}
SIR compartmental models are widely utilized in modeling infectious diseases, but they encounter a significant limitation: the implicit assumption that the time an individual spends in the disease stage follows an exponential distribution. To address this issue, a common method involves subdividing the infectious stage into $n$ number of substages, each substage is exponentially distributed with a constant rate. This approach is known as the Linear Chain Trick (LCT). LCT strategy facilitates an Erlang-shaped stage duration distribution with increased reality. However, this model also faces limitations, including cumbersome parameter optimization processes and the constraint of discrete parameter values, which can affect its performance.

In this study, we propose the SIgR model, which features a fixed number of infectious substages and a geometric distribution rate for these substages, as an alternative to the SInR model that incorporates LCT. In the results section above, we demonstrated that adjusting the parameters for the geometric series $a$ and $r$ allows us to replicate the simulation results generated by the SInR model. Moreover, our model offers enhanced flexibility in shaping the distribution of the infectious stage duration, as it allows for the selection of mean $(M)$ and shape parameters $(\kappa)$ of the distribution such that $M \in \mathbb{R}^+$ and $\kappa \in [\frac{1}{n}, 1)$. Additionally, the fixed structure of our model simplifies the process of finding optimized parameter values, requiring only a one-time process during data fitting. This simplification streamlines the development of computer code for use with ODE solvers in R, ultimately saving time and reducing complexity.

Future work will primarily focus on fitting data to the model we have proposed. This will include both synthetic data generated through simulation and actual real-world data. The results of this fitting process will be compared with those obtained from the SInR model. Later on, we plan to incorporate process error considerations to increase the robustness of our model.

The anticipated outcome of this research is the development of a model that not only exceeds current models in terms of predictive accuracy but also maintains simplicity in implementation. The core design of our model is intended to be easily adaptable across various state transition models. This adaptability is vital for its application in diverse domains, ranging from epidemiology to system dynamics. The idea is designed to be extensible, capable of accommodating a range of distribution types and parameter sets. Such versatility makes it a powerful tool for modeling a wide array of complex systems, with the capability to adjust and evolve in response to new data and insights.

In conclusion, this study suggests that our method could simplify and enhance the fitting of models with flexible time distributions within an Ordinary Differential Equation (ODE) framework. This research has the potential to enhance model performance and to develop user-friendly tools for effective epidemic modeling and prediction.

\bibliographystyle{unsrt}
\bibliography{Reference}

\end{document}

### 2024-01-12 docs/Midyear/Mid_year_report.tex
Submitted Version (maybe)

@@
 
-In this study, we focused on one particular direction: generalize the Erlang distribted stage duration currently achieved in the model to have more diverse shape through elminate the restriction from the shape parameter (K).
-The value of K in the SInR model also subjected to a set of discrete values.
-
-extending the stage duration distribution inherent in the SIR model to approximate a gamma distribution.
-Gamma distributions offer a more versatile distribution shape compared to Erlang distributions, as they allow for non-integer shape parameters.
-The challenge in extending the Linear Chain Trick (LCT) to accommodate Gamma distributions is that their shape parameters cannot be directly decomposed into exponential phases.
-To overcome this limitation, as well as those previously mentioned, we propose a novel approach.
+In this study, we focused on one particular direction: generalizing the Erlang distribution of stage durations currently present in the model.
+The goal was to allow for a more diverse range of shapes, achieved by eliminating the constraints imposed by the shape parameter ($\kappa$).
+The value of $\kappa$ in the SInR model was limited to a set of discrete values, as detailed below.
+To address this limitation, as well as those previously mentioned, we propose a novel approach.
 Instead of using a variable number of substages $(n)$ with a constant rate $(\gamma_i = n\gamma)$ for all exponentially distributed substages, we fix $n$ to a specific integer value and use a series of rates that follow a geometric pattern.
@@
 
-To illustrate this point, consider that the Gamma distribution can be parameterized in terms of its mean $(M)$ and quadratic coefficient of variation (also refer as $\kappa$), which determine the overall shape and position of the distribution.
+To illustrate this point, consider that the Gamma distribution can be parameterized in terms of its mean $(M)$ and shape parameter (quadratic coefficient of variation, $\kappa$), which determine the overall shape and position of the distribution.
 As mentioned above, in subdividing a single stage into n substages, each exponentially distributed substage must have a rate of $n\gamma$ to maintain the overall rate of that stage.
@@
 \end{align*}
-Our approach, using a geometric series of rates, addresses this issue by allowing both the mean and $\kappa$ to take any value.
+Our approach, using a geometric series of rates, addresses this issue by allowing $\kappa$ to take any value within the range $[\frac{1}{n}, 1)$.
 A detailed explanation and the derivation process are provided in section $4.2$.
@@
 
-In this study, we primarily focused on the application of the LCT in the epidemic SIR model, which we refer to as the SI$^n$R model, where $n$ represents the number of substages divided.
+Throughout this research, we primarily focused on the application of the LCT in the epidemic SIR model, which we refer to as the SI$^n$R model, where $n$ represents the number of substages divided.
 The objective of our study is to develop and evaluate the method of using a geometrically distributed substage rate as a replacement for the constant rate, aiming to enhance flexibility and efficiency.
-We have named the model that utilizes this method the SIgR model.
-We hypothesize that this approach will not only be computationally efficient but also provide a better fit to real-world data.
+We have named the model that utilizes this method the SIgR model (temporarily).
+We anticipate that this approach will not only be computationally efficient but also provide a better fit to real-world data.
 In the following sections, we first aim to align the infectious stage duration distribution shapes generated by the SI$^n$R and SIgR models.
-Once we have confirmed that our proposed model can effectively substitute the SInR model, we proceed to allow the SIgR model to generate stage duration distributions with characteristics unattainable by the SI$^n$R model.
-We then invert the process: given a desired mean and $\kappa$ (which can be any real number within a specific range), we develop a computational system to determine the parameter values $(a$ and $r)$ that enable the model to produce a stage duration distribution with these properties.
+Once we have confirmed that our proposed model can effectively substitute the SInR model, we proceed to allow the SIgR model to include stage duration distributions with characteristics unattainable by the SI$^n$R model.
+We then invert the process: given a desired mean and $\kappa$, we develop the computational system to determine the parameter values $(a$ and $r)$ that enable the model to produce a stage duration distribution with these properties.
 Subsequently, we will conduct empirical validation of our model through data fitting.
 This evaluation will cover aspects such as the model’s consistency with dataset and its predictive accuracy in forecasting future trends.
-We anticipate that this approach will enhance the model’s adaptability and practicality, thus paving the way for the development of user-friendly software tools that empower modelers to efficiently select and employ the most suitable models for describing and predicting future epidemic scenarios.
+We predict that this approach will enhance the model’s adaptability and practicality, thus paving the way for the development of user-friendly software tools that empower modelers to efficiently select and employ the most suitable models for describing and predicting future epidemic scenarios.
 
@@
 The idea of dividing the infectious stage into $n$ sub-stages exploits the fact that the sum of a sequence of independent, exponentially distributed random variables follows an Erlang distribution (i.e.
-gamma distributed with integer shape parameter) [Therrien and Tummala 2011].
+gamma distributed with integer shape parameter) \cite{therrien2018probability}.
 Thus, we begin by confirming the uniformity of the results generated by simulating the SI$^n$R model with the Erlang distribution, which has the following form:
@@
 
-We aim to fix the model structure (number of substages) and alter the parameter $a$ and $r$ to determine if this approach is more effective than changing the number of substages.
+We aim to determine if altering the parameter $a$ and $r$ is more effective than changing the number of substages.
 Our starting point involves keeping the mean stage duration $(M)$ constant while varying $r$ to explore the possibility of a match between the SI$^n$R and SIgR models.
-This match will be assessed both visually and by the shape parameter, which is the coefficient of variation squared $(\kappa = \sigma^2)$.
+This match will be assessed both by the mean $(M)$ and by the shape parameter, which is the coefficient of variation squared $(\kappa = \frac{\sigma^2}{M^2})$.
 
@@
 Utilizing the geometric distribution of rates in the ODE model allows for independent changes in $M$ and $\kappa$.
-This enables us to achieve the desired value of $\kappa$ without altering the mean, and vice versa, a flexibility not present in the traditional SI$^n$R model.
+This enables us to achieve the desired value of $\kappa$ without altering the mean, and vice versa.
 Our goal is to reverse the method: starting with given values of $M$ and $\kappa$, we aim to find specific values of $a$ and $r$ that will let the model generate the stage duration distribution that satisfies these properties.
@@
 
-After confirming that the newly proposed model can effectively replace the SInR model by encompassing the same infectious stage duration distributions, we will proceed to empirically test the enhanced performance of the model through data fitting.
+After confirming that the newly proposed model can effectively replace the SI$^n$R model by encompassing the same infectious stage duration distributions, we will proceed to empirically test the enhanced performance of the model through data fitting.
 This process will be divided into two phases and will be conducted in R using the \textbf{bbmle} package.
@@
 We examine the effect of using 2, 4, and 8 infectious substages in the SI$^n$R model $(n = 2, 4, 8)$ on the values of $r$ required for the SIgR model to align formulaically with the shape parameter $(\kappa)$.
-In Figure 2, the plots on the left illustrate the changes in stage duration distribution resulting from varying $r$ within a range that includes the formulaic match for $\kappa$.
+In Figure 2, the plots on the left illustrate the changes in stage duration distribution in SIgR model resulting from varying $r$ within a range that includes the formulaic match for $\kappa$.
 The black curve represent duration distribution from the SI$^n$R model with the associated $n$.
@@
 This analysis confirms that a match is achieved at both the formulaic and numerical levels.
-The visual discrepancies observed between the two distributions in Figure xx can be considered negligible, since both distributions serve as theoretical representations of real-world scenarios and do not require exact correspondence.
+The visual discrepancies observed between the two distributions in Figure 3 can be considered negligible, since both distributions serve as theoretical representations of real-world scenarios and do not require exact correspondence.
 \begin{figure}[h]
@@
 To address this issue, a common method involves subdividing the infectious stage into $n$ number of substages, each substage is exponentially distributed with a constant rate.
-This approach is known as the Linear Chain Trick (LCT).
+This approach is also known as the Linear Chain Trick (LCT).
 LCT strategy facilitates an Erlang-shaped stage duration distribution with increased reality.

### 2024-02-13 Code/fitting.R
temp storage

@@ -134,26 +134,25 @@ plotFit(fitW, df, fixedPar, title = "E (n=2) -> E (n=2)")
 
 print(fitW$fit)
 
-# with different n
+# fit with different n
 ###################################################################### 
 
 fixedPar <- list(n = 6, μ = μ, S0 = S0, I0 = I0, ts = ts, T = T) # what if: initial pop <<>> actual pop
-fitW <- simplFit(startPar, fixedPar, df, sir.nll, "Nelder-Mead")
+fitW <- simplFit(startPar, fixedPar, df, sir.nll)
 plotFit(fitW, df, fixedPar, title = "E (n=2) -> E (n=6)")
 
-fixedPar <- list(n = 10, μ = μ, S0 = S0, I0 = I0, ts = ts, T = T) # what if: initial pop <<>> actual pop
-fitW <- simplFit(startPar, fixedPar, df, sir.nll, "Nelder-Mead")
+fixedPar <- list(n = 10, μ = μ, S0 = S0, I0 = I0, ts = ts, T = T) 
+fitW <- simplFit(startPar, fixedPar, df, sir.nll)
 plotFit(fitW, df, fixedPar, title = "E (n=2) -> E (n=10)")
 
 exp(coef(fitW$fit))
 
-fixedPar <- list(n = 12, μ = μ, S0 = S0, I0 = I0, ts = ts, T = T) # what if: initial pop <<>> actual pop
-fitW <- simplFit(startPar, fixedPar, df, sir.nll, "Nelder-Mead")
+fixedPar <- list(n = 12, μ = μ, S0 = S0, I0 = I0, ts = ts, T = T) 
+fitW <- simplFit(startPar, fixedPar, df, sir.nll)
 plotFit(fitW, df, fixedPar, title = "E (n=2) -> E (n=12)")
 
-
-# Fit Pseudo Erlang with Pseudo Erlang
 ######################################################################
+# Fit Pseudo Erlang with Pseudo Erlang
 ###################################################################### 
 
 kappa <- 2/9
@@ -166,6 +165,7 @@ nfit <- 12
 startPar <- list(logβ=-1, logD=2, logkappa=-1)
 fixedPar <- list(n = nfit, μ = μ, S0 = S0, I0 = I0, ts = ts, T = T) 
 fitW <- simplFit(startPar, fixedPar, df, sir.nll.g)
+print(fitW$fit)
[…]

### 2024-02-15 Code/fitode.R
add fitode

@@ -1,0 +1,86 @@
+# fitode
+library(fitode)
+
+
+n <- 4
+Ivec <- paste0("I",1:n)
+incidence <- sprintf("beta*S*(%s)/N", paste(Ivec, collapse = "+"))
+Itrans <- function(i, death=FALSE) {
+  if (death) return(sprintf("(n*gamma + μ)*I%d", i))
+  return(sprintf("n*gamma*I%d", i))
+}
+vars <- c("S", Ivec, "R", "Cinc")
+resp <- c(sprintf("μ*N - %s - μ*S", incidence),
+          sprintf("%s - (n*gamma + μ)*I1", incidence),
+          sprintf("%s - %s", Itrans((2:n)-1), Itrans(2:n, TRUE)),
+          sprintf("%s - μ*R", Itrans(n)),
+          incidence)
+
+
+## could also use base-R Map(), or a for loop ...
+fs <- purrr::map2(resp, vars, ~reformulate(.x, response = .y))
+fs <- lapply(fs, function(f) { environment(f) <- NULL; f })
+## initial value
+inip <- c("N - i0",
+          "i0",
+          rep(0, n+1))
+
+Ini <- purrr::map2(inip, vars, ~reformulate(.x, response = .y))
+Ini <- lapply(Ini, function(f) { environment(f) <- NULL; f })
+
+#####
+SIR_model <- odemodel(
+  name="SIR (nbinom)",
+  model=fs,
+  observation=list(
+    confirmed ~ dnbinom(mu=Cinc, size=phi)
+  ),
+  initial=Ini,
+  diffnames="Cinc",
+  par=c("beta", "gamma", "μ", "N", "i0", "phi", "n"),
+  #link=c(i0="logit")
+)
+
+#######
+
+SIR_start <- c(beta=0.2, gamma=0.1, μ = 0.01, N=10000, i0=10, phi=1000, n = n)
+ss_SIR <- simulate(SIR_model,
+                   parms=SIR_start, times=timeSeq(1,200, FALSE))
+head(ss_SIR)
+plot(ss_SIR$times, ss_SIR$Cinc, type="l")
+lines(df$Time, df$inc)
+
+###################3
+SIR_model <- odemodel(
+  name="SIR (nbinom)",
+  model=list(
+    S ~ - beta * S * I/N,
+    I ~ beta * S * I/N - gamma * I,
+    R ~ gamma * I
+  ),
[…]

### 2024-02-15 Code/fitode.R
temp storage

@@ -43,44 +43,12 @@ SIR_model <- odemodel(
 
 #######
 
-SIR_start <- c(beta=0.2, gamma=0.1, μ = 0.01, N=10000, i0=10, phi=1000, n = n)
+SIR_start <- c(beta=0.2, gamma=0.1, μ = 0.01, N=10000, i0=10, phi=1000, n = 4)
 ss_SIR <- simulate(SIR_model,
                    parms=SIR_start, times=timeSeq(1,200, FALSE))
 head(ss_SIR)
 plot(ss_SIR$times, ss_SIR$Cinc, type="l")
-lines(df$Time, df$inc)
+lines(df$Time, df$inc, col="red")
 
-###################3
-SIR_model <- odemodel(
-  name="SIR (nbinom)",
-  model=list(
-    S ~ - beta * S * I/N,
-    I ~ beta * S * I/N - gamma * I,
-    R ~ gamma * I
-  ),
-  observation=list(
-    confirmed ~ dnbinom(mu=R, size=phi)
-  ),
-  initial=list(
-    S ~ N * (1 - i0),
-    I ~ N * i0,
-    R ~ 0
-  ),
-  diffnames="R",
-  par=c("beta", "gamma", "N", "i0", "phi"),
-  #link=c(i0="logit")
-)
-SIR_start <- c(beta=70, gamma=60, N=40000, i0=0.0004, phi=6)
-SierraLeone2014b <- rbind(
-  c(times=SierraLeone2014$times[1] -
-      diff(SierraLeone2014$times)[1], confirmed=NA),
-  SierraLeone2014
-)
-ss_SIR <- simulate(SIR_model,
-                   parms=SIR_start, times=SierraLeone2014b$times)
-plot(ss_SIR$times, ss_SIR$S, type="l")
-
-plot(SierraLeone2014)
-lines(ss_SIR$times, ss_SIR$R)
 
 

### 2024-02-16 Code/fitode.R
inc -> Cinc | S0 -> N | 1000 -> nbs | inc compare

@@ -1,54 +1,103 @@
-# fitode
 library(fitode)
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal())
+library(bbmle)
 
+#######################################################################
+
+# generate SInR formula (list)
+sinrFG <- function(n) {
+  Ivec <- paste0("I",1:n)
+  incidence <- sprintf("beta*S*(%s)/N", paste(Ivec, collapse = "+"))
+  Itrans <- function(i, death=FALSE) {
+    if (death) return(sprintf("(n*gamma + μ)*I%d", i))
+    return(sprintf("n*gamma*I%d", i))
+  }
+  vars <- c("S", Ivec, "R", "Cinc")
+  resp <- c(sprintf("μ*N - %s - μ*S", incidence),
+            sprintf("%s - (n*gamma + μ)*I1", incidence),
+            sprintf("%s - %s", Itrans((2:n)-1), Itrans(2:n, TRUE)),
+            sprintf("%s - μ*R", Itrans(n)),
+            incidence)
+  
+  fs <- purrr::map2(resp, vars, ~reformulate(.x, response = .y))
+  fs <- lapply(fs, function(f) { environment(f) <- NULL; f })
+  
+  return(fs)
+}
+
+######################################################################
+
+# generate initial value list
+iniVG <- function(n) {
+  Ivec <- paste0("I",1:n)
+  vars <- c("S", Ivec, "R", "Cinc")
+  inip <- c("N - i0",
+            "i0",
+            rep(0, n+1))
+  
+  Ini <- purrr::map2(inip, vars, ~reformulate(.x, response = .y))
+  Ini <- lapply(Ini, function(f) { environment(f) <- NULL; f })
+  
+  return(Ini)
+}
+
+######################################################################
 
 n <- 4
-Ivec <- paste0("I",1:n)
-incidence <- sprintf("beta*S*(%s)/N", paste(Ivec, collapse = "+"))
-Itrans <- function(i, death=FALSE) {
-  if (death) return(sprintf("(n*gamma + μ)*I%d", i))
-  return(sprintf("n*gamma*I%d", i))
-}
-vars <- c("S", Ivec, "R", "Cinc")
-resp <- c(sprintf("μ*N - %s - μ*S", incidence),
-          sprintf("%s - (n*gamma + μ)*I1", incidence),
-          sprintf("%s - %s", Itrans((2:n)-1), Itrans(2:n, TRUE)),
[…]

### 2024-02-16 Code/fitting.R

@@ -10,7 +10,7 @@ library(pseudoErlang)
 ## Make a data frame with interval incidence, and then observations sampled from that
 simObs <- function(sinr, arp, nbs, seed=NULL) {
   if(!is.null(seed)) set.seed(seed)
-  inc <- diff(sinr[,"inc"]) #/ts
+  inc <- diff(sinr[,"Cinc"]) #/ts
   obs <- rnbinom(mu=arp*inc, size=nbs, n=length(inc))
   df <- data.frame(Time = (sinr[,"time"][-1] + sinr[,"time"][-dim(sinr)[1]])/2,
                    inc = inc,
@@ -21,23 +21,23 @@ simObs <- function(sinr, arp, nbs, seed=NULL) {
 ######################################################################
 
 ## likelihood function with parameter-trajectory tracing 🙂
-## FIXME: size needs to be _passed_ to this function
-sir.nll <- function(logβ, logD, n, μ, S0, I0, ts, T, obs) {
+## FIXME: size needs to be _passed_ to this function √
+sir.nll <- function(logβ, logD, n, μ, N, I0, ts, T, nbs, obs) {
   trace_logbeta <<- c(trace_logbeta, logβ)
   trace_logD <<- c(trace_logD, logD)
   
-  out <- as.data.frame(SInRFlow(β=exp(logβ), D=exp(logD), n=n, μ=μ, S0=S0, I0=I0, ts=ts, T=T))
-  nll <- -sum(dnbinom(x=obs, mu=diff(out$inc), size=1000, log=TRUE))
+  out <- as.data.frame(SInRFlow(β=exp(logβ), D=exp(logD), n=n, μ=μ, N=N, I0=I0, ts=ts, T=T))
+  nll <- -sum(dnbinom(x=obs, mu=diff(out$Cinc), size=nbs, log=TRUE))
 }
 
 ## Similar function for geometric pseudo-Erlang (different params)
-sir.nll.g <- function(logβ, logD, logkappa, n, μ, S0, I0, ts, T, obs) {
+sir.nll.g <- function(logβ, logD, logkappa, n, μ, N, I0, ts, T, nbs, obs) {
   trace_logbeta <<- c(trace_logbeta, logβ)
   trace_logD <<- c(trace_logD, logD)
[…]

### 2024-02-16 Code/fitode.R
temp storage

@@ -2,6 +2,19 @@ library(fitode)
 library(deSolve)
 library(ggplot2); theme_set(theme_minimal())
 library(bbmle)
+library(pseudoErlang)
+
+#######################################################################
+
+simObs <- function(sinr, arp, nbs, seed=NULL) {
+  if(!is.null(seed)) set.seed(seed)
+  inc <- diff(sinr[,"Cinc"]) #/ts
+  obs <- rnbinom(mu=arp*inc, size=nbs, n=length(inc))
+  df <- data.frame(times = sinr[,"time"],
+                   #inc = inc,
+                   obs = c(NA, obs))
+  return(df)
+}
 
 #######################################################################
 
@@ -47,11 +60,12 @@ iniVG <- function(n) {
 n <- 4
 
 # include into a function ##
+SIR_model
 SIR_model <- odemodel(
   name="SInR (nbinom)",
   model=sinrFG(n),
   observation=list(
-    confirmed ~ dnbinom(mu=Cinc, size=nbs)
+    obs ~ dnbinom(mu=Cinc, size=nbs)
   ),
   initial=iniVG(n),
   #diffnames="Cinc",
@@ -76,7 +90,7 @@ arp <- 0.9
 
 SIR_start <- c(beta=β, gamma=1/D, μ = μ, N=10000, i0=I0, nbs=nbs, n = n)
 ss_SIR <- simulate(SIR_model,
-                   parms=SIR_start, times=timeSeq(1,200, FALSE))
+                   parms=SIR_start, times=timeSeq(ts,T, FALSE))
 
 ######################################################################
 
@@ -91,13 +105,80 @@ df <- data.frame(Time = timeSeq(ts, T),
 print(ggplot(df, aes(x=Time))
       + geom_line(aes(y=currentInc, color="original inc"), linewidth = 1.5)
       + geom_line(aes(y=fitodeInc, color="fitode inc"), linetype = "dashed", linewidth = 2) 
-      + labs(y = "Incidence", title = "Original Inc vs. FitOde Inc")
+      + labs(y = "Incidence", title = "Original SInR Inc vs. FitOde SInR Inc (n=4)")
 )
 
+###################################################################### Fit
 
+n <- 4
 
-# change S0 to N
-# set nbs in function instead of 1000
[…]

### 2024-02-16 Code/fitting.R

@@ -1,8 +1,8 @@
 library(deSolve)
 library(ggplot2); theme_set(theme_minimal())
 library(bbmle)
-
-## install_github("Tiffany-Xie/pseudoErlang")
+##library(devtools)
+##install_github("Tiffany-Xie/pseudoErlang")
 library(pseudoErlang)
 
 ######################################################################

### 2024-02-16 Proposal/Proposal.tex
moving stuff around like a crazy

--- 2023-11-02 docs/Proposal.tex
@@
 \section*{Research Design}
-We begin our study by examing the distribution of the stage durations within an SI$^n$R model to determine whether it conforms to an Erlang distribution when dividing the infectious stage into a sequence of n substages, each exponentially distributed with a mean recovery rate of $n\gamma$.
+We begin our study by examing the distribution of the stage durations
+\jd{What does this mean? I think we're interested in the overall duration, not the stage durations.}
+within an SI$^n$R model to determine whether it conforms to an Erlang distribution when dividing the infectious stage into a sequence of n substages, each exponentially distributed with a mean recovery rate of $n\gamma$.
+\jd{This is a good step for us to confirm what we think is already known, right? We're not determining that it conforms, just that we're doing it right, I think.}
 To achieve this, a simplified I$^n$R model has been created, assuming a fixed number of infected individuals at time 0, with no new infections occurring throughout the study period, and confers permanent immunity upon recovery:
@@
 By setting the initial number of infected individuals to 1, we derived the cumulative distribution function (CDF) for both infected and recovered individuals.
+\jd{I think we have the CDF for recovered individuals only.
+If I were being more technical, what we're really interested in is the CDF of who has \emph{left} the infectious class.
+It corresponds to recovered in this case, but it would be good to consider the question more generally.}
 Taking the derivative of the CDF functions obtained form the model would give us the probability density function (PDF), which facilitates a comparison with the Erlang distribution, characterized by shape parameter (number of substages) $n$ and scale parameter (recovery rate of each substage) $n\gamma$.
@@
 Determining the range of values for the variable $r$ will faciliate the process of identifying the specific $r$ value, within that range, at which $K$ equals $\frac{1}{n_E}$.
+\jd{It would be great to rewrite the math here to be simple and clear, matching what we did when we were coding together.
+The key points are: there are simple equations; we can match $\kappa$ without worrying about the mean; we can the scale the mean to the desired value without changing $\kappa$.
+Next of course we want robust functions that can validate that the simulation gives us the desired values of $\mu$ and $\kappa$, and then we want to compare what the pseudo-Erlang and Erlang distributions look like.}
 
@@
 Subsequently, the goodness-of-fit of each model will be calculate to compare their respective performance and suitability for the given dataset.
+\jd{I am not sure why you are talking about goodness-of-fit here; it's OK to be very brief, more important to be clear than to write long sections.}
 
 \section*{Concluding Remarks}
-Our proposed projects and experiments will comprehensively assess the effectiveness of our novel model, which replaces the established constant substage recovery rate with a geometrically distributed one.
-The outcomes of these experiments promise to faciliate the selection process for modelers when characterizing diseases by reduce the data fitting processes and transition the model's parameter focus from discrete to continuous values, consequently augmenting its adaptability and practicality.
+Our proposed projects and experiments will assess the effectiveness of our model.
+Eventually, these methods may making it easier for modelers to fit model with flexible time distributions in an ODE framework.
 This research holds the potential to guide more effective public health responses and decision-making during future epidemics.

### 2024-02-16 Code/fitting.R
Weekly meeting

@@ -20,8 +20,7 @@ simObs <- function(sinr, arp, nbs, seed=NULL) {
 
 ######################################################################
 
-## likelihood function with parameter-trajectory tracing 🙂
-## FIXME: size needs to be _passed_ to this function √
+## likelihood function with parameter-trajectory tracing
 sir.nll <- function(logβ, logD, n, μ, N, I0, ts, T, nbs, obs) {
   trace_logbeta <<- c(trace_logbeta, logβ)
   trace_logD <<- c(trace_logD, logD)
@@ -123,7 +122,6 @@ nbs <- 1000
 sinr <- SInRFlow(β, D, n, μ, N, I0, ts, T)
 df = simObs(sinr, arp, nbs, seed = 73)
 
-## FIXME: don't call log β βe!! etc!! √
 startPar <- list(logβ=-1.5, logD=2)
 fixedPar <- list(n = n, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T)
 # Consider estimating I0 (while fixing total pop size)

### 2024-02-16 Code/foFits.R

@@ -2,6 +2,7 @@ library(fitode)
 library(deSolve)
 library(ggplot2); theme_set(theme_minimal())
 library(bbmle)
+library(purrr)
 library(pseudoErlang)
 
 #######################################################################
@@ -33,7 +34,7 @@ sinrFG <- function(n) {
             sprintf("%s - μ*R", Itrans(n)),
             incidence)
   
-  fs <- purrr::map2(resp, vars, ~reformulate(.x, response = .y))
+  fs <- map2(resp, vars, ~reformulate(.x, response = .y))
   fs <- lapply(fs, function(f) { environment(f) <- NULL; f })
   
   return(fs)
@@ -49,7 +50,7 @@ iniVG <- function(n) {
             "i0",
             rep(0, n+1))
   
-  Ini <- purrr::map2(inip, vars, ~reformulate(.x, response = .y))
+  Ini <- map2(inip, vars, ~reformulate(.x, response = .y))
   Ini <- lapply(Ini, function(f) { environment(f) <- NULL; f })
   
   return(Ini)
@@ -60,7 +61,6 @@ iniVG <- function(n) {
 n <- 4
 
 # include into a function ##
-SIR_model
 SIR_model <- odemodel(
   name="SInR (nbinom)",
   model=sinrFG(n),

### 2024-02-16 Midyear/Mid_year_report.tex
Leaving office

--- 2024-01-12 docs/Midyear/Mid_year_report.tex
@@
 This can be seen in a simplified equation modeling the dynamics of the infected population
+
+JD: This is what I mean by the cohort equation (see below). Also, probably it's good to have more numbered equations and fewer anonymous ones like this.
 \begin{align*}
@@
 In reality, a delay between the successful infection of an individual and the recovery from the disease is inevitable, as shown in studies conducted by Greenhalgh and Day (2017) \cite{greenhalgh2017time} using measles outbreak data.
+JD: “inevitable” does not match well with “shown”.
 
@@
 This inflexibility in specifying probability distributions to describe the duration individuals spend in a given state is a nonnegligible shortcoming of the SIR model.
+JD: This seems to be going too deep; the “basic” ODE framework is assuming exponential.
 
@@
 As $n$ approaches infinity, the distribution converges to a delta function (a deterministic value, implying that all individuals who become infectious at time $t$ recover at exactly time $t + \frac{1}{\gamma}$), and the system evolves into a delay differential equation \cite{krylova2013effects}\cite{hethcote1980integral}.
+JD: This \P\ is very long and doesn't really explain well the key point.
+I might explain in the other direction: LCT equations are much simpler than IDEs, and can produce dwell distributions that are more realistic than the basic model.
+The fact that the the n-box LCT formulation is an Erlang is kind of a detail.
 
@@
 \end{align*}
+
+JD: I feel there's some confusion below.
+When thinking about the cohort model (like the very first displayed equation), we focus on dynamics of recovered.
+That's how we construct our infectious-period distributions.
+But when thinking about a transmission model (like the one below), we typically focus on incidence (cumulative, interval and instantaneous).
+For your thesis, it might be good to explicitly write out cohort models to support the stage-distribution stuff above, and to clarify the distinction between cohort models and transmission models.
+The reason I call them cohort models is that each “cohort” of
 

### 2024-02-17 Midyear/Mid_year_report.tex
modified:   Mid_year_report.tex

@@
 \renewcommand{\headrulewidth}{0pt}
+
+\newcommand{\perday}{\ensuremath{/mathrm{day}}
 
@@
 
+JD: It would be good to have a clearer explanation of the goal. “Effective” is not very explicit here.
 We aim to determine if altering the parameter $a$ and $r$ is more effective than changing the number of substages.
@@
 This match will be assessed both by the mean $(M)$ and by the shape parameter, which is the coefficient of variation squared $(\kappa = \frac{\sigma^2}{M^2})$.
+JD: Not sure why you're using $M$, and whether you really mean the mean “stage” duration. We should be focused on the mean total duration, and call it $D$, I think.
 
@@
 As the starting point, we will employ a deterministic model and assume only observation noise.
-Our first step involves coding to generate a 'fake' dataset, treating the disease incidence time series from the SIgR model as the actual case, and introducing deviations (noises) into them using a mean reporting ratio and a negative binomial dispersion parameter.
+Our first step involves coding to generate a `fake' dataset, treating the disease incidence time series from the SIgR model as the actual case, and introducing deviations (noises) into them using a mean reporting ratio and a negative binomial dispersion parameter.
 The model parameters will then be estimated using the Maximum Likelihood Estimation (MLE) method, applied to the fake dataset.
@@
 In this model, the mean recovery time is set to be 10 days $(\gamma = \frac{1}{10})$ and the infectious stage is divided into $n$ substages $(n = 4, 8, 12, 18)$, with each substage having a rate of $\frac{n}{10}$.
-The corresponding shape and scale parameters for the gamma distribution are $n$ and $\frac{n}{10}$, respectively.
+The corresponding shape and scale parameters for the gamma distribution are $n$ and $\frac{n}{10\perday}$, respectively.
 The plot demonstrates a perfect match between the two distributions, despite being generated by different methods.
@@
 \end{figure}
+JD: Respect the units!
 
@@
 
+JD: I would frame this differently. Both PE and Erlang \emph{can} change the value of $n$; the real difference is that PE can be flexible even when $n$ doesn't change.
 Importantly, it should be noted that the value of $\kappa$ is not affected by changes in the number of substages, as $n_f$ is fixed.
@@
 \end{table}
+
+JD: A table caption should have a clear explanation of what's in the table for people who are skimming (your figure captions should probably be improved, too).
+It's not clear what you're showing here, or whether it's right.
+Maybe show target values as well as observed values?
+Also, there seems to be a mismatch between $n=12$ and the corresponding $\kappa$?
+Also, there are too many significant figures here.
+We don't need to fix this table, since we don't need it moving forward, these are just general tips.
 

### 2024-03-07 thesis/example.tex
not working!!

--- 2024-02-17 Midyear/Mid_year_report.tex
@@
 \lhead{Ningrui Xie}
-\chead{BIO 4C12 Midyear Report}
+\chead{BIO 4C12 Example document}
 \rhead{ID: 400323416}
@@
 
-\newcommand{\perday}{\ensuremath{/mathrm{day}}
+\newcommand{\perday}{\ensuremath{/mathrm{day}}}
 
@@
     \centering
-    \includegraphics[width= 0.9\textwidth]{4.png}
+    \includegraphics{figures/test.Rout.pdf}
     \caption{Probability Distribution Comparison Between Erlang and SI$^n$R Model}
 \end{figure}
-JD: Respect the units!
-
-\subsection{Possible match between SI$^n$R and SIgR}
-In the initial step following the successful construction of the SIgR model, which employs a fixed number of substages and a geometrically distributed substage rate, we want to evaluate the theoretical possibility of a match between the stage duration distribution produced by the SIgR model and that of the traditional SInR model.
-This match is defined in terms of two key parameters: the mean $(M)$ and the shape parameter $(\kappa)$.
-
-The substage number of the SIgR model $(n_f)$ has been fixed at 12, and the mean stage duration for the entire infectious period is set to 10 days.
-We examine the effect of using 2, 4, and 8 infectious substages in the SI$^n$R model $(n = 2, 4, 8)$ on the values of $r$ required for the SIgR model to align formulaically with the shape parameter $(\kappa)$.
-In Figure 2, the plots on the left illustrate the changes in stage duration distribution in SIgR model resulting from varying $r$ within a range that includes the formulaic match for $\kappa$.
-The black curve represent duration distribution from the SI$^n$R model with the associated $n$.
-Plots on the right depict the variations in $\kappa$, calculated using the corresponding r values, where the red dashed line is the target value of $\kappa$.
-It is evident from the plots that, within the specified range of r, there is a concurrence in both the mean and shape parameter values between the SInR and SIgR model.
-These findings suggest that the results generated by the traditional mode are replicable using the model we propose.
-
-\begin{figure}[h!]
-    \centering
-
-    \begin{subfigure}{0.6\linewidth}
-       \includegraphics[width=\linewidth]{4.1(n=2).png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.6\linewidth}
-       \includegraphics[width=\linewidth]{4.1(n=4).png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.6\linewidth}
-        \includegraphics[width=\linewidth]{4.1(n=8).png}
-     \end{subfigure}
-
-    \caption{Adjusting $r$ to Evaluate Match While Preserving Mean Value}
-    \label{fig:all_plots}
- \end{figure}
-
-\subsection{Simple Equations Facilitating Independent and Flexible Changes in Mean and $\kappa$}
-In the SIgR model, the calculation of the Mean $(M)$ and Variance $(\sigma^2)$ for the infectious stage duration distribution is accomplished by adding the corresponding properties of each substage, as follows:
-\begin{align}
-    Mean: \quad  &\sum_{i=1}^{n_f} \frac{1}{\gamma_i} = \frac{\frac{1}{a} (\frac{1}{r^n} - 1)}{\frac{1}{r} -1} \\
-    Variance: \quad &\sum_{i=1}^{n_f} \left(\frac{1}{\gamma_i} \right)^2 = \frac{\frac{1}{a^2} (\frac{1}{r^{2n}} - 1)}{\frac{1}{r^2} -1}
-\end{align}
-The shape parameter $\kappa$, determined by $\frac{\sigma^2}{M}$, can thus be simplified to:
-\begin{align}
-    \kappa = \frac{(1+r^{n_f}) (1-r)}{(1+r) (1-r^{n_f})}
-\end{align}
-This simplification reveals that the value of $\kappa$ is exclusively dependent on the parameter $r$, thereby granting the model independence in property changes.
-This independence allows for the alignment of $\kappa$ without altering the Mean, achievable through adjustments to parameter $a$.
-Similarly, the Mean can be scaled to a desired value without impacting $\kappa$.
-
-JD: I would frame this differently. Both PE and Erlang \emph{can} change the value of $n$; the real difference is that PE can be flexible even when $n$ doesn't change.
-Importantly, it should be noted that the value of $\kappa$ is not affected by changes in the number of substages, as $n_f$ is fixed.
-Consequently, $\kappa$ can be any real number within the range $[\frac{1}{n_f}, 1)$, thereby significantly enhancing the model's flexibility.
-In contrast, in the SI$^n$R model, the value of $\kappa$ depends on how the substages are partitioned.
-Following the same steps of calculation, we obtain: $\kappa = \frac{1}{n}$.
-The shape parameter $\kappa$ in the SI$^n$R model is therefore restricted to a set of discrete values:
-\begin{align*}
-    \kappa \in \left\{1, \frac{1}{2}, \frac{1}{3}, \dots \right\}
-\end{align*}
-This limitation potentially reduces the capability of the SI$^n$R model to accurately reflect real-world scenarios.
-
-\subsection{Identifying Parameters $a$ and $r$ to Align SI$^n$R and SIgR with Given Mean and $\kappa$}
-Utilizing formulas (4) and (6), we have developed a computational system to reverse-engineer the process.
-The parameters $a$ and $r$, generated by this system, were reintegrated into the SIgR model.
-This enable us to compare the stage duration distribution with that of the SI$^n$R model, which possesses identical mean and shape parameters, as depicted in Figure 3.
-Although the formulaic $M$ and $\kappa$ are computed to be the same, we also conducted a numerical calculation of both properties, with results presented in Table 1.
-This analysis confirms that a match is achieved at both the formulaic and numerical levels.
-The visual discrepancies observed between the two distributions in Figure 3 can be considered negligible, since both distributions serve as theoretical representations of real-world scenarios and do not require exact correspondence.
-\begin{figure}[h]
-    \centering
-    \includegraphics[width= \textwidth]{4.3.1.png}
-    \caption{Theoretical Alignment of Mean and Shape Parameters in SI$^n$R and SIgR Models with Calculated Parameter Values $a$ and $r$}
-\end{figure}
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{ l | c | r }
-      \hline
-      Substages & Mean & $\kappa$ \\
-      \hline
-      n = 2 & 10.000000 & 0.4999971 \\
-      n = 4 & 10.000000 & 0.2500002 \\
-      n = 8 & 10.000 & 0.125 \\
-      n = 12 & 10.000000 & 0.0999986 \\
-      \hline
-    \end{tabular}
-    \caption{Numerically Calculated Parameter Values From the SIgR Model}
-\end{table}
-
-JD: A table caption should have a clear explanation of what's in the table for people who are skimming (your figure captions should probably be improved, too).
-It's not clear what you're showing here, or whether it's right.
-Maybe show target values as well as observed values?
-Also, there seems to be a mismatch between $n=12$ and the corresponding $\kappa$?
-Also, there are too many significant figures here.
-We don't need to fix this table, since we don't need it moving forward, these are just general tips.
-
-By incorporating the infectious stage duration distribution back into the entire SI$^n$R and SIgR models, we also present the dynamics of the infectious population density generated directly from the simulations.
-In Figure 4, using $n=2$ and $n=4$ for the SI$^n$R model and corresponding $a$ and $r$ values ( $r_2=2.9999604, a_2=0.1500007, r_4 = 1.6626998, a_4 = 0.2503359$) for the SIgR model, a close match is observed between the results generated by the two different methods.
-This indicates that the results of the SInR model are reproducible using the SIgR model, but with enhanced flexibility in parameter value changes.
-\begin{figure}[h]
-    \centering
-    \includegraphics[width= \textwidth]{4.3.2.png}
-    \caption{Simulation Comparison of the Population Dynamics Between SI$^n$R and SIgR Model with Equivalent Mean and Shape Parameter}
-\end{figure}
-
-\subsection{Enhanced Flexibility in Controlling Stage Duration Distribution Shapes}
-Besides offering a simpler approach to constructing an epidemic model with efficacy comparable to the SInR model, the SIgR model further enhances our ability to manipulate the shape of the stage duration distribution.
-This is achieved by allowing the selection of any $\kappa$ value within the range $(\frac{1}{n_f}, 1)$.
-Figure 5 showcases the probability distribution plots generated by the SIgR model, given a desired mean $(M=10)$ and selected $\kappa$ value $(\kappa = 0.25, 0.45, 0.65, 0.85)$.
-This chosen $\kappa$ value (except 0.25) is unattainable in the SI$^n$R model, distinguishing the SIgR model's unique capabilities.
-The new model adeptly incorporates a wider array of distribution shapes into the SIR framework.
-It emerges as a versatile tool for capturing a more extensive range of dynamic patterns in disease spread and control.
-\begin{figure}[h]
-    \centering
-    \includegraphics[width= \textwidth]{4.4.2.png}
-    \caption{Generating Stage Duration Distributions with Custom Mean and Shape Parameters Using the SIgR Model}
-\end{figure}
-
-\section{Discussion}
-SIR compartmental models are widely utilized in modeling infectious diseases, but they encounter a significant limitation: the implicit assumption that the time an individual spends in the disease stage follows an exponential distribution.
-To address this issue, a common method involves subdividing the infectious stage into $n$ number of substages, each substage is exponentially distributed with a constant rate.
-This approach is also known as the Linear Chain Trick (LCT).
-LCT strategy facilitates an Erlang-shaped stage duration distribution with increased reality.
-However, this model also faces limitations, including cumbersome parameter optimization processes and the constraint of discrete parameter values, which can affect its performance.
-
-In this study, we propose the SIgR model, which features a fixed number of infectious substages and a geometric distribution rate for these substages, as an alternative to the SInR model that incorporates LCT.
-In the results section above, we demonstrated that adjusting the parameters for the geometric series $a$ and $r$ allows us to replicate the simulation results generated by the SInR model.
-Moreover, our model offers enhanced flexibility in shaping the distribution of the infectious stage duration, as it allows for the selection of mean $(M)$ and shape parameters $(\kappa)$ of the distribution such that $M \in \mathbb{R}^+$ and $\kappa \in [\frac{1}{n}, 1)$.
-Additionally, the fixed structure of our model simplifies the process of finding optimized parameter values, requiring only a one-time process during data fitting.
-This simplification streamlines the development of computer code for use with ODE solvers in R, ultimately saving time and reducing complexity.
-
-Future work will primarily focus on fitting data to the model we have proposed.
-This will include both synthetic data generated through simulation and actual real-world data.
-The results of this fitting process will be compared with those obtained from the SInR model.
-Later on, we plan to incorporate process error considerations to increase the robustness of our model.
-
-The anticipated outcome of this research is the development of a model that not only exceeds current models in terms of predictive accuracy but also maintains simplicity in implementation.
-The core design of our model is intended to be easily adaptable across various state transition models.
-This adaptability is vital for its application in diverse domains, ranging from epidemiology to system dynamics.
-The idea is designed to be extensible, capable of accommodating a range of distribution types and parameter sets.
-Such versatility makes it a powerful tool for modeling a wide array of complex systems, with the capability to adjust and evolve in response to new data and insights.
-
-In conclusion, this study suggests that our method could simplify and enhance the fitting of models with flexible time distributions within an Ordinary Differential Equation (ODE) framework.
-This research has the potential to enhance model performance and to develop user-friendly tools for effective epidemic modeling and prediction.
 
 \bibliographystyle{unsrt}
-\bibliography{Reference}
+\bibliography{Reference.bib}
 

### 2024-03-07 figures/test.R

@@ -1,0 +1,4 @@
+library(shellpipes)
+startGraphics()
+
+plot(1:20)

### 2024-03-07 thesis/example.tex
meeting

@@
 \section{Results}
+
 In Figure 1, we present the comparison plots between the infectious stage duration distribution generated by the SI$^n$R model with the Erlang distribution.
@@
 
+In Figure 2, we present more excellent stuff!
+\begin{figure}[h]
+    \centering
+    \includegraphics{figures/excellent.Rout-1.pdf}
+    \caption{Probability Distribution Comparison Between Erlang and SI$^n$R Model}
+\end{figure}
+
 \bibliographystyle{unsrt}

### 2024-03-07 Code/hardtest_error.R

@@ -2,11 +2,13 @@ library(deSolve)
 library(ggplot2); theme_set(theme_minimal())
 library(bbmle)
 
-##library(devtools)
+##library(remotes)
 ##install_github("Tiffany-Xie/pseudoErlang")
 library(pseudoErlang)
 
 library(shellpipes)
+## This statement will do nothing inside the pipeline, but allows to run this script interactively
+rpcall("hardtest_error.Rout .pipestar Code/hardtest_error.R tempfunc.rda")
 
 loadEnvironments()
 

### 2024-03-07 figures/excellent.R

@@ -1,0 +1,81 @@
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal())
+library(bbmle)
+library(pseudoErlang)
+
+library(shellpipes)
+startGraphics(height=4)
+
+set.seed(223)
+
+numObs <- 1000
+kappa <- 0.4
+D <- 5
+
+## k = 1/shape
+## D = shape*scale = scale/kappa
+
+v <- round(rgamma(numObs, shape=1/kappa, scale=D*kappa))
+
+print(mean(v))
+print(sd(v)^2/mean(v)^2)
+
+ggplot(data.frame(v = v), aes(x = v)) + geom_histogram()
+
+## start/end times
+
+## start is a uniform
+## end is start + v [above]
+## round them both before subtracting
+
+gammaDis <- function(time, mean, kappa) {
+  alpha = 1/kappa
+  beta = 1/(kappa*mean)
+  beta^alpha / factorial(alpha-1) * time^(alpha-1) * exp(-beta*time)  
+}
+
+
+sir.nll <- function(logmean, logkappa, ts,T, obs) {
+  out <- gammaDis(timeSeq(ts, T, FALSE), exp(logmean), exp(logkappa))
+  nll <- -sum(dnorm(x=obs, mean=out, sd=1, log=TRUE))
+}
+
+simplFit <- function(startPar, fixedPar, obs, likelihood.m, optMethod="Nelder-Mead") {
+  
+  fit0 <- mle2(likelihood.m, 
+               data=list(obs=obs),
+               start=startPar,
+               fixed=fixedPar,
+               method=optMethod,
+               control=list(trace = 0))
+  
+  return(fit0)
+}
+
+ts = 1; T = 30; mean = 10; kappa = 1/2
+g <- gammaDis(timeSeq(ts, T, FALSE), mean, kappa)
+
+startPar = list(logmean = 2, logkappa = -0.5)
+fixedPar = list(ts=1, T=30)
+fit = simplFit(startPar, fixedPar, g, sir.nll)
+
+df <- data.frame(time = timeSeq(ts, T, FALSE), 
+                 obs=g, fitted = gammaDis(timeSeq(ts, T, FALSE), 
+                                          exp(coef(fit)[["logmean"]]),
+                                          exp(coef(fit)[["logkappa"]])))
+
+ggplot(df, aes(time)) +
+  geom_line(aes(y = obs, col = "obs")) + 
+  geom_line(aes(y = fitted, col = "fitted")) + 
[…]

### 2024-03-07 figures/test.R

@@ -1,4 +1,4 @@
 library(shellpipes)
 startGraphics()
 
-plot(1:20)
+plot((1:10)^2)

### 2024-03-13 Code/Interval.R
temp storage

@@ -38,6 +38,7 @@ gammaDis <- function(time, mean, kappa) {
 
 
 sir.nll <- function(logmean, logkappa, ts,T, obs) {
+  #obs = round(obs)
   out <- gammaDis(timeSeq(ts, T, FALSE), exp(logmean), exp(logkappa))
   nll <- -sum(dnorm(x=obs, mean=out, sd=1, log=TRUE))
 }

### 2024-03-14 Code/Interval.R
temp storage

@@ -3,11 +3,15 @@ library(ggplot2); theme_set(theme_minimal())
 library(bbmle)
 library(pseudoErlang)
 
+######################################################################
+## gamma -> gamma
+######################################################################
+
 set.seed(223)
 
 numObs <- 1000
-kappa <- 0.4
-D <- 5
+kappa <- 0.3
+D <- 7
 
 ## k = 1/shape
 ## D = shape*scale = scale/kappa
@@ -15,69 +19,59 @@ D <- 5
 v <- round(rgamma(numObs, shape=1/kappa, scale=D*kappa))
 
 print(mean(v))
-print(sd(v)^2/mean(v)^2)
-
-df <- data.frame(v = v)
-ggplot(df, aes(x = v)) + geom_histogram()
+print(sd(v)^2/mean(v)^2) # nice!
 
 ## start/end times
-
 ## start is a uniform
 ## end is start + v [above]
 ## round them both before subtracting
 
-gammaDelay <- function(n, mean, kappa){
-  
-}
-
-gammaDis <- function(time, mean, kappa) {
-  alpha = 1/kappa
-  beta = 1/(kappa*mean)
-  beta^alpha / factorial(alpha-1) * time^(alpha-1) * exp(-beta*time)  
-}
-
-
-sir.nll <- function(logmean, logkappa, ts,T, obs) {
-  #obs = round(obs)
-  out <- gammaDis(timeSeq(ts, T, FALSE), exp(logmean), exp(logkappa))
-  nll <- -sum(dnorm(x=obs, mean=out, sd=1, log=TRUE))
-}
-
-simplFit <- function(startPar, fixedPar, obs, likelihood.m, optMethod="Nelder-Mead") {
-  
-  fit0 <- mle2(likelihood.m, 
-               data=list(obs=obs),
-               start=startPar,
-               fixed=fixedPar,
-               method=optMethod,
-               control=list(trace = 0))
-  
-  return(fit0)
-}
-
-ts = 1; T = 30; mean = 10; kappa = 1/2
-g <- gammaDis(timeSeq(ts, T, FALSE), mean, kappa)
-
-startPar = list(logmean = 2, logkappa = -0.5)
-fixedPar = list(ts=1, T=30)
-fit = simplFit(startPar, fixedPar, g, sir.nll)
-
-df <- data.frame(time = timeSeq(ts, T, FALSE), 
-                 obs=g, fitted = gammaDis(timeSeq(ts, T, FALSE), 
-                                          exp(coef(fit)[["logmean"]]),
[…]

### 2024-03-14 Code/hardtest_error.R

@@ -1,16 +1,9 @@
 library(deSolve)
 library(ggplot2); theme_set(theme_minimal())
 library(bbmle)
-
-##library(remotes)
+##library(devtools)
 ##install_github("Tiffany-Xie/pseudoErlang")
 library(pseudoErlang)
-
-library(shellpipes)
-## This statement will do nothing inside the pipeline, but allows to run this script interactively
-rpcall("hardtest_error.Rout .pipestar Code/hardtest_error.R tempfunc.rda")
-
-loadEnvironments()
 
 ######################################################################
  

### 2024-03-14 Code/tempfunc.R

@@ -1,6 +1,4 @@
 
-library(shellpipes)
-rpcall("tempfunc.Rout .pipestar Code/tempfunc.R")
 
 simObs <- function(sinr, arp, nbs, seed=NULL) {
   if(!is.null(seed)) set.seed(seed)
@@ -96,4 +94,7 @@ plotTrace <- function(fitW) {
     labs(x = "log(D)", y = "log(beta)", color = "Time")
 }
 
-saveEnvironment()
+######################################################################
+
+
+

### 2024-03-15 Code/Interval.R
temp storage

@@ -67,6 +67,77 @@ ggplot(df) +
 ######################################################################
 ## gamma -> Erlang
 ######################################################################
+gE.nll <- function(logmean, logkappa, interval) { 
+  ts=max(interval)/length(interval)
+  sinr <- as.data.frame(SInRFlow(β=0, D=exp(logmean), n=round(1/exp(logkappa)), μ=0, 
+                                 N=1, I0=1, 
+                                 ts=ts, T=max(interval)))
+  -sum(dnorm(x=interval, mean=diff(sinr$R)/ts*1000, log=TRUE))
+}
+
+#sinr = as.data.frame(SInRFlow(0,7,4,0,1,1,ts,25))
+#plot(timeSeq(ts,25), diff(sinr$R)/ts*n)
+
+startPar = list(logmean = 2, logkappa = -1)
+fit <- mle2(gE.nll, 
+            data = list(interval = g_interval),
+            start = startPar,
+            method = "Nelder-Mead",
+            control = list(maxit = 10000))
+
+time = seq(0, max(g_interval), length.out=n)
+fitmean = exp(coef(fit)[["logmean"]])
+fitkappa = exp(coef(fit)[["logkappa"]])
+ts = max(g_interval)/length(g_interval)
+df <- data.frame(Time = time,
+                 interval = g_interval,
+                 gamma = dgamma(time, shape=1/kappa, scale=mean*kappa)*n,
+                 fit_gamma = diff(SInRFlow(β=0, D=fitmean, n=round(1/fitkappa), μ=0, 
+                                      N=1, I0=1, 
+                                      ts=ts, T=max(g_interval))[,"R"])/ts*1000)
+ggplot(df) + 
+  geom_histogram(aes(x=interval)) +
+  geom_line(aes(x=Time, y=gamma, color="Gamma"), linewidth=1.5) +
+  geom_line(aes(x=Time, y=fit_gamma, color="Gamma after fit"), linewidth=1.5) +
+  labs(x = "Interval", y = "Count", title = "gamma -> Erlang")
+
+######################################################################
+## gamma -> Pseudo Erlang
+######################################################################
+gPE.nll <- function(logmean, logkappa, interval) { 
+  ts=max(interval)/length(interval)
[…]

### 2024-03-15 Code/Interval.R
temp storage

@@ -64,15 +64,22 @@ ggplot(df) +
   geom_line(aes(x=Time, y=fit_gamma, color="Gamma after fit"), linewidth=1.5) +
   labs(x = "Interval", y = "Count", title = "gamma -> gamma")
 
+quit()
 ######################################################################
 ## gamma -> Erlang
 ######################################################################
-gE.nll <- function(logmean, logkappa, interval) { 
-  ts=max(interval)/length(interval)
+count_interval <- as.data.frame(table(g_interval))
+names(count_interval) <- c("Interval", "Count")
+count_interval$Interval <- as.numeric(count_interval$Interval)
+count_interval <- rbind(data.frame(Interval=0, Count=0), count_interval)
+
+ts=max(count_interval$Interval)/dim(count_interval)[1]
+tmax= max(count_interval$Interval)
+gE.nll <- function(logmean, logkappa, c_interval) { 
   sinr <- as.data.frame(SInRFlow(β=0, D=exp(logmean), n=round(1/exp(logkappa)), μ=0, 
                                  N=1, I0=1, 
-                                 ts=ts, T=max(interval)))
-  -sum(dnorm(x=interval, mean=diff(sinr$R)/ts*1000, log=TRUE))
+                                 ts=ts, T=tmzx))
+  -sum(dnorm(x=c_interval, mean=diff(sinr$R)/ts*1000, log=TRUE))
 }
 
 #sinr = as.data.frame(SInRFlow(0,7,4,0,1,1,ts,25))
@@ -80,7 +87,7 @@ gE.nll <- function(logmean, logkappa, interval) {
 
 startPar = list(logmean = 2, logkappa = -1)
 fit <- mle2(gE.nll, 
-            data = list(interval = g_interval),
+            data = list(interval = count_interval$Interval),
             start = startPar,
             method = "Nelder-Mead",
             control = list(maxit = 10000))

### 2024-03-16 Code/Pre_plot.R
+ Pre_plot

@@ -1,0 +1,329 @@
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal(base_size = 17))
+library(bbmle)
+##library(devtools)
+##install_github("Tiffany-Xie/pseudoErlang")
+library(pseudoErlang)
+
+######################################################################
+
+erlang <- function(x, n, γ) {
+  (n*γ)^n*x^(n-1)*exp(-n*γ*x)/factorial(n-1)
+} 
+
+time <- timeSeq(0.1, 40)
+N <- c(1,2,4,10,20,70)
+γ <- 0.1
+
+df <- as.data.frame(expand.grid(Time=time,n=paste0("n=",N)))
+PD <- c()
+for (n in N) {
+  PD <- c(PD, erlang(time, n, γ))
+}
+df$PD <- PD
+
+ggplot(df, aes(x=Time, y = PD, color = n)) + 
+  geom_line(linewidth = 1) +
+  scale_color_manual(values = c('red', 'blue', 'green', 'yellow2', 'pink', 'magenta4')) + 
+  labs(x="Stage duration, days", y = "Probability Density")
+  
+
+
+
+######################################################################
+
+
+compareSDD <- function(D, n, nfix, ts, T, otherPar) {
+  sinr <- SInRFlow(otherPar[["β"]], D, n, 
+                   otherPar[["μ"]], otherPar[["N"]], 
+                   otherPar[["I0"]], ts, T)
+  sinner <- sinnerFlow(otherPar[["β"]], D, kappa=1/n, nfix, 
+                       otherPar[["μ"]], otherPar[["N"]],
+                       otherPar[["I0"]], ts, T)
+  
+  df <- data.frame(time = timeSeq(ts, T),
+                   SInR = diff(sinr[,"R"])/ts,
+                   SIgR = diff(sinner[,"R"])/ts)
+  #flow =diff(sinr[,"R"])
+  #mu <- sum(flow*df$time)
+  #S <- sum(flow*(df$time)^2)
+  #numkappa <- S/mu^2 - 1
+  
+  ggplot(df, aes(x = time)) + 
+    geom_line(aes(y = SInR, color = "SInR"), linewidth = 1) + 
+    geom_line(aes(y = SIgR, color = "SIgR"), linewidth = 1) + 
[…]

### 2024-03-22 First_draft/mytest1.R
Add test plot code

@@ -1,0 +1,11 @@
+library(ggplot2)
+library(shellpipes)
+startGraphics()
+
+df <- data.frame(A = seq(1,10), B = seq(1,10)^2)
+
+ggplot(df, aes(x = A, y = B)) +
+geom_line() + 
+labs(title = "This is a test plot")
+
+# not working qswl....

### 2024-03-22 First_draft/mytest2.R

@@ -1,0 +1,36 @@
+library(shellpipes)
+startGraphics()
+library(pseudoErlang)
+
+# loadEnvironments()
+######################################################################
+
+β <- 0.2
+D <- 10
+n <- 2
+μ <- 0.01  
+N <- 10000
+I0 <- 10
+ts <- 1
+T <- 200
+
+arp <- 0.9
+nbs <- 1000
+
+######################################################################
+# hard test (large kappa) 
+
+kappa <- 0.8 # maximum = 0.97
+nfix <- 12
+
+sigr <- sinnerFlow(β, D, kappa, nfix, μ, N, I0, ts, T)
+df <- simObs(sigr, arp, nbs, seed = 72)
+
+nfit <- 12
+startPar <- list(logβ=-1, logD=2, logkappa=-0.1)
+fixedPar <- list(n = nfit, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T) 
+fitW <- simplFit(startPar, fixedPar, df, sir.nll.g)
+#print(fitW$fit) 
+plotFit(fitW, df, fixedPar, type="SIgR", title = paste0("PE (n=12) -> PE (n=12) | kappa = " , kappa))
+
+

### 2024-03-22 figures/mytest2.R
Moving stuff around!

@@ -1,8 +1,13 @@
+library(pseudoErlang)
+library(deSolve)
+library(bbmle)
+library(ggplot2)
+
 library(shellpipes)
 startGraphics()
-library(pseudoErlang)
 
-# loadEnvironments()
+loadEnvironments()
+
 ######################################################################
 
 β <- 0.2

### 2024-03-22 Code/tempfunc.R
modified:   Code/tempfunc.R

@@ -1,4 +1,5 @@
 
+library(shellpipes)
 
 simObs <- function(sinr, arp, nbs, seed=NULL) {
   if(!is.null(seed)) set.seed(seed)
@@ -96,5 +97,6 @@ plotTrace <- function(fitW) {
 
 ######################################################################
 
+saveEnvironment()
 
 

### 2024-03-22 figures/mytest1.R
changed mytest1

@@ -6,6 +6,6 @@ df <- data.frame(A = seq(1,10), B = seq(1,10)^2)
 
 ggplot(df, aes(x = A, y = B)) +
 geom_line() + 
-labs(title = "This is a test plot")
+labs(title = "THIS is a test plot")
 
 # not working qswl....

### 2024-03-25 thesis/draft.tex
new file:   draft.tex

--- 2024-01-12 docs/Midyear/Mid_year_report.tex
@@
 \lhead{Ningrui Xie}
-\chead{BIO 4C12 Midyear Report}
+\chead{Thesis First Draft}
 \rhead{ID: 400323416}
@@
 \renewcommand{\headrulewidth}{0pt}
+\newcommand{\perday}{\ensuremath{\mathrm{/day}}}
 
 \begin{document}
+
 \title{Mathematical Approaches for Simulating Epidemic Progression: Addressing Limitations of the Linear Chain Trick in ODE Models}
@@
 \\McMaster University\\Hamiltion, Ontario, Canada L8S 4K1}
-\date {January 1, 2024}
+\date {March 22, 2024}
 \maketitle
@@
 \section{Abstract}
-The SIR framework, because of its simplicity and ability to describe dynamic changes, is widely used in infectious disease modeling.
-One issue with the traditional SIR model is its implicit assumption of an exponentially distributed infectious stage duration, which often diverges from real-world scenarios.
-A common solution is the application of the Linear Chain Trick (LCT), wherein the single infectious stage is divided into several substages, each following an exponential distribution.
-The incorporation of LCT results in an Erlang distributed stage duration.
-However, despite its widespread application, LCT presents challenges, including the complex determination of substage numbers and rigid parameter values.
-Here, we propose a novel approach, maintaining a fixed number of substages but employing a geometrically distributed substage transition rate.
-This method, while akin to LCT in its foundational principles, offers greater flexibility in defining the stage duration shape.
-In contrast to the recently introduced Generalized Linear Chain Trick (GLCT), which broadens the scope to encompass a wider range of phase-type distributions, our method provides a more straightforward and efficient means to achieve model flexibility.
-Although our discussion focuses on epidemic modeling, the potential applications of this approach extend to various forms of dynamic modeling, offering advantages in terms of computational efficiency, ease of parameter estimation, and adaptability to diverse modeling scenarios.
 
 \section{Introduction}
-Ordinary differential equations (ODEs) are a popular choice for mathematical modeling in biology due to their ability to accurately represent the dynamic processes of systems over time.
-ODEs provide a powerful framework for describing phenomena such as population dynamics \cite{salisbury2011mathematical}\cite{sego2021generation}, the spread of infectious diseases \cite{Anderson1991}\cite{Diekmann2000}\cite{Feng2016}, and cell proliferation (e.g., cancer) \cite{baker1998modelling}\cite{jarrett2018mathematical}, which are characterized by continuous state changes.
-The inherent flexibility and simplicity of ODEs make them not only easy to solve and understand but also easy to manipulate and accommodate to various situations.
+%Ordinary Differential Equations (ODEs) are a popular choice fo...
+\textbf{This is a test file!}
 
-In the field of epidemiology, ordinary differential equations (ODEs) serve as a powerful tool to characterize the progression of diseases and to make predictions.
-A commonly used model is the SIR compartmental model \cite{Anderson1991}\cite{kermack1927contribution}, which divides the population into three states: Susceptible, Infectious, and Recovered.
-In this model, a susceptible individual, upon contact with the infected population at a rate $\beta$, transitions immediately to the infectious state.
-An infected individual then moves to the recovered state at a recovery rate of $\gamma$.
-However, one major limitation of this model is its implicit assumption that the duration of the infectious stage follows an exponential distribution.
-This can be seen in a simplified equation modeling the dynamics of the infected population
-\begin{align*}
-    \frac{dI}{dt} = -\gamma I
-\end{align*}
-Solving this equation explicitly yields
-\begin{align*}
-    I(t) = I_0 e^{-\gamma t}
-\end{align*}
-which implies a stochastic state transition model where the time spent in the infectious stage $I$ is exponentially distributed with a mean stage duration of $\frac{1}{\gamma}$.
-This dwell time distribution suggests that an individual is most likely to recover immediately after infection, which is often not realistic.
-In reality, a delay between the successful infection of an individual and the recovery from the disease is inevitable, as shown in studies conducted by Greenhalgh and Day (2017) \cite{greenhalgh2017time} using measles outbreak data.
+This is a citation \cite{Feng2016}
 
-The constraint of the basic ODE framework is that it primarily offers a singular approach to assuming specific dwell time distributions: the first event time distribution of a nonhomogeneous Poisson process, including the exponential distribution as a special case \cite{hurtado2019generalizations}.
-This inflexibility in specifying probability distributions to describe the duration individuals spend in a given state is a nonnegligible shortcoming of the SIR model.
+Another citation \cite{Anderson1991}
 
-It is important to note that even when the mean is constant, the shape of the stage duration distributions (i.e., the timing of stage transitions) can significantly impact the dynamics and the outcomes of predictive models in applied settings \cite{krylova2013effects}\cite{keeling2002understanding}\cite{wearing2005appropriate}\cite{nguyen2008noise}.
-Consequently, it becomes crucial to consider the type of distribution characterizing an individual in a specific state when constructing these models.
-A standard approach for developing models that incorporate different stage duration distributions is through the use of integro-differential equations (IDEs) or integral equations (IEs) \cite{hurtado2019generalizations}\cite{kermack1927contribution}\cite{hethcote1980integral}.
-IDEs facilitate the extension of the SIR model to include arbitrary distributions for the duration of infectiousness, as discussed in various sources \cite{feng2000endemic}\cite{hethcote1980integral}\cite{ma2006generality}.
-This approach allows for the capture of the 'memory effect', where the current rate of change in each compartment of the model (Susceptible, Infectious, Recovered) is influenced by the entire history of the epidemic, rather than relying on a constant rate assumption.
+Here's a simple plot (with no dependence)
 
-However, while integro-differential equations enhance the model's flexibility, they are often challenging to formulate, analyze mathematically, and simulate, especially when compared to ordinary differential equations \cite{krylova2013effects}\cite{hurtado2019generalizations}\cite{burton2005volterra}.
-There have been efforts to simplify IDEs into systems of ODEs \cite{macdonald2008biological}\cite{ponosov2004thew}\cite{burton2005volterra}\cite{goltser2013reducing}\cite{diekmann2018finite}, with the most notable example being the reduction of IDEs that assume Erlang-distributed stage durations \cite{hurtado2019generalizations}.
-Unlike the exponential distribution, a more biologically plausible assumption might be a probability distribution where the likelihood of exiting the infectious stage starts small, increases as it nears the mean duration of infection, and eventually diminishes to zero as time extends indefinitely \cite{sartwell1966incubation}\cite{bailey1954statistical}.
-Due to its central tendency towards the mean value, the Erlang distribution, defined as a gamma distribution with an integer shape parameter, is a more realistic choice for depicting the duration of the infectious stage.
-The integration of the Erlang distribution into the ODE system is facilitated by the Linear Chain Trick (LCT) \cite{macdonald1978time}\cite{smith2011introduction}.
-In this approach, the single infected state is divided into $n$ substages, each following an exponential distribution with a rate of $n\gamma$.
-This technique is based on the principle that the sum of a series of independent exponentially distributed random variables follows a gamma distribution \cite{krylova2013effects}\cite{therrien2018probability}.
-When the LCT is applied to the SIR model, the number of substages $n$ determines the shape of the stage duration distribution.
-As $n$ increases, the variance decreases, resulting in a distribution that is more peaked and narrow around the mean value $\frac{1}{\gamma}$.
-As $n$ approaches infinity, the distribution converges to a delta function (a deterministic value, implying that all individuals who become infectious at time $t$ recover at exactly time $t + \frac{1}{\gamma}$), and the system evolves into a delay differential equation \cite{krylova2013effects}\cite{hethcote1980integral}.
-
-The Linear Chain Trick (LCT) is extensively employed in modeling biological processes, not only in quantifying disease progression \cite{lloyd2001destabilization}\cite{lloyd2001realistic}, but also in analyzing virus \cite{lloyd2001dependence}\cite{kakizoe2015method} and population \cite{cushing2013integrodifferential} dynamics.
-Studies have demonstrated that switching the stage duration distribution from an exponential to an Erlang distribution leads to predictions that align more closely with real data points \cite{kakizoe2015method}.
-However, despite the increased realism they bring to modeling disease processes, employing LCT in ODE models presents several limitations.
-One primary challenge is determining the appropriate number of substages $n$ to divide, which is a cumbersome and time-consuming process.
-Multiple models with varying substages must be constructed and subjected to repeated data-fitting processes to identify the optimized parameter value.
-Prior research has investigated approaches for estimating the number of states in the linear chain, employing techniques such as profile likelihood \cite{raue2009structural} and model reduction \cite{maiwald2016driving}.
-Yet, a comparative study of these methods indicates that the typical quality of experimental data is insufficient for reliably estimating the chain length using these approaches \cite{hauber2020estimating}.
-Another inherent limitation of using LCT is the inflexibility of the model's parameters.
-To integrate $n$ independent exponentially distributed random variables into the ODE system, $n$ must be an integer.
-As the subdivision process is done without altering the overall mean of the stage, the rate of each substage $n\gamma$ is also constrained to discrete values.
-These limitations significantly restrict the adaptability and practical application performance of the model.
-
-The recently proposed Generalized Linear Chain Trick (GLCT) enhances the Linear Chain Trick (LCT) by enabling models to incorporate a broader phase-type family of distributions \cite{hurtado2019generalizations}\cite{hurtado2021building}\cite{bladt2017phase}, which includes exponential, Erlang, hypoexponential, and Coxian distributions, thus allowing for greater flexibility and the ability to capture more complexity.
-The GLCT achieves this by creating a higher-dimensional Markov chain, enabling a more accurate representation of the complex nature of dwell time distributions.
-However, this process can be mathematically heavy and challenging to interpret.
-Additionally, this approach may lead to a increase in the number of variables or states within the model, potentially escalating computational demands.
-Therefore, finding a simpler and more efficient method to enhance the flexibility of the ODE model remains essential.
-
-In this study, we focused on one particular direction: generalizing the Erlang distribution of stage durations currently present in the model.
-The goal was to allow for a more diverse range of shapes, achieved by eliminating the constraints imposed by the shape parameter ($\kappa$).
-The value of $\kappa$ in the SInR model was limited to a set of discrete values, as detailed below.
-To address this limitation, as well as those previously mentioned, we propose a novel approach.
-Instead of using a variable number of substages $(n)$ with a constant rate $(\gamma_i = n\gamma)$ for all exponentially distributed substages, we fix $n$ to a specific integer value and use a series of rates that follow a geometric pattern.
-This series is represented by the formula $\gamma_i = ar^{i-1}$, governed by parameters $a$ and $r$.
-Consequently, the parameter of interest shifts from discrete values $(n$ and $ng)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability.
-
-To illustrate this point, consider that the Gamma distribution can be parameterized in terms of its mean $(M)$ and shape parameter (quadratic coefficient of variation, $\kappa$), which determine the overall shape and position of the distribution.
-As mentioned above, in subdividing a single stage into n substages, each exponentially distributed substage must have a rate of $n\gamma$ to maintain the overall rate of that stage.
-While the mean of the Erlang distribution in the ODE system can be chosen freely, $\kappa$ is constrained to discrete values, shown as follows:
-\begin{align*}
-    Mean: \quad &\sum_{i=1}^{n_f} \frac{1}{n\gamma} = \frac{1}{\gamma}\\
-    Variance: \quad &\sum_{i=1}^{n_f} \frac{1}{n^2\gamma^2} = \frac{1}{n\gamma^2} \\
-    \kappa: \quad &\frac{\sigma^2}{M^2} = \frac{1}{n} \quad n \in \mathbb{N}
-\end{align*}
-Our approach, using a geometric series of rates, addresses this issue by allowing $\kappa$ to take any value within the range $[\frac{1}{n}, 1)$.
-A detailed explanation and the derivation process are provided in section $4.2$.
-Another advantage of this method is that it enables a fixed model structure where only the parameters change, eliminating the need to determine optimal substage numbers through the construction of multiple models, a process currently required in the ODE model that employs the LCT.
-
-Throughout this research, we primarily focused on the application of the LCT in the epidemic SIR model, which we refer to as the SI$^n$R model, where $n$ represents the number of substages divided.
-The objective of our study is to develop and evaluate the method of using a geometrically distributed substage rate as a replacement for the constant rate, aiming to enhance flexibility and efficiency.
-We have named the model that utilizes this method the SIgR model (temporarily).
-We anticipate that this approach will not only be computationally efficient but also provide a better fit to real-world data.
-In the following sections, we first aim to align the infectious stage duration distribution shapes generated by the SI$^n$R and SIgR models.
-Once we have confirmed that our proposed model can effectively substitute the SInR model, we proceed to allow the SIgR model to include stage duration distributions with characteristics unattainable by the SI$^n$R model.
-We then invert the process: given a desired mean and $\kappa$, we develop the computational system to determine the parameter values $(a$ and $r)$ that enable the model to produce a stage duration distribution with these properties.
-Subsequently, we will conduct empirical validation of our model through data fitting.
-This evaluation will cover aspects such as the model’s consistency with dataset and its predictive accuracy in forecasting future trends.
-We predict that this approach will enhance the model’s adaptability and practicality, thus paving the way for the development of user-friendly software tools that empower modelers to efficiently select and employ the most suitable models for describing and predicting future epidemic scenarios.
-
-\section{Method}
-Several assumptions have been made for the unforced SIR model (Susceptible, Infectious, and Recovered) that our study is based on: 
-\begin{enumerate}
-    \item \textbf{Closed Population}: A constant population size with equal birth and death rates and no disease-related mortalities.
-    \item \textbf{Homogeneous mixing}: Individuals have an equal chance of coming into contact with any other individual in the population. 
-    \item \textbf{Ignoring Latency}: Individuals move immediately from a susceptible state to an infected state, ignoring any incubation period associated with the disease.
-    \item \textbf{Lifelong immunity}: Once an individual recovers, they do not become susceptible or infectious again.
-\end{enumerate}
-We also assume that the state variables have been scaled by the total population size $N$ so that $S + I + R = 1$.
-The SIR model can be expressed as a simple system of nonlinear ordinary differential equations:
-\begin{align}
-    \frac{dS}{dt} &= \mu - \beta SI - \mu S \\
-    \frac{dI}{dt} &= \beta SI - \gamma I - \mu I \\
-    \frac{dR}{dt} &= \gamma I - \mu R
-\end{align}
-where $\mu$ is the birth and natural mortality rate, $\beta$ is the disease transmission rate, and $\gamma$ is the recovery rate.
-The mean recovery time for an infected individual is $\frac{1}{\gamma}$.
-
-The idea of dividing the infectious stage into $n$ sub-stages exploits the fact that the sum of a sequence of independent, exponentially distributed random variables follows an Erlang distribution (i.e.
-gamma distributed with integer shape parameter) \cite{therrien2018probability}.
-Thus, we begin by confirming the uniformity of the results generated by simulating the SI$^n$R model with the Erlang distribution, which has the following form:
-\begin{equation*}
-    f(x; n,n\gamma) = \frac{(n\gamma)^n}{(n-1)!} x^{n-1} e^{-n\gamma x} \quad x>0, n \in \mathbb{N}
-\end{equation*}
-
-By breaking a single stage into $n$ substages, each exponentially distributed with a rate of $n\gamma$ ($\gamma_i = n\gamma$), the SI$^n$R model can be represented by a system of ODEs:
-\begin{align*}
-    \frac{dS}{dt} &= \mu - \beta SI - \mu S \\
-    \frac{dI_1}{dt} &= \beta SI - (n \gamma + \mu) I_1 \\
-    \frac{dI_2}{dt} &= n\gamma I_1 - (n \gamma + \mu) I_2 \\
-    &\vdots \\
-    \frac{dI_n}{dt} &= n\gamma I_{n-1} - (n \gamma + \mu) I_n \\
-    \frac{dR}{dt} &= n \gamma I_n - \mu R
-\end{align*}
-
-Note that this decomposition process will not change the mean stage duration, as the sum of the means of the $n$ exponentially distributed random variables still equals $\frac{1}{\gamma}$
-\begin{align*}
-    \frac{1}{\gamma} = \sum_{i=1}^{n} \frac{1}{n\gamma}
-\end{align*}
-
-For consistency, we focused exclusively on the dynamics of the recovered compartment.
-The simulation results generated by the SI$^n$R model represent the cumulative distribution of the recovered population.
-Therefore, we need to take the derivative to convert it into a probability density function, allowing for comparison with the Erlang distribution.
-After confirming that the stage duration distribution generated by the SI$^n$R model aligns with a gamma distribution with an integer shape parameter, we constructed the SIgR model.
-This model has a fixed number of substages $n_f$ and features substage rate that are geometrically distributed ($\gamma_i = ar^{i-1}$), shown as follow
-\begin{align*}
-    \frac{dS}{dt} &= \mu - \beta SI - \mu S \\
-    \frac{dI_1}{dt} &= \beta SI - (a + \mu) I_1 \\
-    \frac{dI_2}{dt} &= a I_1 - (ar + \mu) I_2 \\
-    &\vdots \\
-    \frac{dI_n}{dt} &= ar^{n-1} I_{n-1} - (ar^n + \mu) I_n \\
-    \frac{dR}{dt} &= ar^n I_n - \mu R
-\end{align*}
-
-We aim to determine if altering the parameter $a$ and $r$ is more effective than changing the number of substages.
-Our starting point involves keeping the mean stage duration $(M)$ constant while varying $r$ to explore the possibility of a match between the SI$^n$R and SIgR models.
-This match will be assessed both by the mean $(M)$ and by the shape parameter, which is the coefficient of variation squared $(\kappa = \frac{\sigma^2}{M^2})$.
-
-After determining the potential match, we have formulated the formulas for $M$ and $\kappa$ in terms of $a$ and $r$.
-Utilizing the geometric distribution of rates in the ODE model allows for independent changes in $M$ and $\kappa$.
-This enables us to achieve the desired value of $\kappa$ without altering the mean, and vice versa.
-Our goal is to reverse the method: starting with given values of $M$ and $\kappa$, we aim to find specific values of $a$ and $r$ that will let the model generate the stage duration distribution that satisfies these properties.
-Due to the complexity introduced by the high powers in the formula, solving it explicitly is challenging.
-To address this, we have developed a system of functions in the R programming language, employing the uniroot() function to obtain a numerical solution for r.
-We then reintegrate this parameter value into the SIgR model and compare it with the SInR model, using the same $M$ and $\kappa$ values.
-Subsequently, the mean and the shape parameter of both simulation results are calculated numerically to confirm the validity of the match.
-
-After confirming that the newly proposed model can effectively replace the SI$^n$R model by encompassing the same infectious stage duration distributions, we will proceed to empirically test the enhanced performance of the model through data fitting.
-This process will be divided into two phases and will be conducted in R using the \textbf{bbmle} package.
-As the starting point, we will employ a deterministic model and assume only observation noise.
-Our first step involves coding to generate a 'fake' dataset, treating the disease incidence time series from the SIgR model as the actual case, and introducing deviations (noises) into them using a mean reporting ratio and a negative binomial dispersion parameter.
-The model parameters will then be estimated using the Maximum Likelihood Estimation (MLE) method, applied to the fake dataset.
-The objective here is to accurately recover the original parameter values that were used to generate the disease incidence data.
-In the next phase of the study, we will fit the model using an real-world dataset.
-We plan to compare the fitting results obtained from both the SI$^n$R and SIgR models.
-This will allow us to quantitatively assess the efficacy of our proposed method in improving the model's accuracy in reflecting actual epidemiological scenarios.
-
-\section{Results}
-In Figure 1, we present the comparison plots between the infectious stage duration distribution generated by the SI$^n$R model with the Erlang distribution.
-In this model, the mean recovery time is set to be 10 days $(\gamma = \frac{1}{10})$ and the infectious stage is divided into $n$ substages $(n = 4, 8, 12, 18)$, with each substage having a rate of $\frac{n}{10}$.
-The corresponding shape and scale parameters for the gamma distribution are $n$ and $\frac{n}{10}$, respectively.
-The plot demonstrates a perfect match between the two distributions, despite being generated by different methods.
-In the subsequent sections, we will refer to these two distributions interchangeably.
 \begin{figure}[h]
     \centering
-    \includegraphics[width= 0.9\textwidth]{4.png}
-    \caption{Probability Distribution Comparison Between Erlang and SI$^n$R Model}
+    \includegraphics{figures/mytest1.Rout.pdf}
+    \caption{Mytest1 plot}
 \end{figure}
 
-\subsection{Possible match between SI$^n$R and SIgR}
-In the initial step following the successful construction of the SIgR model, which employs a fixed number of substages and a geometrically distributed substage rate, we want to evaluate the theoretical possibility of a match between the stage duration distribution produced by the SIgR model and that of the traditional SInR model.
-This match is defined in terms of two key parameters: the mean $(M)$ and the shape parameter $(\kappa)$.
+Another plot (depend on tempfunc.R)
 
-The substage number of the SIgR model $(n_f)$ has been fixed at 12, and the mean stage duration for the entire infectious period is set to 10 days.
-We examine the effect of using 2, 4, and 8 infectious substages in the SI$^n$R model $(n = 2, 4, 8)$ on the values of $r$ required for the SIgR model to align formulaically with the shape parameter $(\kappa)$.
-In Figure 2, the plots on the left illustrate the changes in stage duration distribution in SIgR model resulting from varying $r$ within a range that includes the formulaic match for $\kappa$.
-The black curve represent duration distribution from the SI$^n$R model with the associated $n$.
-Plots on the right depict the variations in $\kappa$, calculated using the corresponding r values, where the red dashed line is the target value of $\kappa$.
-It is evident from the plots that, within the specified range of r, there is a concurrence in both the mean and shape parameter values between the SInR and SIgR model.
-These findings suggest that the results generated by the traditional mode are replicable using the model we propose.
-
-\begin{figure}[h!]
+\begin{figure}[h]
     \centering
-
-    \begin{subfigure}{0.6\linewidth}
-       \includegraphics[width=\linewidth]{4.1(n=2).png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.6\linewidth}
-       \includegraphics[width=\linewidth]{4.1(n=4).png}
-    \end{subfigure}
-
-    \begin{subfigure}{0.6\linewidth}
-        \includegraphics[width=\linewidth]{4.1(n=8).png}
-     \end{subfigure}
-
-    \caption{Adjusting $r$ to Evaluate Match While Preserving Mean Value}
-    \label{fig:all_plots}
+    \includegraphics{figures/mytest2.Rout.pdf}
+    \caption{Mytest2 plot}
 \end{figure}
 
-\subsection{Simple Equations Facilitating Independent and Flexible Changes in Mean and $\kappa$}
-In the SIgR model, the calculation of the Mean $(M)$ and Variance $(\sigma^2)$ for the infectious stage duration distribution is accomplished by adding the corresponding properties of each substage, as follows:
-\begin{align}
-    Mean: \quad  &\sum_{i=1}^{n_f} \frac{1}{\gamma_i} = \frac{\frac{1}{a} (\frac{1}{r^n} - 1)}{\frac{1}{r} -1} \\
-    Variance: \quad &\sum_{i=1}^{n_f} \left(\frac{1}{\gamma_i} \right)^2 = \frac{\frac{1}{a^2} (\frac{1}{r^{2n}} - 1)}{\frac{1}{r^2} -1}
-\end{align}
-The shape parameter $\kappa$, determined by $\frac{\sigma^2}{M}$, can thus be simplified to:
-\begin{align}
-    \kappa = \frac{(1+r^{n_f}) (1-r)}{(1+r) (1-r^{n_f})}
-\end{align}
-This simplification reveals that the value of $\kappa$ is exclusively dependent on the parameter $r$, thereby granting the model independence in property changes.
-This independence allows for the alignment of $\kappa$ without altering the Mean, achievable through adjustments to parameter $a$.
-Similarly, the Mean can be scaled to a desired value without impacting $\kappa$.
+Maybe another plot from a different folder:
 
-Importantly, it should be noted that the value of $\kappa$ is not affected by changes in the number of substages, as $n_f$ is fixed.
-Consequently, $\kappa$ can be any real number within the range $[\frac{1}{n_f}, 1)$, thereby significantly enhancing the model's flexibility.
-In contrast, in the SI$^n$R model, the value of $\kappa$ depends on how the substages are partitioned.
-Following the same steps of calculation, we obtain: $\kappa = \frac{1}{n}$.
-The shape parameter $\kappa$ in the SI$^n$R model is therefore restricted to a set of discrete values:
-\begin{align*}
-    \kappa \in \left\{1, \frac{1}{2}, \frac{1}{3}, \dots \right\}
-\end{align*}
-This limitation potentially reduces the capability of the SI$^n$R model to accurately reflect real-world scenarios.
-
-\subsection{Identifying Parameters $a$ and $r$ to Align SI$^n$R and SIgR with Given Mean and $\kappa$}
-Utilizing formulas (4) and (6), we have developed a computational system to reverse-engineer the process.
-The parameters $a$ and $r$, generated by this system, were reintegrated into the SIgR model.
-This enable us to compare the stage duration distribution with that of the SI$^n$R model, which possesses identical mean and shape parameters, as depicted in Figure 3.
-Although the formulaic $M$ and $\kappa$ are computed to be the same, we also conducted a numerical calculation of both properties, with results presented in Table 1.
-This analysis confirms that a match is achieved at both the formulaic and numerical levels.
-The visual discrepancies observed between the two distributions in Figure 3 can be considered negligible, since both distributions serve as theoretical representations of real-world scenarios and do not require exact correspondence.
 \begin{figure}[h]
     \centering
-    \includegraphics[width= \textwidth]{4.3.1.png}
-    \caption{Theoretical Alignment of Mean and Shape Parameters in SI$^n$R and SIgR Models with Calculated Parameter Values $a$ and $r$}
+    \includegraphics{figures/test.Rout.pdf}
+    \caption{test plot}
 \end{figure}
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{ l | c | r }
-      \hline
-      Substages & Mean & $\kappa$ \\
-      \hline
-      n = 2 & 10.000000 & 0.4999971 \\
-      n = 4 & 10.000000 & 0.2500002 \\
-      n = 8 & 10.000 & 0.125 \\
-      n = 12 & 10.000000 & 0.0999986 \\
-      \hline
-    \end{tabular}
-    \caption{Numerically Calculated Parameter Values From the SIgR Model}
-\end{table}
-
-By incorporating the infectious stage duration distribution back into the entire SI$^n$R and SIgR models, we also present the dynamics of the infectious population density generated directly from the simulations.
-In Figure 4, using $n=2$ and $n=4$ for the SI$^n$R model and corresponding $a$ and $r$ values ( $r_2=2.9999604, a_2=0.1500007, r_4 = 1.6626998, a_4 = 0.2503359$) for the SIgR model, a close match is observed between the results generated by the two different methods.
-This indicates that the results of the SInR model are reproducible using the SIgR model, but with enhanced flexibility in parameter value changes.
-\begin{figure}[h]
-    \centering
-    \includegraphics[width= \textwidth]{4.3.2.png}
-    \caption{Simulation Comparison of the Population Dynamics Between SI$^n$R and SIgR Model with Equivalent Mean and Shape Parameter}
-\end{figure}
-
-\subsection{Enhanced Flexibility in Controlling Stage Duration Distribution Shapes}
-Besides offering a simpler approach to constructing an epidemic model with efficacy comparable to the SInR model, the SIgR model further enhances our ability to manipulate the shape of the stage duration distribution.
-This is achieved by allowing the selection of any $\kappa$ value within the range $(\frac{1}{n_f}, 1)$.
-Figure 5 showcases the probability distribution plots generated by the SIgR model, given a desired mean $(M=10)$ and selected $\kappa$ value $(\kappa = 0.25, 0.45, 0.65, 0.85)$.
-This chosen $\kappa$ value (except 0.25) is unattainable in the SI$^n$R model, distinguishing the SIgR model's unique capabilities.
-The new model adeptly incorporates a wider array of distribution shapes into the SIR framework.
-It emerges as a versatile tool for capturing a more extensive range of dynamic patterns in disease spread and control.
-\begin{figure}[h]
-    \centering
-    \includegraphics[width= \textwidth]{4.4.2.png}
-    \caption{Generating Stage Duration Distributions with Custom Mean and Shape Parameters Using the SIgR Model}
-\end{figure}
-
-\section{Discussion}
-SIR compartmental models are widely utilized in modeling infectious diseases, but they encounter a significant limitation: the implicit assumption that the time an individual spends in the disease stage follows an exponential distribution.
-To address this issue, a common method involves subdividing the infectious stage into $n$ number of substages, each substage is exponentially distributed with a constant rate.
-This approach is also known as the Linear Chain Trick (LCT).
-LCT strategy facilitates an Erlang-shaped stage duration distribution with increased reality.
-However, this model also faces limitations, including cumbersome parameter optimization processes and the constraint of discrete parameter values, which can affect its performance.
-
-In this study, we propose the SIgR model, which features a fixed number of infectious substages and a geometric distribution rate for these substages, as an alternative to the SInR model that incorporates LCT.
-In the results section above, we demonstrated that adjusting the parameters for the geometric series $a$ and $r$ allows us to replicate the simulation results generated by the SInR model.
-Moreover, our model offers enhanced flexibility in shaping the distribution of the infectious stage duration, as it allows for the selection of mean $(M)$ and shape parameters $(\kappa)$ of the distribution such that $M \in \mathbb{R}^+$ and $\kappa \in [\frac{1}{n}, 1)$.
-Additionally, the fixed structure of our model simplifies the process of finding optimized parameter values, requiring only a one-time process during data fitting.
-This simplification streamlines the development of computer code for use with ODE solvers in R, ultimately saving time and reducing complexity.
-
-Future work will primarily focus on fitting data to the model we have proposed.
-This will include both synthetic data generated through simulation and actual real-world data.
-The results of this fitting process will be compared with those obtained from the SInR model.
-Later on, we plan to incorporate process error considerations to increase the robustness of our model.
-
-The anticipated outcome of this research is the development of a model that not only exceeds current models in terms of predictive accuracy but also maintains simplicity in implementation.
-The core design of our model is intended to be easily adaptable across various state transition models.
-This adaptability is vital for its application in diverse domains, ranging from epidemiology to system dynamics.
-The idea is designed to be extensible, capable of accommodating a range of distribution types and parameter sets.
-Such versatility makes it a powerful tool for modeling a wide array of complex systems, with the capability to adjust and evolve in response to new data and insights.
-
-In conclusion, this study suggests that our method could simplify and enhance the fitting of models with flexible time distributions within an Ordinary Differential Equation (ODE) framework.
-This research has the potential to enhance model performance and to develop user-friendly tools for effective epidemic modeling and prediction.
 

### 2024-03-25 figures/mytest1.R
test make by changing mytest1

@@ -6,6 +6,6 @@ df <- data.frame(A = seq(1,10), B = seq(1,10)^2)
 
 ggplot(df, aes(x = A, y = B)) +
 geom_line() + 
-labs(title = "THIS is a test plot")
+labs(title = "THIS is a test plot (TESE MAKE")
 
 # not working qswl....

### 2024-03-25 figures/mytest1.R
some changes for plot position (√)

@@ -1,11 +1,12 @@
 library(ggplot2)
 library(shellpipes)
-startGraphics()
+startGraphics(height=4)
+
 
 df <- data.frame(A = seq(1,10), B = seq(1,10)^2)
 
 ggplot(df, aes(x = A, y = B)) +
 geom_line() + 
-labs(title = "THIS is a test plot (TESE MAKE")
+labs(title = "THIS is a test plot (TESE)")
 
 # not working qswl....

### 2024-03-26 figures/mytest1.R
change plot size (crazy)

@@ -1,6 +1,6 @@
 library(ggplot2)
 library(shellpipes)
-startGraphics(height=4)
+startGraphics(height=1, width=1)
 
 
 df <- data.frame(A = seq(1,10), B = seq(1,10)^2)

### 2024-03-26 thesis/draft.tex
some changes for plot position (√)
temp storage

@@
 \section{Introduction}
-%Ordinary Differential Equations (ODEs) are a popular choice fo...
-\textbf{This is a test file!}
+Ordinary differential equations (ODEs) are extensively utilized for mathematical modeling in the biological sciences, offering a robust framework to accurately describe the temporal dynamics of systems.
+Their capacity to model continuous state changes makes them essential in capturing complex biological phenomena such as population dynamics \cite{salisbury2011mathematical}\cite{sego2021generation}, epidemiological patterns of infectious diseases \cite{Anderson1991}\cite{Diekmann2000}\cite{Feng2016}, and cellular processes like proliferation in oncogenesis \cite{baker1998modelling}\cite{jarrett2018mathematical}.
+The inherent flexibility and simplicity of ODEs make them not only easy to solve and understand but also easy to manipulate and accommodate to various situations.
+
+In the field of epidemiology, ordinary differential equations (ODEs) serve as a crucial tool for characterizing disease progression and making predictions.
+During the COVID-19 pandemic, they were extensively utilized to guide policy-making, especially in disease control and public health interventions \cite{thompson2020epidemiological}.
+A commonly used model is the SIR compartmental model \cite{Anderson1991}\cite{kermack1927contribution}, which divides the population into three states: Susceptible $(S)$, Infectious $(I)$, and Recovered $(R)$.
+In the model, a susceptible individual, upon contact with the infected population at a rate $\beta$ $(\frac{1}{time})$, transitions immediately to the infectious state.
+An infected individual then moves to the recoverd state at a recovery rate of $\gamma$ $(\frac{1}{time})$.
+However, one major limitation of this model it its implicit assumption that the duration of the infectious stage follows an exponential distribution.
+This can be observed in a simplified equation that models the dynamics of the infected population.
+It assumes the entire population was infected at the outset, and individuals can only transition out of the infectious stage at a rate of $\gamma$:
+\begin{align}
+    \frac{dI}{dt} = -\gamma I
+\end{align}
+Solving this equation explicitly yields
+\begin{align}
+    I(t) = I_0 e^{-\gamma t}
+\end{align}
+where $I_0$ represent the initial infected population.
+Formula (2) implies a stochastic state transition model where the time spend in the infectious $I$ is exponentially distributed with a mean stage duration of $\frac{1}{\gamma}$.
+This dwell time distribution suggests that an individual has the highest probability of recovering from the disease immediately after infection, which is often not realistic.
+In reality, a delay between the successful infection of an individual and the recovery from the disease is inevitable.
+The work conducted by Byrne AW et al.
+\cite{byrne2020inferred} presents the simulation of parameter distribution inferred for the duration of the infectious period, and the shape of this distribution closely resembles a gamma distribution.
+The study by Greenhalgh and Day (2017) \cite{greenhalgh2017time}, utilizing measles outbreak data, also demonstrates that the recovery rate from an infectious disease is not static but can vary based on a wide range of factors.
+
+It is important to note that even when the mean is constant, the shape of the stage duration distributions (i.e., the timing of stage transitions) can significantly impact the dynamics and the outcomes of predictive models in applied settings \cite{krylova2013effects}\cite{keeling2002understanding}\cite{wearing2005appropriate}\cite{nguyen2008noise}.
+Consequently, it becomes crucial to consider the type of distribution characterizing an individual in a specific state when constructing these models.
+A standard approach for developing models that incorporate different stage duration distributions is through the use of integro-differential equations (IDEs) or integral equations (IEs) \cite{hurtado2019generalizations}\cite{kermack1927contribution}\cite{hethcote1980integral}.
+IDEs facilitate the extension of the SIR model to include arbitrary distributions for the duration of infectiousness, as discussed in various sources \cite{feng2000endemic}\cite{hethcote1980integral}\cite{ma2006generality}.
+This approach allows for the capture of the 'memory effect', where the current rate of change in each compartment of the model (Susceptible, Infectious, Recovered) is influenced by the entire history of the epidemic, rather than relying on a constant rate assumption.
+
+However, while integro-differential equations increase the model's flexibility, they present challenges in formulation, mathematical analysis, and simulation, particularly in comparison to ODEs \cite{krylova2013effects}\cite{hurtado2019generalizations}\cite{burton2005volterra}.
+An alternative method, the Linear Chain Trick (LCT), overcomes some of these challenges, offering a simpler and more direct approach to modifying the shape of distributions \cite{macdonald1978time}\cite{smith2011introduction}.
+This approach subdivides the single infected state into n substages, each characterized by an identical exponential distribution with a rate of $n\gamma$.
+This technique is based on the principle that the sum of a series of independent exponentially distributed random variables follows an Erlang distribution \cite{krylova2013effects}\cite{therrien2018probability}, which is a special case of the gamma distribution with an integer shape parameter.
+By employing this method, the distribution of the infectious stage duration transitions from exponential to Erlang.
+The central tendency of the Erlang distribution towards its mean value offers the model a more accurate representation of real-world scenarios.
+
+When applying LCT into the SIR model, the number of substages $n$ determines the shape of the stage duration distribution.
+As $n$ increases, the variance decreases, resulting in a distribution that is more peaked and narrow around the mean value $\frac{1}{\gamma}$.
+As $n$ approaches infinity, the distribution converges to a delta function (a deterministic value, implying that all individuals who become infectious at time $t$ recover at exactly time $t + \frac{1}{\gamma}$), and the system evolves into a delay differential equation \cite{krylova2013effects}\cite{hethcote1980integral}.
+
+The Linear Chain Trick (LCT) is extensively employed in modeling biological processes, not only in quantifying disease progression \cite{lloyd2001destabilization}\cite{lloyd2001realistic}, but also in analyzing virus \cite{lloyd2001dependence}\cite{kakizoe2015method} and population \cite{cushing2013integrodifferential} dynamics.
+Studies have demonstrated that switching the stage duration distribution from an exponential to an Erlang distribution leads to predictions that align more closely with real data points \cite{kakizoe2015method}.
+However, despite the increased realism they bring to modeling disease processes, employing LCT in ODE models presents several limitations.
+One primary challenge is the inflexibility of the model's parameters.
+To integrate independent exponentially distributed random variables into the ODE system, $n$ must be an integer.
+As mentioned above, since the shape of the stage duration distribution is determined by $n$, it would also be constrained to discrete changes.
+Another inherent limitation of using LCT is the cumbersome and time-consuming process of determining the appropriate number of substage to divide.
+Multiple models with varying substages must be constructed and subjected to repeated data-fitting processes to identify the optimized parameter value.
+Prior research has investigated approaches for estimating the number of states in the linear chain, employing techniques such as profile likelihood \cite{raue2009structural} and model reduction \cite{maiwald2016driving}.
+Yet, a comparative study of these methods indicates that the typical quality of experimental data is insufficient for reliably estimating the chain length using these approaches \cite{hauber2020estimating}.
+These limitations significantly restrict the adaptability and practical application performance of the model.
+
+The recently proposed Generalized Linear Chain Trick (GLCT) enhances the Linear Chain Trick (LCT) by enabling models to incorporate a broader phase-type family of distributions \cite{hurtado2019generalizations}\cite{hurtado2021building}\cite{bladt2017phase}, which includes exponential, Erlang, hypoexponential, and Coxian distributions, thus allowing for greater flexibility and the ability to capture more complexity.
+The GLCT achieves this by creating a higher-dimensional Markov chain, enabling a more accurate representation of the complex nature of dwell time distributions.
+However, this process can be mathematically heavy and challenging to interpret.
+Additionally, this approach may lead to a increase in the number of variables or states within the model, potentially escalating computational demands.
+Therefore, finding a simpler and more efficient method to enhance the flexibility of the ODE model remains essential.
+
+In this study, we focused on one particular direction: generalizing the Erlang distribution of stage durations currently present in the model.
+The goals were to have a more effective way for parameter estimation and to allow for a more diverse range of shapes, achieved by eliminating the constraints imposed by the shape parameter ($\kappa$).
+To address these limitations, we proposed the Geometric Chain Trick (GCT).
+Instead of having a variable number of substages $(n)$ with a constant transition rate between them $(\gamma_i = n\gamma)$, we fix $n$ to a specific integer value and use a series of rates that follow a geometric pattern.
+This series is represented by the formula $\gamma_i = ar^{i-1}$, governed by parameters $a$ and $r$.
+This approach enables a uniform model structure with the parameter of interest transitioning from discrete values $(n$ and $ng)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability (details in Section 4.x).
+
+\section{Method}
+
+\section*{Results}
+
+\section{Discussion}
+
+%%%%%%%%
 
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
     \includegraphics{figures/mytest1.Rout.pdf}
-    \caption{Mytest1 plot}
+    \caption{Mytest1 plot (are you working??)}
 \end{figure}

### 2024-03-27 thesis/draft.tex
temp storage

@@
 
+To illustrate this point, consider that the Gamma distribution can be parameterized in terms of its mean $(D)$ and shape parameter (quadratic coefficient of variation, $\kappa$), which determine the overall shape and position of the distribution.
+As mentioned above, in subdividing a single stage into n substages, each exponentially distributed substage must have a rate of $n\gamma$ to maintain the overall rate of that stage.
+While the mean of the Erlang distribution in the ODE system can be chosen freely, $\kappa$ is constrained to discrete values, shown as follows:
+\begin{align*}
+    Mean: \quad &\sum_{i=1}^{n_f} \frac{1}{n\gamma} = \frac{1}{\gamma}\\
+    Variance: \quad &\sum_{i=1}^{n_f} \frac{1}{n^2\gamma^2} = \frac{1}{n\gamma^2} \\
+    \kappa: \quad &\frac{\sigma^2}{M^2} = \frac{1}{n} \quad n \in \mathbb{N}
+\end{align*}
+Our approach, using a geometric series of rates, addresses this issue by allowing $\kappa$ to take any value within the range $\left[\ \frac{1}{n}, 1 \right)$.
+A detailed explanation and the derivation process are provided in section $4.2$.
+Another advantage of this method is that it enables a fixed model structure where only the parameters change, eliminating the need to determine optimal substage numbers through the construction of multiple models, a process currently required in the ODE model that employs the LCT.
+
+Throughout this research, we primarily focused on the application of the LCT in the epidemic SIR model, which we refer to as the SI$^n$R model (or LCT model), where $n$ represents the number of substages divided.
+The objective of our study is to develop and evaluate the method of using a fixed model structure and a geometrically distributed substage rate as a replacement for the constant rate, aiming to enhance flexibility and efficiency.
+We have named the model that utilizes this method the GCT model.
+We anticipate that this approach will not only be computationally efficient but also provide a better fit to real-world data.
+In the following sections, we first analyze the model parameters by calculating the formulae for two essential properties: the mean and shape parameter.
+We then invert the process: given a desired $D$ and $\kappa$, we develop a computational system to determine the parameter values $(a and r)$ that enable the model to produce a stage duration distribution with these properties.
+The obtained parameter values are then used to compare the GCT model with the LCT model through simulation.
+Subsequently, we conduct empirical validation of our model through data fitting.
+This fitting confirmed that the GCT model enables the use of a uniform model structure with enhanced flexibility and practicality, thus paving the way for the development of user-friendly software tools that empower modelers to efficiently select and employ the most suitable models for describing and predicting future epidemic scenarios.
+
 \section{Method}
+Several assumptions have been made for the unforced SIR model (Susceptible, Infectious, and Recovered) that our study is based on: 
+\begin{enumerate}
+    \item \textbf{Closed Population}: A constant population size with equal birth and death rates and no disease-related mortalities.
+    \item \textbf{Homogeneous mixing}: Individuals have an equal chance of coming into contact with any other individual in the population. 
+    \item \textbf{Ignoring Latency}: Individuals move immediately from a susceptible state to an infected state, ignoring any incubation period associated with the disease.
+    \item \textbf{Lifelong immunity}: Once an individual recovers, they do not become susceptible or infectious again.
+    \item \textbf{Fixed rate of transmission and recovery}: The rate of disease spread and the rate at which infected individuals recover and become immune are constant over time.
+\end{enumerate}
+
+The dynamic of the population within the three compartment can be expressed as a system of nonlinear ordinary differential equations:
+\begin{align}
+    \frac{dS}{dt} &= \mu N - \frac{\beta SI}{N} - \mu S \\
+    \frac{dI}{dt} &= \frac{\beta SI}{N} - \gamma I - \mu I \\
+    \frac{dR}{dt} &= \gamma I - \mu R
+\end{align}
+where $S$, $I$, and $R$ represent the susceptible, recovered, and infectious population.
+$N$ is the size of the total population, and in this case $S + I + R = N$.
+The $\mu$, $\beta$ and $\gamma$ are the rates of natural birth (death), transmission and recovery, respectively.
+$\beta$ is the rate of new infections per unit time, so $\beta SI$ is the number of new infections that occur per unit time (the incidence rate).
+Here, the mean recovery time for an infected individual is $\frac{1}{\gamma}$.
+By assuming that the entire population was infected at the onset and considering a zero natural birth (death) rate, we can focus solely on the dynamics of the recovered population, as indicated by formula (1).
+
+The idea of dividing the infectious stage into $n$ sub-stages exploits the fact that the sum of a sequence of independent, exponentially distributed random variables follows an Erlang distribution (i.e.
+gamma distributed with integer shape parameter) \cite{therrien2018probability}
+\begin{equation}
+    f(x; n,n\gamma) = \frac{(n\gamma)^n}{(n-1)!} x^{n-1} e^{-n\gamma x} \quad x>0, n \in \mathbb{N}
+\end{equation}
+Utilizing the principle of the Linear Chain Trick (LCT), we divide the single infectious stage into $n$ substages, each with a constant rate of $n\gamma$.
+Therefore, formula (1) will be replaced by the following set of equations, which we refer to as the cohort model:
+\begin{align}
+    \frac{dI_1}{dt} &= - n \gamma I_1 \\
+    \frac{dI_2}{dt} &= n\gamma I_1 - n \gamma I_2 \\
+    &\vdots \\
+    \frac{dI_n}{dt} &= n\gamma I_{n-1} - n \gamma I_n \\
+    \frac{dR}{dt} &= n \gamma I_n
+\end{align}
+Note that this decomposition process will not change the mean stage duration, as the sum of the means of the $n$ exponentially distributed random variables still equals $\frac{1}{\gamma}$
+\begin{align*}
+    \sum_{i=1}^{n} \frac{1}{n\gamma} = \frac{1}{\gamma}
+\end{align*}
+
+The Geometric Chain Trick (GCT), on the other hand, features a fixed number of substages along with a set of inter-substage transition rates characterized by a geometric series, rather than a constant rate, as illustrated below:
+\begin{align}
+    \frac{dI_1}{dt} &= - a I_1 \\
+    \frac{dI_2}{dt} &= a I_1 - ar I_2 \\
+    \frac{dI_3}{dt} &= ar I_2 - ar^2 I_3 \\
+    &\vdots \\
+    \frac{dI_n}{dt} &= ar^{n-2} I_{n-1} - ar^{n-1} I_n \\
+    \frac{dR}{dt} &= ar^{n-1} I_n
+\end{align}
+
+Having established the model, our initial step involved analyzing the parameters within the GCT model.
+Our objective is to ascertain whether employing a fixed model structure (i.e., a fixed number of substages) and manipulating the parameters $a$ and $r$, which define the geometric transition rate, can replicate the effects of varying the number of substages.
+Additionally, we aim to evaluate whether the GCT structure offers greater flexibility in the choice of $\kappa$ (CV).
+By calculating the mean, variance, and $\kappa$, we are able to represent all these properties of the infectious stage duration distribution in terms of a and r.
+Similar to the LCT, utilizing the GCT facilitates independent alterations in the mean and $\kappa$.
+Given this, we can invert the process: starting with specific values of mean and $\kappa$, we can identify the corresponding a and r that imbue the model with such characteristics.
+However, due to the complexity of the formula, solving it explicitly poses a challenge.
+To overcome this, we have developed a system of functions in the R programming language, employing uniroot() to determine the values of a and r.
+These parameters are then reapplied to the GCT model to compare simulation outcomes with those of the LCT model, examining alignment in both the infectious stage duration distribution and the incidence over time.
 

### 2024-03-27 thesis/draft.tex
temp storage

@@
 Having established the model, our initial step involved analyzing the parameters within the GCT model.
-Our objective is to ascertain whether employing a fixed model structure (i.e., a fixed number of substages) and manipulating the parameters $a$ and $r$, which define the geometric transition rate, can replicate the effects of varying the number of substages.
-Additionally, we aim to evaluate whether the GCT structure offers greater flexibility in the choice of $\kappa$ (CV).
-By calculating the mean, variance, and $\kappa$, we are able to represent all these properties of the infectious stage duration distribution in terms of a and r.
+We aim to determine whether employing a fixed model structure (i.e., a fixed number of substages) and manipulating the parameters $a$ and $r$, which define the geometric transition rate, can replicate the effects of varying the number of substages.
+Additionally, we aim to evaluate whether the GCT structure offers greater flexibility in the choice of $\kappa$ $(CV^2)$.
+By calculating the mean $(D)$, variance $(\sigma)$, and shape parameter $(\kappa)$, we are able to represent all these properties of the infectious stage duration distribution in terms of $a$ and $r$.
 Similar to the LCT, utilizing the GCT facilitates independent alterations in the mean and $\kappa$.
-Given this, we can invert the process: starting with specific values of mean and $\kappa$, we can identify the corresponding a and r that imbue the model with such characteristics.
-However, due to the complexity of the formula, solving it explicitly poses a challenge.
-To overcome this, we have developed a system of functions in the R programming language, employing uniroot() to determine the values of a and r.
-These parameters are then reapplied to the GCT model to compare simulation outcomes with those of the LCT model, examining alignment in both the infectious stage duration distribution and the incidence over time.
+Given this, we can invert the process: starting with specific values of $D$ and $\kappa$, we can identify the corresponding $a$ and $r$ that endow the model with such characteristics.
+However, due to the complexity of the formula, solving it explicitly is challenging.
+To overcome this, we have developed a system of functions in the \textbf{R} programming language, employing \verb|uniroot()| to determine the values of $a$ and $r$.
+These parameters are then integrated to the GCT model to compare simulation outcomes with those of the LCT model, examining alignment in both the infectious stage duration distribution and the incidence over time.
 

### 2024-03-27 thesis/draft.tex
temp storage

@@
 \end{align*}
-Our approach, using a geometric series of rates, addresses this issue by allowing $\kappa$ to take any value within the range $\left[\ \frac{1}{n}, 1 \right)$.
+Our approach, using a geometric series of rates, addresses this issue by allowing $\kappa$ to take any value within the range $\left[ \frac{1}{n}, 1 \right)$.
 A detailed explanation and the derivation process are provided in section $4.2$.
@@
 $\beta$ is the rate of new infections per unit time, so $\beta SI$ is the number of new infections that occur per unit time (the incidence rate).
+Together, formulas $(3,4,5)$ will be referred to as the transmission model.
 Here, the mean recovery time for an infected individual is $\frac{1}{\gamma}$.
@@
 
-\section*{Results}
+After confirming that the GCT model, which employs a fixed structure and varying parameters $a$ and $r$, achieves comparable efficacy to the LCT model, we proceeded to empirically test the model's enhanced performance.
+Specifically, we focused on its improved flexibility in the kappa parameter through data fitting.
+Initially, we utilized simulated data as a replacement for actual data.
+This incidence data was generated through transmission model simulations, with deviations (noise) introduced via a mean reporting ratio and a negative binomial dispersion parameter.
+We performed model fitting using the \textbf{bbmle} package in R, estimating the model parameters through Maximum Likelihood Estimation (MLE).
+Our goal was to accurately retrieve the parameter values used to generate the incidence data, assessing the fitting quality through the mean squared error (MSE).
+
+We began by applying the same model type (LCT or GCT) for both data generation and model fitting.
+Data were generated by simulating the LCT model with two substages $(n=2)$ and then fitted using the LCT model, evaluating both identical and modified structures.
+This methodology was similarly applied to the GCT model.
+To further test the GCT model, we increased the fitting challenge by generating incidence data using the LCT model and then fitting it with the GCT model, varying the number of substages.
+Subsequently, we modified both the fitting model's structure and the input data to observe the impact on fitting outcomes.
+Finally, we extracted the parameters estimated through fitting and compared them with the actual values (those used for simulation).
+
+In the next phase of our study, we will approach the fitting of interval data from two perspectives: the rounding paradigm, which interprets 2 days as meaning any duration between 1.5 and 2.5 days, and the start/end paradigm, which considers potential starting and ending times.
+Initially, we will generate interval data using various distributions, such as gamma and lognormal, and subsequently apply both the LCT and GCT models for fitting.
+Although the GCT model can replicate certain properties (such as $D$ and $\kappa$) observed with the gamma distribution, its inherent distribution does not match the gamma distribution precisely.
+Thus, it's necessary for us to develop a new likelihood function for fitting purposes.
+Furthermore, we intend to incorporate real-world data, including incidence and interval data, to evaluate the GCT model's performance comprehensively.
+By comparing the fitting results from both the LCT and GCT models, we aim to quantitatively determine the effectiveness of our proposed method in enhancing model accuracy and its reflection of actual epidemiological scenarios.
+
+\section{Results}
+\subsection{Independent and Flexible Modification of Mean and Kappa in Stage Duration Distribution Using the GCT Model}
+As discussed in previous sections, the duration distribution of the stage can be parameterized by its mean (D) and shape parameter $(\kappa^2 = CV^2)$.
+In the GCT model, the calculation of mean and variance is achieved by summing the respective properties of each substage, shown as follows:
+\begin{align}
+    Mean: \quad  &\sum_{i=1}^{n_f} \frac{1}{\gamma_i} = \frac{\frac{1}{a} (\frac{1}{r^n} - 1)}{\frac{1}{r} -1} \\
+    Variance: \quad &\sum_{i=1}^{n_f} \left(\frac{1}{\gamma_i} \right)^2 = \frac{\frac{1}{a^2} (\frac{1}{r^{2n}} - 1)}{\frac{1}{r^2} -1}
+\end{align}
+The shape parameter $\kappa$, determined by $\frac{\sigma^2}{D}$, can thus be obtained:
+\begin{align}
+    \kappa = \frac{(1+r^{n}) (1-r)}{(1+r) (1-r^{n})}
+\end{align}
+This simplification revealed that the value of $\kappa$ is solely determined by the parameter $r$, thus enabling the alignment of kappa without affecting the mean, which can be adjusted through alterations to parameter $a$.
+Similarly, the mean can be scaled to a desired value without impacting $\kappa$.
+This grants the model independence in modifying properties.
+
+Importantly, within the GCT model, even though the number of substages can be modified $(n)$, the value of $\kappa$ becomes independent of this number.
+It should also be noted that, according to formula (20), the range of $\kappa$ can encompass any real number within the range $\left[ \frac{1}{n}, 1 \right)$ for $r > 0$, thereby significantly enhancing the model's flexibility.
+In contrast, when applying a similar calculation for $\kappa$ in the LCT model, we find that $\kappa = \frac{1}{n}$ and the shape parameter is therefore constrained to a series of discrete values, as follows:
+\begin{align*}
+    \kappa \in \left\{ 1, \frac{1}{2}, \frac{1}{3}, \dots \right\}
+\end{align*}
+The transition from discrete to continuous values in kappa significantly enhances the model's adaptability to the shape of the stage duration distribution, enabling a more precise fit to the observed data and allowing for a nuanced interpretation of temporal patterns.
+
+\subsection{Matching GCT and LCT Model Performance through Parameter Adjustment within a Uniform Structure}
+
+Using formulas (18) and (20), we developed code to reverse-engineer the process: given specific values of D and kappa, we can identify the corresponding $a$ and $r$ that endow the model with such characteristics.
+These $a$ and $r$ values were subsequently integrated into the GCT model.
+Our initial goal was to determine if the stage duration distribution could match between the two models, assuming identical $D$ and $\kappa$ values.
+The number of substages in the LCT model was set to 2, 4, 6, and 8, while the substage number for the GCT model was fixed at 12.
+We selected a mean of 10 days.
+Applying the mean and kappa values $\left( \frac{1}{2}, \frac{1}{4}, \frac{1}{6}, \frac{1}{8} \right)$ to the system, we plotted the stage duration distributions generated from the LCT cohort model alongside those from the GCT with the corresponding a and r values.
+As depicted in Figure x.x, the stage duration distributions exhibit slightly different shapes, despite having identical mean and shape parameters.
+To further examine the match of properties, D and kappa were numerically calculated from the output distributions, revealing identical values across both models.
+This indicates that while the distributions' shapes may not completely match, they can possess the same properties.
+
+By reintegrating the infectious stage duration distribution into the whole dynamic system, we present the dynamics of newly infected cases (incidence) as simulated by the transmission model.
+Figure x.x illustrates the incidence time series for the LCT model with 4 and 8 substages, represented by yellow and blue lines, respectively.
+The corresponding kappa values of the stage duration distribution for them are $\frac{1}{4}$ and $\frac{1}{8}$.
+Employing the GCT model with a fixed 12 substages and utilizing $a$ and $r$ values derived from the reverse-engineering system for the targeted kappa values, we successfully approximate the incidence patterns observed in the LCT model across different substage configurations, as shown by the black dashed line.
+This demonstrates the GCT model's capability to emulate the LCT model's performance by adjusting parameters within a constant structural framework.
+Compared to modifying the model structure to achieve desired kappa values, the GCT model simplifies parameter estimation and model fitting processes.
+
+\subsection{Substage Configuration Minimally Impacts Model Fitting}
+\textbf{Only temporarily. I'm going to change all of this...}
+
+After establishing the theoretical foundation, we proceeded with model fitting to assess the GCT model's performance.
+The incidence data were obtained from the model simulation, and the noises were added using a mean reporting ratio of 0.9 and a negative binomial dispersion parameter of 1000.
+Initially, we employed the same model type for both data generation and fitting.
+In Figure x.x.A, the data, generated by the LCT model with 2 substages, were fitted using an LCT model of the identical structure, as the plot title indicates.
+The quality of the fitting was evaluated based on the mean squared error (MSE).
+With an MSE of 1.2, the fitting can be considered satisfactory.
+Figure x.x.B showcases the parameters estimated post-fitting.
+Conversely, employing the same incidence data but altering the fitting model's structure from an LCT with 2 substages to an LCT with 6 substages, as depicted in Figure x.x.C, resulted in less favorable outcomes, with the MSE escalating to 25.
+The estimated parameters, displayed in Figure x.x.D, showed significant discrepancies.
+
+On the other hand, employing the GCT model with 12 substages for both generating and fitting incidence data resulted in a satisfactory fit, as indicated by an MSE of 0.01 (Figure x.x.A), and parameters post-fitting were close to the exact values (Figure x.x.B).
+However, altering the fitting model's structure from 12 substages to 6 substages, the model still exhibited a close match with the original in terms of both shape (Figure x.x.C) and parameters (Figure x.x.D), albeit with a slightly increased MSE (Figure x.x.C).
+This indicates that the GCT model's fitting performance is minimally impacted by changes in the model structure.
+
+\subsection{Adaptability of the GCT Model to Data Input and Structural Changes in Fitting}
+Proceeding to a more challenging test, we utilized a different model type for data generation and model fitting.
+As demonstrated in Figure x.x, the data generated by the LCT model with 2 substages were fitted using the GCT model with 12 substages.
+Despite the significant difference in model structure, the fitting results were reasonably favorable, with an MSE of 2.4.
+Altering the fitting model from a GCT with 2 substages to a GCT with 6 substages had negligible impact on the fitting outcomes (Figure x.x).
+Similarly, changing the input data from an LCT with 2 substages to an LCT with 7 substages resulted in slightly improved fitting, though the difference was not substantial (Figure x.x).
+These findings suggest that regardless of changes to the fitting model structure or the data used, the GCT model is capable of adapting to these variations, providing satisfactory fitting results.
+
+\subsection{Parameter Estimation}
+In this section, we discuss the parameter estimation results obtained through model fitting.
+The incidence data were generated using the GCT model with 6 substages, setting the $D$ to 10 and $\kappa$ to $\frac{2}{9}$.
+The model was fitted using the GCT framework with 2 to 10 substages.
+As depicted in Figure x.x, the precise values for each parameter are marked by a red line, with post-fitting parameter values displayed alongside their 95\% confidence intervals.
+It can be observed that the parameter estimation is not perfect, with kappa and beta consistently overestimated, and D underestimated.
+However, these discrepancies in parameter estimation counterbalance each other, resulting in a fitting outcome characterized by a low mean squared error.
+The absence of confidence intervals for models with 2 and 4 substages is attributed to structural inadequacy.
+As previously mentioned, the minimum possible value for kappa in the model is $\frac{1}{n}$.
+The model encounters limitations in aligning with a kappa value of 2/9 when using 2 and 4 substages.
+It is important to acknowledge the imprecision of parameter estimation, which will be a focus of our subsequent efforts.
 

### 2024-03-27 thesis/draft.tex
temp storage

@@
 \section{Abstract}
+Simple dynamical models based on Ordinary Differential Equations (ODE) are widely utilized to forecast the spread of infectious diseases and to analyze other biological systems.
+A limitation of this framework is its implicit assumption that residence time in any given state, such as the infectious state, follows an exponentially distributed pattern when a constant rate is employed.
+This assumption is often not realistic.
+A common solution is the application of the Linear Chain Trick (LCT), wherein a single stage is subdivided into multiple substages, each follows an identical exponential distribution, thereby yielding a more realistic Erlang distribution for the duration.
+However, the LCT poses challenges by restricting parametric flexibility, particularly resulting in less flexible dwell time distributions, and it often necessitates comparisons between models with varying structures (i.e.
+different numbers of substages).
+Here, we propose a Geometric Chain Trick (GCT) employing a fixed number of substages with transition rates forming a geometric sequence.
+Through mathematical analysis, simulation, and model fitting, we demonstrate that our approach retains the principal benefits of the LCT while providing increased flexibility and facilitating easier data fitting.
+Unlike the recently introduced Generalized Linear Chain Trick (GLCT), which incorporates higher-dimensional auxiliary stages to model delays in transitions between main compartments, our method offers a more direct means of achieving flexibility.
+Although this study focuses on the application of the LCT and GCT in the SIR (susceptible-infectious-recovered) model, the potential applications of our approach span a broad array of dynamical models, including those for infectious diseases and other systems (e.g.
+population dynamics, biochemical reaction networks), offering advantages in computational efficiency, ease of parameter estimation, and the flexibility of time distributions.
 

### 2024-03-28 thesis/draft.tex
temp storage

@@
 \section{Discussion}
+In this study, we introduce the Geometric Chain Trick (GCT), characterized by a fixed number of substages and a geometric series of transition rates.
+This approach aims to address the limitations identified in the Linear Chain Trick (LCT), including rigid parameter values and a cumbersome structure determination process.
+Our results indicate that the GCT model can utilize a uniform structure to achieve efficacy comparable to that of the LCT model with varying substage structures.
+The comparison is based on the consistency of simulation results, examining both the cohort model (stage duration distribution) and the transmission model (incidence over time).
+This enhancement simplifies the optimization of parameter values, requiring only a one-time process during data fitting.
+Furthermore, by theoretically analyzing and empirically validating the model through data fitting, we demonstrate that our model offers enhanced flexibility, specifically a broader range for the possible values of $\kappa$, in modeling the distribution of infectious stage duration.
+This flexibility allows for the selection of mean $(M)$ and shape parameters $(\kappa)$ for the distribution, with $M \in \mathbb{R}^+$ and $\kappa \in [\frac{1}{n}, 1)$.
+
+Compared to current studies that aim to provide a more flexible dwell time distribution within the ODE framework, such as the Generalized Linear Chain Trick (GLCT) proposed by Hurtado and Kirosingh \cite{hurtado2019generalizations}, our method offers a simpler concept and structure for constructing, simulating, and analyzing models.
+This simplicity makes our approach more accessible to a wide range of scientists and fields of study.
+Additionally, our method allows for considering a more diverse substage transition pattern, for instance, a higher dimension transition matrix (e.g.
+Markov Chain).
+This idea is inspired by the GLCT \cite{hurtado2019generalizations}\cite{hurtado2021building}, yet it is less complex.
+In our Geometric Chain Trick (GCT), transitions between substages follow a sequential pattern, where each substage receives a unique input and progresses to another distinct substage, with no cycles among them.
+Modifying this pattern of transitions between substages could possibly yield a wider variety of stage duration distributions.
+
+There are potential challenges that need to be addressed in the current GCT model.
+As discussed in Section 4.x, the parameters estimated after data fitting do not precisely reflect the actual values, yet these values are often crucial for understanding disease dynamics.
+For example, the basic reproduction number, $R_0$, which indicates how many secondary infections can be expected from a single infected individual in a fully susceptible population, is essential for assessing the potential for disease outbreak and epidemic spread.
+The calculation of $R_0$ requires the use of the contact rate $(\beta)$, recovery rate $(\gamma)$, and birth/death rate $(\mu)$.
+Therefore, achieving accurate parameter values from model fitting is necessary, and more study is needed.
+
+In addition to seeking solutions to the challenges discussed and exploring the diverse transition patterns mentioned above, future efforts will also involve fitting the GCT model to real-world incidence and interval data to assess its performance.
+We also plan to integrate this framework into the SEIR (Susceptible, Exposed, Infectious, Recovered) model to better account for delayed responses.
+Many infectious diseases exhibit a notable incubation period during which the pathogen replicates within the host without making the host infectious.
+Incorporating an exposed phase into the SEIR model enables a more accurate representation of disease transmission dynamics across populations.
+
+In conclusion, this study suggests that our method simplifies the application of a uniform model architecture for data fitting and exhibits heightened adaptability for parameter value selection within an Ordinary Differential Equation (ODE) framework.
+Specifically, in the field of epidemiology, our research holds the potential to improve model performance and facilitate the creation of user-friendly tools for robust epidemic forecasting and simulation.
+The method demonstrates considerable flexibility in its structure, making it suitable to a wide array of applications across different biological systems.
 

### 2024-03-28 figures/distributionCompare.R
add distribution compare (try)

@@ -1,0 +1,35 @@
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal(base_size = 17))
+library(bbmle)
+library(pseudoErlang)
+library(gridExtra) #
+
+library(shellpipes)
+startGraphics(height=5, width=7)
+
+######################################################################
+
+compare <- function(D, n, nfix, ts, T) {
+  time <- timeSeq(ts, T)
+  sddLCT <- erlang(time, n, 1/D)
+  sigr <- sinnerFlow(β=0, D=10, kappa=1/n, n=nfix, μ=0, N=1, I0=1, ts=0.1, T=60)
+  df <- data.frame(Time = time,
+                   ssdLCT = sddLCT,
+                   ssdGCT = diff(sigr[,"R"])/ts)
+  ggplot(df, aes(x=Time)) +
+    geom_line(aes(y=ssdLCT, color="LCT"), linewidth=1) +
+    geom_line(aes(y=ssdGCT, color="GCT"), linewidth=1) +
+    labs(x="Stage duration, days", y="Probability Density",
+         title = paste0("n = ", n))
+}
+
+######################################################################
+
+ts <- 0.1
+T <- 60
+D <- 10
+nfix <- 12
+nn <- c(2,4,6,8)
+
+plots <- lapply(nn, function(nn) compare(D, nn, nfix, ts, T))
+grid.arrange(grobs = plots, ncol = 2, nrow = 2)

### 2024-03-28 figures/distributionCompare.R
temp storage

@@ -1,5 +1,5 @@
 library(deSolve)
-library(ggplot2); theme_set(theme_minimal(base_size = 17))
+library(ggplot2); theme_set(theme_minimal(base_size = 12))
 library(bbmle)
 library(pseudoErlang)
 library(gridExtra) #

### 2024-03-28 figures/mytest1.R

@@ -1,12 +1,12 @@
 library(ggplot2)
 library(shellpipes)
-startGraphics(height=1, width=1)
+startGraphics(height=7, width=7)
 
 
 df <- data.frame(A = seq(1,10), B = seq(1,10)^2)
 
 ggplot(df, aes(x = A, y = B)) +
 geom_line() + 
-labs(title = "THIS is a test plot (TESE)")
+labs(title = "THIS is a test plot")
 
 # not working qswl....

### 2024-03-28 figures/incidenceCompare.R
add incidence compare

@@ -1,0 +1,52 @@
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal(base_size = 12))
+library(bbmle)
+library(pseudoErlang)
+
+library(shellpipes)
+startGraphics(height=5, width=7)
+
+######################################################################
+
+β <- 0.2
+D <- 10
+fixn <- 12
+μ <- 0.001
+ts <- 0.1
+T <- 100
+
+N = 1
+I0 = 0.01
+
+n <- 4
+kappa <- 1/n
+time <- timeSeq(ts, T)
+sinner <- sinnerFlow(β, D, kappa, fixn, μ, N=N, I0=I0, ts, T)
+sinr <- SInRFlow(β, D, n, μ, N=N, I0=I0, ts, T)
+dfn4 <- data.frame(Time = time, 
+                   LCT = diff(sinr[,"Cinc"])/ts,
+                   GCT = diff(sinner[,"Cinc"])/ts)
+n <- 8
+kappa <- 1/n
+time <- timeSeq(ts, T)
+sinner <- sinnerFlow(β, D, kappa, fixn, μ, N=N, I0=I0, ts, T)
+sinr <- SInRFlow(β, D, n, μ, N=N, I0=I0, ts, T)
+dfn8 <- data.frame(Time = time, 
+                   LCT = diff(sinr[,"Cinc"])/ts,
+                   GCT = diff(sinner[,"Cinc"])/ts)
+
+
+df <- data.frame(Time = time, LCT4 = dfn4$LCT, LCT8 = dfn8$LCT,
+                 GCT4 = dfn4$GCT, GCT8 = dfn8$GCT)
+
+ggplot(df, aes(x=Time)) +
+  geom_line(aes(y=LCT4, color = "n = 4"), linewidth=1) +
+  geom_line(aes(y=LCT8, color = "n = 8"), linewidth=1) +
+  scale_color_manual(values = c("n = 4" = "blue", "n = 8" = "orange")) +
+  labs(color = "LCT", x = "Time, days",
+       y = "Newly Infected Cases (proportion)") +
+  geom_line(aes(y=GCT4, color = "black"), linewidth=1, linetype="dashed") +
+  geom_line(aes(y=GCT8, color = "black"), linewidth=1, linetype="dashed")
+
+
+

### 2024-03-28 figures/sametypeFitting.R
same type fitting plot

@@ -1,0 +1,92 @@
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal(base_size = 12))
+library(bbmle)
+library(pseudoErlang)
+
+library(shellpipes)
+startGraphics(height=5, width=9)
+
+loadEnvironments()
+
+######################################################################
+
+β <- 0.2
+D <- 10
+n <- 2
+μ <- 0.01  
+N <- 10000
+I0 <- 10
+ts <- 1
+T <- 200
+
+arp <- 0.9
+nbs <- 1000
+
+startPar <- list(logβ=-1, logD=2)
+
+######################################################################
+
+onetimeParE <- function(fit, fitPar) { # fitPar = c(N = 1, D = 10)
+  dfs <- list()
+  plots <- list()
+  
+  for (i in names(fitPar)) {
+    
+    log_est <- coef(fitW$fit)[[paste0("log", i)]]
+    log_stde <- coef(summary(fitW$fit))[paste0("log", i), 'Std. Error']
+    lower <- exp(log_est - 1.96*log_stde)
+    upper <- exp(log_est + 1.96*log_stde)
+    
+    df <- data.frame(variable = i,
+                     estimate = exp(log_est),
+                     lower = lower,
+                     upper = upper)
+    #print(df)
+    
+    #dfs[[paste("df", i, sep = "_")]] <- df
+    
+    p <- ggplot(df, aes(x=variable, y=estimate)) +
+      geom_point() +
+      geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
+      geom_hline(yintercept=fitPar[[i]], col='red', linewidth=1) +
+      labs(x = element_blank(), y = "Estimate Parameter")
+    #print(p)
+    
+    plots <- c(plots, list(p))
+  }
+  #grid.arrange(grobs = plots, ncol = length(fitPar))
+  return(plots)
+}
+
+###################################################################### Same n
+
+sinr <- SInRFlow(β, D, n, μ, N, I0, ts, T)
+df = simObs(sinr, arp, nbs, seed = 73)
+
+fixedPar <- list(n = n, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T)
+fitW <- simplFit(startPar, fixedPar, df, sir.nll)
[…]

### 2024-03-28 figures/sametypeFitting.R
why...

@@ -2,6 +2,7 @@ library(deSolve)
 library(ggplot2); theme_set(theme_minimal(base_size = 12))
 library(bbmle)
 library(pseudoErlang)
+library(patchwork)
 
 library(shellpipes)
 startGraphics(height=5, width=9)

### 2024-03-28 figures/sametypeFitting.R
try to solve

@@ -82,7 +82,7 @@ parEd <- onetimeParE(fitW, fitPar)
 row1 <- (incs | (parEs[[1]] | parEs[[2]]))
 row2 <- (incd | (parEd[[1]] | parEd[[2]]))
 
-row1 / row2
+print(row1 / row2)
 
 
 

### 2024-03-29 figures/sametypeFitting.R
temp storage

@@ -1,93 +1,0 @@
-library(deSolve)
-library(ggplot2); theme_set(theme_minimal(base_size = 12))
-library(bbmle)
-library(pseudoErlang)
-library(patchwork)
-
-library(shellpipes)
-startGraphics(height=5, width=9)
-
-loadEnvironments()
-
-######################################################################
-
-β <- 0.2
-D <- 10
-n <- 2
-μ <- 0.01  
-N <- 10000
-I0 <- 10
-ts <- 1
-T <- 200
-
-arp <- 0.9
-nbs <- 1000
-
-startPar <- list(logβ=-1, logD=2)
-
-######################################################################
-
-onetimeParE <- function(fit, fitPar) { # fitPar = c(N = 1, D = 10)
-  dfs <- list()
-  plots <- list()
-  
-  for (i in names(fitPar)) {
-    
-    log_est <- coef(fitW$fit)[[paste0("log", i)]]
-    log_stde <- coef(summary(fitW$fit))[paste0("log", i), 'Std. Error']
-    lower <- exp(log_est - 1.96*log_stde)
-    upper <- exp(log_est + 1.96*log_stde)
-    
-    df <- data.frame(variable = i,
-                     estimate = exp(log_est),
-                     lower = lower,
-                     upper = upper)
-    #print(df)
-    
-    #dfs[[paste("df", i, sep = "_")]] <- df
-    
-    p <- ggplot(df, aes(x=variable, y=estimate)) +
-      geom_point() +
-      geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
-      geom_hline(yintercept=fitPar[[i]], col='red', linewidth=1) +
-      labs(x = element_blank(), y = "Estimate Parameter")
-    #print(p)
-    
-    plots <- c(plots, list(p))
-  }
-  #grid.arrange(grobs = plots, ncol = length(fitPar))
-  return(plots)
-}
-
-###################################################################### Same n
-
-sinr <- SInRFlow(β, D, n, μ, N, I0, ts, T)
-df = simObs(sinr, arp, nbs, seed = 73)
-
-fixedPar <- list(n = n, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T)
-fitW <- simplFit(startPar, fixedPar, df, sir.nll)
[…]

### 2024-03-29 figures/difftypeFitting.R
add plots (hopefully all)

@@ -1,0 +1,61 @@
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal(base_size = 12))
+library(bbmle)
+library(pseudoErlang)
+library(patchwork)
+
+library(shellpipes)
+startGraphics(height=5, width=9)
+
+loadEnvironments()
+
+######################################################################
+β <- 0.2
+D <- 10
+n <- 2
+μ <- 0.01  
+N <- 10000
+I0 <- 10
+ts <- 1
+T <- 200
+
+sinr <- SInRFlow(β, D, n, μ, N, I0, ts, T)
+df = simObs(sinr, arp, nbs, seed = 73)
+
+######################################################################
+
+nfit <- 12
+startPar <- list(logβ=-1, logD=2, logkappa=-0.2)
+fixedPar <- list(n = nfit, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T) 
+fitW <- simplFit(startPar, fixedPar, df, sir.nll.g)
+org <- plotFit(fitW, df, fixedPar, type="SIgR", title = "LCT (n=2) -> GCT (n=12)")
+
+######################################################################
+
+nfit <- 6
+startPar <- list(logβ=-1, logD=2, logkappa=-0.2)
+fixedPar <- list(n = nfit, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T) 
+fitW <- simplFit(startPar, fixedPar, df, sir.nll.g)
+cfitshape <- plotFit(fitW, df, fixedPar, type="SIgR", title = "LCT (n=2) -> GCT (n=6)")
+
+######################################################################
+
+n <- 7
+sinr <- SInRFlow(β, D, n, μ, N, I0, ts, T)
+df = simObs(sinr, arp, nbs, seed = 73)
+
+nfit <- 12
+startPar <- list(logβ=-1, logD=2, logkappa=-0.2)
+fixedPar <- list(n = nfit, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T) 
+fitW <- simplFit(startPar, fixedPar, df, sir.nll.g)
+cdata <- plotFit(fitW, df, fixedPar, type="SIgR", title = "LCT (n=7) -> GCT (n=12)")
+
+######################################################################
+
+(org | (cfitshape / cdata)) + plot_layout(widths = c(2.5, 1))
+
+
+
+
+
+

### 2024-03-29 figures/parEst.R

@@ -1,0 +1,71 @@
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal(base_size = 12))
+library(bbmle)
+library(pseudoErlang)
+library(patchwork)
+
+library(shellpipes)
+startGraphics(height=3, width=9)
+
+loadEnvironments()
+
+######################################################################
+
+## save time version...
+
+β <- 0.2
+D <- 10
+μ <- 0.01  
+N <- 10000
+I0 <- 10
+ts <- 1
+T <- 200
+nbs<-1000
+arp<-0.9
+kappa=2/9
+
+######################################################################
+
+parEst <- function(nfix, cnfit, fitPar) {
+  sigr <- sinnerFlow(β, D, kappa, nfix, μ, N, I0, ts, T)
+  df <- simObs(sigr, arp, nbs, seed = 73)
+  startPar <- list(logβ=-1, logD=2, logkappa=-0.1)
+  
+  plots <- list()
+  for (i in names(fitPar)) {
+    data <- data.frame(variable = integer(),
+                       estimate = numeric(),
+                       lower = numeric(),
+                       upper = numeric())
+    for (n in cnfit) {
+      fixedPar <- list(n = n, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T) 
+      fitW <- simplFit(startPar, fixedPar, df, sir.nll.g)
+      
+      log_est <- coef(fitW$fit)[[paste0("log", i)]]
+      log_stde <- coef(summary(fitW$fit))[paste0("log", i), 'Std. Error']
+      lower <- exp(log_est - 1.96*log_stde)
+      upper <- exp(log_est + 1.96*log_stde)
+      data <- rbind(data, data.frame(variable=n, estimate=exp(log_est), lower=lower, upper=upper))
+    }
+    
+    p <- ggplot(data, aes(x=variable, y=estimate)) + 
+      geom_point() +
+      coord_flip()+
+      geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
+      geom_hline(yintercept=fitPar[[i]], col='red', linewidth=1) + 
+      labs(x="Fitting Substages", y = paste0("Estimate ", i))
+    plots <- c(plots, list(p))
+  }
+  return(plots)
+}
+
+nfix <- 5
+cnfit <- seq(5,10)
+fitPar = c(β=0.2, D=10, kappa=2/9)
+plots <- parEst(nfix, cnfit, fitPar)
+
[…]

### 2024-03-29 figures/sametypeFittingG.R

@@ -1,0 +1,88 @@
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal(base_size = 12))
+library(bbmle)
+library(pseudoErlang)
+library(patchwork)
+
+library(shellpipes)
+startGraphics(height=5, width=9)
+
+loadEnvironments()
+
+######################################################################
+
+onetimeParE <- function(fit, fitPar) { # fitPar = c(N = 1, D = 10)
+  dfs <- list()
+  plots <- list()
+  
+  for (i in names(fitPar)) {
+    
+    log_est <- coef(fitW$fit)[[paste0("log", i)]]
+    log_stde <- coef(summary(fitW$fit))[paste0("log", i), 'Std. Error']
+    lower <- exp(log_est - 1.96*log_stde)
+    upper <- exp(log_est + 1.96*log_stde)
+    
+    df <- data.frame(variable = i,
+                     estimate = exp(log_est),
+                     lower = lower,
+                     upper = upper)
+    #print(df)
+    
+    #dfs[[paste("df", i, sep = "_")]] <- df
+    
+    p <- ggplot(df, aes(x=variable, y=estimate)) +
+      geom_point() +
+      geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
+      geom_hline(yintercept=fitPar[[i]], col='red', linewidth=1) +
+      labs(x = element_blank(), y = "Estimate Parameter")
+    #print(p)
+    
+    plots <- c(plots, list(p))
+  }
+  #grid.arrange(grobs = plots, ncol = length(fitPar))
+  return(plots)
+}
+
+######################################################################
+
+β <- 0.2
+D <- 10
+nfix <- 12
+μ <- 0.01  
+N <- 10000
+I0 <- 10
+ts <- 1
+T <- 200
+kappa <- 2/9
+
+arp <- 0.9
+nbs <- 1000
+
+startPar <- list(logβ=-1, logD=2, logkappa=-1)
+
+###################################################################### Same n
+
+sigr <- sinnerFlow(β, D, kappa, nfix, μ, N, I0, ts, T)
+df <- simObs(sigr, arp, nbs, seed = 72)
+
+nfit = nfix
+fixedPar <- list(n = nfit, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T) 
[…]

### 2024-03-29 figures/sametypeFittingL.R

@@ -1,0 +1,93 @@
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal(base_size = 12))
+library(bbmle)
+library(pseudoErlang)
+library(patchwork)
+
+library(shellpipes)
+startGraphics(height=5, width=9)
+
+loadEnvironments()
+
+######################################################################
+
+β <- 0.2
+D <- 10
+n <- 2
+μ <- 0.01  
+N <- 10000
+I0 <- 10
+ts <- 1
+T <- 200
+
+arp <- 0.9
+nbs <- 1000
+
+startPar <- list(logβ=-1, logD=2)
+
+######################################################################
+
+onetimeParE <- function(fit, fitPar) { # fitPar = c(N = 1, D = 10)
+  dfs <- list()
+  plots <- list()
+  
+  for (i in names(fitPar)) {
+    
+    log_est <- coef(fitW$fit)[[paste0("log", i)]]
+    log_stde <- coef(summary(fitW$fit))[paste0("log", i), 'Std. Error']
+    lower <- exp(log_est - 1.96*log_stde)
+    upper <- exp(log_est + 1.96*log_stde)
+    
+    df <- data.frame(variable = i,
+                     estimate = exp(log_est),
+                     lower = lower,
+                     upper = upper)
+    #print(df)
+    
+    #dfs[[paste("df", i, sep = "_")]] <- df
+    
+    p <- ggplot(df, aes(x=variable, y=estimate)) +
+      geom_point() +
+      geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
+      geom_hline(yintercept=fitPar[[i]], col='red', linewidth=1) +
+      labs(x = element_blank(), y = "Estimate Parameter")
+    #print(p)
+    
+    plots <- c(plots, list(p))
+  }
+  #grid.arrange(grobs = plots, ncol = length(fitPar))
+  return(plots)
+}
+
+###################################################################### Same n
+
+sinr <- SInRFlow(β, D, n, μ, N, I0, ts, T)
+df = simObs(sinr, arp, nbs, seed = 73)
+
+fixedPar <- list(n = n, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T)
+fitW <- simplFit(startPar, fixedPar, df, sir.nll)
[…]

### 2024-03-29 thesis/draft.tex
temp storage
modify makefile for newplot depend on tempfunc
I think I broke Makefile...

@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width= \textwidth]{figures/distributionCompare.Rout.pdf}
+    \caption{IMPORTANT}
+\end{figure}
+
 By reintegrating the infectious stage duration distribution into the whole dynamic system, we present the dynamics of newly infected cases (incidence) as simulated by the transmission model.
@@
 Compared to modifying the model structure to achieve desired kappa values, the GCT model simplifies parameter estimation and model fitting processes.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width= \textwidth]{figures/incidenceCompare.Rout.pdf}
+    \caption{IMPORTANT}
+\end{figure}
 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width= \textwidth]{figures/sametypeFittingL.Rout.pdf}
+    \caption{IMPORTANT}
+\end{figure}
+
 On the other hand, employing the GCT model with 12 substages for both generating and fitting incidence data resulted in a satisfactory fit, as indicated by an MSE of 0.01 (Figure x.x.A), and parameters post-fitting were close to the exact values (Figure x.x.B).
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width= \textwidth]{figures/sametypeFittingR.Rout.pdf}
+    \caption{IMPORTANT}
+\end{figure}
+
 \subsection{Adaptability of the GCT Model to Data Input and Structural Changes in Fitting}
 Proceeding to a more challenging test, we utilized a different model type for data generation and model fitting.
-As demonstrated in Figure x.x, the data generated by the LCT model with 2 substages were fitted using the GCT model with 12 substages.
+As demonstrated in Figure x.x.A, the data generated by the LCT model with 2 substages were fitted using the GCT model with 12 substages.
 Despite the significant difference in model structure, the fitting results were reasonably favorable, with an MSE of 2.4.
-Altering the fitting model from a GCT with 2 substages to a GCT with 6 substages had negligible impact on the fitting outcomes (Figure x.x).
-Similarly, changing the input data from an LCT with 2 substages to an LCT with 7 substages resulted in slightly improved fitting, though the difference was not substantial (Figure x.x).
+Altering the fitting model from a GCT with 2 substages to a GCT with 6 substages had negligible impact on the fitting outcomes (Figure x.x.B).
+Similarly, changing the input data from an LCT with 2 substages to an LCT with 7 substages resulted in slightly improved fitting, though the difference was not substantial (Figure x.x.C).
 These findings suggest that regardless of changes to the fitting model structure or the data used, the GCT model is capable of adapting to these variations, providing satisfactory fitting results.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width= \textwidth]{figures/difftypeFitting.Rout.pdf}
+    \caption{IMPORTANT}
+\end{figure}
 
@@
 In this section, we discuss the parameter estimation results obtained through model fitting.
-The incidence data were generated using the GCT model with 6 substages, setting the $D$ to 10 and $\kappa$ to $\frac{2}{9}$.
-The model was fitted using the GCT framework with 2 to 10 substages.
+The incidence data were generated using the GCT model with 5 substages, setting the $D$ to 10 and $\kappa$ to $\frac{2}{9}$.
+The model was fitted using the GCT framework with 5 to 10 substages.
 As depicted in Figure x.x, the precise values for each parameter are marked by a red line, with post-fitting parameter values displayed alongside their 95\% confidence intervals.
@@
 However, these discrepancies in parameter estimation counterbalance each other, resulting in a fitting outcome characterized by a low mean squared error.
-The absence of confidence intervals for models with 2 and 4 substages is attributed to structural inadequacy.
-As previously mentioned, the minimum possible value for kappa in the model is $\frac{1}{n}$.
-The model encounters limitations in aligning with a kappa value of 2/9 when using 2 and 4 substages.
-It is important to acknowledge the imprecision of parameter estimation, which will be a focus of our subsequent efforts.
+It is possible for the parameter to lack confidence intervals, which could be attributed to structural inadequacies.
+As previously mentioned, the minimum possible value for kappa in the model with $n$ substages is $\frac{1}{n}$.
+The model encounters limitations when aligning with a kappa value of less than $\frac{1}{n}$.
+Overall, it is important to acknowledge the imprecision of parameter estimation, which will be a focus of our subsequent efforts.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width= \textwidth]{figures/parEst.Rout.pdf}
+    \caption{IMPORTANT}
+\end{figure}
 
@@
 
-This is a citation \cite{Feng2016}
-
-Another citation \cite{Anderson1991}
-
 Here's a simple plot (with no dependence)
@@
     \centering
-    \includegraphics{figures/mytest1.Rout.pdf}
+    \includegraphics[width= \textwidth]{figures/mytest1.Rout.pdf}
     \caption{Mytest1 plot (are you working??)}

### 2024-03-29 thesis/draft.tex
Move piping to thesis directory
first draft

@@
 Unlike the recently introduced Generalized Linear Chain Trick (GLCT), which incorporates higher-dimensional auxiliary stages to model delays in transitions between main compartments, our method offers a more direct means of achieving flexibility.
-Although this study focuses on the application of the LCT and GCT in the SIR (susceptible-infectious-recovered) model, the potential applications of our approach span a broad array of dynamical models, including those for infectious diseases and other systems (e.g.
+Although this study focuses on the application of the LCT and GCT in the epidemic SIR (susceptible-infectious-recovered) model, the potential applications of our approach span a broad array of dynamical models, including those for infectious diseases and other systems (e.g.
 population dynamics, biochemical reaction networks), offering advantages in computational efficiency, ease of parameter estimation, and the flexibility of time distributions.
@@
 In this study, we focused on one particular direction: generalizing the Erlang distribution of stage durations currently present in the model.
-The goals were to have a more effective way for parameter estimation and to allow for a more diverse range of shapes, achieved by eliminating the constraints imposed by the shape parameter ($\kappa$).
+The goals were to have a more effective way for parameter estimation and to eliminate the constraints imposed by the shape parameter ($\kappa$).
 To address these limitations, we proposed the Geometric Chain Trick (GCT).
@@
 This series is represented by the formula $\gamma_i = ar^{i-1}$, governed by parameters $a$ and $r$.
-This approach enables a uniform model structure with the parameter of interest transitioning from discrete values $(n$ and $ng)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability (details in Section 4.x).
+This approach enables a uniform model structure with the parameter of interest transitioning from discrete values $(n$ and $n\gamma)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability (details in Section 4.x).
 
@@
 Our approach, using a geometric series of rates, addresses this issue by allowing $\kappa$ to take any value within the range $\left[ \frac{1}{n}, 1 \right)$.
-A detailed explanation and the derivation process are provided in section $4.2$.
+A detailed explanation and the derivation process are provided in Section $4.1$.
 Another advantage of this method is that it enables a fixed model structure where only the parameters change, eliminating the need to determine optimal substage numbers through the construction of multiple models, a process currently required in the ODE model that employs the LCT.
 
-Throughout this research, we primarily focused on the application of the LCT in the epidemic SIR model, which we refer to as the SI$^n$R model (or LCT model), where $n$ represents the number of substages divided.
+Throughout this research, we primarily focused on the application of the LCT in the epidemic SIR model, which we refer to as the LCT model (or SI$^n$R model, where $n$ represents the number of substages divided).
 The objective of our study is to develop and evaluate the method of using a fixed model structure and a geometrically distributed substage rate as a replacement for the constant rate, aiming to enhance flexibility and efficiency.
@@
 In the following sections, we first analyze the model parameters by calculating the formulae for two essential properties: the mean and shape parameter.
-We then invert the process: given a desired $D$ and $\kappa$, we develop a computational system to determine the parameter values $(a and r)$ that enable the model to produce a stage duration distribution with these properties.
+We then invert the process: given a desired $D$ and $\kappa$, we develop a computational system to determine the parameter values $a$ and $r$ that enable the model to produce a stage duration distribution with these properties.
 The obtained parameter values are then used to compare the GCT model with the LCT model through simulation.
@@
 $N$ is the size of the total population, and in this case $S + I + R = N$.
-The $\mu$, $\beta$ and $\gamma$ are the rates of natural birth (death), transmission and recovery, respectively.
-$\beta$ is the rate of new infections per unit time, so $\beta SI$ is the number of new infections that occur per unit time (the incidence rate).
+$\mu$, $\beta$ and $\gamma$ are the rates of natural birth (death), transmission, and recovery, respectively.
+$\beta$ is the rate of new infections per unit time, so $\frac{\beta SI}{N}$ is the number of new infections that occur per unit time (the incidence rate).
 Together, formulas $(3,4,5)$ will be referred to as the transmission model.
@@
 
-Importantly, within the GCT model, even though the number of substages can be modified $(n)$, the value of $\kappa$ becomes independent of this number.
-It should also be noted that, according to formula (20), the range of $\kappa$ can encompass any real number within the range $\left[ \frac{1}{n}, 1 \right)$ for $r > 0$, thereby significantly enhancing the model's flexibility.
+Importantly, within the GCT model, the value of $\kappa$ becomes independent of this number.
+It should also be noted that, according to formula (20), the range of $\kappa$ can encompass any real number within the range $\left[ \frac{1}{n}, 1 \right)$ for $r > 0$, thereby enhancing the model's flexibility.
 In contrast, when applying a similar calculation for $\kappa$ in the LCT model, we find that $\kappa = \frac{1}{n}$ and the shape parameter is therefore constrained to a series of discrete values, as follows:
@@
 \end{align*}
-The transition from discrete to continuous values in kappa significantly enhances the model's adaptability to the shape of the stage duration distribution, enabling a more precise fit to the observed data and allowing for a nuanced interpretation of temporal patterns.
+The transition from discrete to continuous values in kappa increases the model's adaptability to the shape of the stage duration distribution, enabling a more precise fit to the observed data and allowing for a nuanced interpretation of temporal patterns.
 
@@
 Applying the mean and kappa values $\left( \frac{1}{2}, \frac{1}{4}, \frac{1}{6}, \frac{1}{8} \right)$ to the system, we plotted the stage duration distributions generated from the LCT cohort model alongside those from the GCT with the corresponding a and r values.
-As depicted in Figure x.x, the stage duration distributions exhibit slightly different shapes, despite having identical mean and shape parameters.
+As depicted in Figure \ref{distributionCompare}, the stage duration distributions exhibit slightly different shapes, despite having identical mean and shape parameters.
 To further examine the match of properties, D and kappa were numerically calculated from the output distributions, revealing identical values across both models.
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
-    \includegraphics[width= \textwidth]{figures/distributionCompare.Rout.pdf}
-    \caption{IMPORTANT}
+    \includegraphics[width=\textwidth]{figures/distributionCompare.Rout.pdf}
+    \caption{\textbf{Stage Duration distribution Comparsion between LCT and GCT.} \\ Duration of the infectious stage derived from the derivative of infectious population dynamics in simulations of the GCT (Geometric Chain Trick) and LCT (Linear Chain Trick) cohort models, with a mean infectious recovery period of 10 days $(\gamma = \frac{1}{10})$.
+The GCT model features a fixed substage number of 12, whereas the LCT model's substage number varies and is indicated in each figure's title.
+$\kappa$ values are calculated as $\frac{1}{n}$ for each plot.
+Parameters for the GCT model were adjusted to standardize the mean $(D)$ and shape parameter $(\kappa)$ across all distributions.}
+    \label{distributionCompare}
 \end{figure}
@@
 By reintegrating the infectious stage duration distribution into the whole dynamic system, we present the dynamics of newly infected cases (incidence) as simulated by the transmission model.
-Figure x.x illustrates the incidence time series for the LCT model with 4 and 8 substages, represented by yellow and blue lines, respectively.
+Figure \ref{incidenceCompare} illustrates the incidence time series for the LCT model with 4 and 8 substages, represented by yellow and blue lines, respectively.
 The corresponding kappa values of the stage duration distribution for them are $\frac{1}{4}$ and $\frac{1}{8}$.
 Employing the GCT model with a fixed 12 substages and utilizing $a$ and $r$ values derived from the reverse-engineering system for the targeted kappa values, we successfully approximate the incidence patterns observed in the LCT model across different substage configurations, as shown by the black dashed line.
-This demonstrates the GCT model's capability to emulate the LCT model's performance by adjusting parameters within a constant structural framework.
+This demonstrates the GCT model's capability to match the LCT model's performance by adjusting parameters within a constant structural framework.
 Compared to modifying the model structure to achieve desired kappa values, the GCT model simplifies parameter estimation and model fitting processes.
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
-    \includegraphics[width= \textwidth]{figures/incidenceCompare.Rout.pdf}
-    \caption{IMPORTANT}
+    \includegraphics[width=\textwidth]{figures/incidenceCompare.Rout.pdf}
+    \caption{\textbf{Incidence comparsion between GCT and LCT with different substage numbers} \\ Incidence over time for GCT and LCT transmission models, depicted under a mean recovery period of 10 days $(\gamma = \frac{1}{10})$, with $\beta=0.2$, $\mu=0.001$, $N=1$, $I_0=0.01$.
+The incidence from the LCT model is illustrated with a blue line for 4 substages and a yellow line for 8 substages.
+The black dashed line represents the GCT model incidence, featuring a uniform 12 substages, with adjusted $a$ and $r$ values to align the $\kappa$ value with that of the LCT model.
+This ensures both models have comparable $D$ and $\kappa$ values across the depicted incidences.}
+    \label{incidenceCompare}
 \end{figure}
@@
 Initially, we employed the same model type for both data generation and fitting.
-In Figure x.x.A, the data, generated by the LCT model with 2 substages, were fitted using an LCT model of the identical structure, as the plot title indicates.
+In Figure \ref{sametypeFittingL}.A, the data, generated by the LCT model with 2 substages, were fitted using an LCT model of the identical structure, as the plot title indicates.
 The quality of the fitting was evaluated based on the mean squared error (MSE).
-With an MSE of 1.2, the fitting can be considered satisfactory.
-Figure x.x.B showcases the parameters estimated post-fitting.
-Conversely, employing the same incidence data but altering the fitting model's structure from an LCT with 2 substages to an LCT with 6 substages, as depicted in Figure x.x.C, resulted in less favorable outcomes, with the MSE escalating to 25.
-The estimated parameters, displayed in Figure x.x.D, showed significant discrepancies.
+With an MSE of 1.435, the fitting can be considered satisfactory.
+Figure \ref{sametypeFittingL}.B showcases the parameters estimated post-fitting.
+Conversely, employing the same incidence data but altering the fitting model's structure from an LCT with 2 substages to an LCT with 6 substages, as depicted in Figure \ref{sametypeFittingL}.C, resulted in less favorable outcomes, with the MSE escalating to 25.521.
+The estimated parameters, displayed in Figure \ref{sametypeFittingL}.D, showed significant discrepancies.
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
     \includegraphics[width= \textwidth]{figures/sametypeFittingL.Rout.pdf}
-    \caption{IMPORTANT}
+    \caption{\textbf{Model fitting and parameter estimation in LCT model.} \\ Data generated form the LCT transmission model featured 2 substages, with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$.
+The parameters fitted are $\beta$ and $D$.
+Plots A and C show the fitting results (red), alongside the target values (green) and the input data to the model, after adding noise (blue), using 2 and 6 substages, respectively.
+Plots B and D present the estimated parameters post-fitting.
+Each parameter's 95\% confidence interval is shown, with the exact values highlighted by a red line.}
+    \label{sametypeFittingL}
 \end{figure}
 
-On the other hand, employing the GCT model with 12 substages for both generating and fitting incidence data resulted in a satisfactory fit, as indicated by an MSE of 0.01 (Figure x.x.A), and parameters post-fitting were close to the exact values (Figure x.x.B).
-However, altering the fitting model's structure from 12 substages to 6 substages, the model still exhibited a close match with the original in terms of both shape (Figure x.x.C) and parameters (Figure x.x.D), albeit with a slightly increased MSE (Figure x.x.C).
+On the other hand, employing the GCT model with 12 substages for both generating and fitting incidence data resulted in a close match, as indicated by an MSE of 0.378 (Figure \ref{sametypeFittingG}.A), and parameters post-fitting were close to the exact values (Figure \ref{sametypeFittingG}.B).
+In this case, altering the fitting model's structure from 12 substages to 6 substages, the model still exhibited a close match with the original in terms of both shape (Figure \ref{sametypeFittingG}.C) and parameters (Figure \ref{sametypeFittingG}.D), albeit with a slightly increased MSE (Figure \ref{sametypeFittingG}.C).
 This indicates that the GCT model's fitting performance is minimally impacted by changes in the model structure.
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
-    \includegraphics[width= \textwidth]{figures/sametypeFittingR.Rout.pdf}
-    \caption{IMPORTANT}
+    \includegraphics[width= \textwidth]{figures/sametypeFittingG.Rout.pdf}
+    \caption{\textbf{Model fitting and parameter estimation in GCT model.} \\ Data generated from the GCT model featured 12 substages, with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$ and $\kappa=\frac{2}{9}$.
+Remaining details are as described in Figure \ref{sametypeFittingL}.}
+    \label{sametypeFittingG}
 \end{figure}
@@
 Proceeding to a more challenging test, we utilized a different model type for data generation and model fitting.
-As demonstrated in Figure x.x.A, the data generated by the LCT model with 2 substages were fitted using the GCT model with 12 substages.
-Despite the significant difference in model structure, the fitting results were reasonably favorable, with an MSE of 2.4.
-Altering the fitting model from a GCT with 2 substages to a GCT with 6 substages had negligible impact on the fitting outcomes (Figure x.x.B).
-Similarly, changing the input data from an LCT with 2 substages to an LCT with 7 substages resulted in slightly improved fitting, though the difference was not substantial (Figure x.x.C).
+As demonstrated in Figure \ref{difftypeFitting}.A, the data generated by the LCT model with 2 substages were fitted using the GCT model with 12 substages.
+Despite the significant difference in model structure, the fitting results were reasonably favorable, with an MSE of 4.583.
+Altering the fitting model from a GCT with 2 substages to a GCT with 6 substages had negligible impact on the fitting outcomes (Figure \ref{difftypeFitting}.B).
+Similarly, changing the input data from an LCT with 2 substages to an LCT with 7 substages resulted in slightly improved fitting, though the difference was not substantial (Figure \ref{difftypeFitting}.C).
 These findings suggest that regardless of changes to the fitting model structure or the data used, the GCT model is capable of adapting to these variations, providing satisfactory fitting results.
@@
     \includegraphics[width= \textwidth]{figures/difftypeFitting.Rout.pdf}
-    \caption{IMPORTANT}
+    \caption{\textbf{Model fitting using different model types.} \\  Input data for Plots A and B were generated using a 2-substages LCT transmission model simulation with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$.
+The parameter fitted include $\beta$, $D$, and $\kappa$.
+Model fitting for Plot A utilized a GCT model with 12 substages, while Plot B also using GCT but with 6 substages.
+For Plot C, data derived from an LCT model with 7 substages were fitted using a GCT model with 12 substages.}
+    \label{difftypeFitting}
 \end{figure}
@@
 The model was fitted using the GCT framework with 5 to 10 substages.
-As depicted in Figure x.x, the precise values for each parameter are marked by a red line, with post-fitting parameter values displayed alongside their 95\% confidence intervals.
+As depicted in Figure \ref{parEst}, the precise values for each parameter are marked by a red line, with post-fitting parameter values displayed alongside their 95\% confidence intervals.
 It can be observed that the parameter estimation is not perfect, with kappa and beta consistently overestimated, and D underestimated.
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
     \includegraphics[width= \textwidth]{figures/parEst.Rout.pdf}
-    \caption{IMPORTANT}
+    \caption{\textbf{GCT parameter estimation for various structures.} \\ Input data for fittings were derived from the GCT transmission model featuring 5 subtages, with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$.
+Fittings were conducted for 5 to 10 substages within the GCT framework.
+The targeted parameter values are marked by red lines, with the 95\% confidence intervals depicted by black lines.}
+    \label{parEst}
 \end{figure}
@@
 This idea is inspired by the GLCT \cite{hurtado2019generalizations}\cite{hurtado2021building}, yet it is less complex.
-In our Geometric Chain Trick (GCT), transitions between substages follow a sequential pattern, where each substage receives a unique input and progresses to another distinct substage, with no cycles among them.
+In our Geometric Chain Trick (GCT) model, transitions between substages follow a sequential pattern, where each substage receives a unique input and progresses to another distinct substage, with no cycles among them.
 Modifying this pattern of transitions between substages could possibly yield a wider variety of stage duration distributions.
@@
 There are potential challenges that need to be addressed in the current GCT model.
-As discussed in Section 4.x, the parameters estimated after data fitting do not precisely reflect the actual values, yet these values are often crucial for understanding disease dynamics.
+As discussed in Section 4.5, the parameters estimated after data fitting do not precisely reflect the actual values, yet these values are often crucial for understanding disease dynamics.
 For example, the basic reproduction number, $R_0$, which indicates how many secondary infections can be expected from a single infected individual in a fully susceptible population, is essential for assessing the potential for disease outbreak and epidemic spread.
@@
 
-%%%%%%%%
-
-Here's a simple plot (with no dependence)
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width= \textwidth]{figures/mytest1.Rout.pdf}
-    \caption{Mytest1 plot (are you working??)}
-\end{figure}
-
-Another plot (depend on tempfunc.R)
-
-\begin{figure}[h]
-    \centering
-    \includegraphics{figures/mytest2.Rout.pdf}
-    \caption{Mytest2 plot}
-\end{figure}
-
-Maybe another plot from a different folder:
-
-\begin{figure}[h]
-    \centering
-    \includegraphics{figures/test.Rout.pdf}
-    \caption{test plot}
-\end{figure}
-
 \bibliographystyle{unsrt}

### 2024-03-29 figures/difftypeFitting.R

@@ -1,5 +1,5 @@
 library(deSolve)
-library(ggplot2); theme_set(theme_minimal(base_size = 12))
+library(ggplot2); theme_set(theme_minimal(base_size = 9))
 library(bbmle)
 library(pseudoErlang)
 library(patchwork)
@@ -18,6 +18,8 @@ N <- 10000
 I0 <- 10
 ts <- 1
 T <- 200
+arp <- 0.9
+nbs <- 1000
 
 sinr <- SInRFlow(β, D, n, μ, N, I0, ts, T)
 df = simObs(sinr, arp, nbs, seed = 73)

### 2024-03-29 figures/parEst.R

@@ -53,7 +53,7 @@ parEst <- function(nfix, cnfit, fitPar) {
       coord_flip()+
       geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
       geom_hline(yintercept=fitPar[[i]], col='red', linewidth=1) + 
-      labs(x="Fitting Substages", y = paste0("Estimate ", i))
+      labs(x="Fitting Substages", y = paste0("Estimate ", ifelse(i == "β", "beta", i)))
     plots <- c(plots, list(p))
   }
   return(plots)

### 2024-03-29 figures/sametypeFittingG.R

@@ -5,7 +5,7 @@ library(pseudoErlang)
 library(patchwork)
 
 library(shellpipes)
-startGraphics(height=5, width=9)
+startGraphics(height=7, width=9)
 
 loadEnvironments()
 
@@ -22,7 +22,7 @@ onetimeParE <- function(fit, fitPar) { # fitPar = c(N = 1, D = 10)
     lower <- exp(log_est - 1.96*log_stde)
     upper <- exp(log_est + 1.96*log_stde)
     
-    df <- data.frame(variable = i,
+    df <- data.frame(variable = ifelse(i == "β", "beta", i),
                      estimate = exp(log_est),
                      lower = lower,
                      upper = upper)
@@ -82,7 +82,7 @@ incd <- plotFit(fitW, df, fixedPar, type="SIgR", title = "GCT (n=12) -> GCT (n=6
 
 parEd <- onetimeParE(fitW, fitPar)
 
-row1 <- (incs | (parEs[[1]] | parEs[[2]]))
-row2 <- (incd | (parEd[[1]] | parEd[[2]]))
+row1 <- (incs | (parEs[[1]] | parEs[[2]] | parEs[[3]]))
+row2 <- (incd | (parEd[[1]] | parEd[[2]] | parEd[[3]]))
 
 print(row1 / row2)

### 2024-03-29 figures/sametypeFittingL.R

@@ -5,7 +5,7 @@ library(pseudoErlang)
 library(patchwork)
 
 library(shellpipes)
-startGraphics(height=5, width=9)
+startGraphics(height=7, width=9)
 
 loadEnvironments()
 
@@ -38,7 +38,7 @@ onetimeParE <- function(fit, fitPar) { # fitPar = c(N = 1, D = 10)
     lower <- exp(log_est - 1.96*log_stde)
     upper <- exp(log_est + 1.96*log_stde)
     
-    df <- data.frame(variable = i,
+    df <- data.frame(variable = ifelse(i == "β", "beta", i),
                      estimate = exp(log_est),
                      lower = lower,
                      upper = upper)

### 2024-04-19 Code/realfitting.R
temp storage

@@ -69,6 +69,121 @@ plot(data$Time, data$Cases, type="l", xlab="Time", ylab="New Cases")
 data <- as.data.frame(data)
 ggplot(data, aes(x=Time, y=Cases)) + 
   geom_line() +
-  geom_point(data=data.frame(x=data$Time[c(299,402)], y=data$Cases[c(299,402)]), aes(x,y), color="red",size=3)
+  geom_point(data=data.frame(x=data$Time[c(349, 454)], y=data$Cases[c(349, 454)]), aes(x,y), color="red",size=3)
+
+which.min(subset(data, Time>1956.5 & Time<1957.5)$Cases)
+subset(data, Time>1956.5 & Time<1957.5)[11,]
+onepeak <- data[349:454,] #299:402
+ggplot(onepeak, aes(x=Time, y=Cases)) + geom_line()
+
+
+nfit <- 12
+startPar <- list(logβ=-1, logD=2, logkappa=-0.2)
+fixedPar <- list(n = nfit, μ = 0.01, N = 160180000, I0 = 1, ts = 7, nbs = 1000, T = 742) # 171745
+
+sir.nll.g <- function(logβ, logD, logkappa, n, μ, N, I0, ts, T, nbs, obs) {
+  
+  out <- as.data.frame(sinnerFlow(β=exp(logβ), D=exp(logD), kappa=exp(logkappa), n=n, μ=μ, N=N, I0=I0, ts=ts, T=T))
+  nll <- -sum(dnbinom(x=obs, mu=diff(out$Cinc), size=nbs, log=TRUE))
+}
+
+fit0 <- mle2(sir.nll.g, 
+             data=list(obs=onepeak$Cases),
+             start=startPar, 
+             fixed=fixedPar,
+             method="Nelder-Mead",
+             control=list(trace = 0))
+
+logβ <- coef(fit0)[["logβ"]]
+logD <- coef(fit0)[["logD"]]
+logkappa <- coef(fit0)[["logkappa"]]
+mod.prep <- as.data.frame(sinnerFlow(β=exp(logβ), D=exp(logD), kappa=exp(logkappa),
+                                     n=fixedPar$n, μ=fixedPar$μ, N=fixedPar$N, 
+                                     I0=fixedPar$I0, ts=fixedPar$ts, T=fixedPar$T))
+
[…]

### 2024-04-19 Code/Pre_plot.R
temp storage

@@ -298,94 +298,88 @@ ts <- 1
 T <- 200
 nbs<-1000
 arp<-0.9
-kappa=2/9
+kappa<- 0.3
 
 ######################################################################
 
-parEst <- function(nfix, cnfit, fitPar) {
+parEst <- function(nfix, cnfit, fitPar, seeds) {
   sigr <- sinnerFlow(β, D, kappa, nfix, μ, N, I0, ts, T)
-  df <- simObs(sigr, arp, nbs, seed = 73)
-  startPar_g <- list(logβ=-1, logD=2, logkappa=-0.1)
-  startPar_l <- list(logβ=-1, logD=2)
-  fixedPar <- list(n = n, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T)
-  
-  for (par in names(fitPar)) {
-    data <- data.frame(variable = integer(),
-                       estimate_g = numeric(),
-                       lower_g = numeric(),
-                       upper_g = numeric(),
-                       estimate_l = numeric(),
-                       lower_l = numeric(),
-                       upper_l = numeric())
-    assign(paste("data", par, sep="_"), data)
-  }
-  
-  for (n in cnfit) {
-    fitW_g <- simplFit(startPar_g, fixedPar, df, sir.nll.g)
-    fitW_l <- simplFit(startPar_l, fixedPar, df, sir.nll)
-    
-    log_est_beta_l <- coef(fitW_l$fit)[["logβ"]]
-    log_est_beta_g <- coef(fitW_g$fit)[["logβ"]]
-    log_stde_beta_l <- coef(summary(fitW_l$fit))[paste0("log", i), 'Std. Error']
-    lower_beta_l <- exp(log_est - 1.96*log_stde)
-    upper <- exp(log_est + 1.96*log_stde)
-  }
-}
-
-parEst <- function(nfix, cnfit, fitPar) {
-  sigr <- sinnerFlow(β, D, kappa, nfix, μ, N, I0, ts, T)
-  df <- simObs(sigr, arp, nbs, seed = 73)
-  startPar_g <- list(logβ=-1, logD=2, logkappa=-0.1)
+  startPar_g <- list(logβ=-1, logD=2, logkappa=-1)
   startPar_l <- list(logβ=-1, logD=2)
   
   plots <- list()
   for (i in names(fitPar)) {
     data_g <- data.frame(variable = integer(),
-                       estimate_g = numeric(),
-                       lower_g = numeric(),
-                       upper_g = numeric())
[…]

### 2024-04-19 Code/realfitting.R

@@ -8,7 +8,7 @@ library(dplyr)
 ## data visualization
 ######################################################################
 
-path <- "https://raw.githubusercontent.com/Tiffany-Xie/Thesis_NX/main/data/WHO_covid.csv?token=GHSAT0AAAAAACN7IGL3P7DCW6BOB7ZBLLDAZRABGEQ"
+path <- "https://covid19.who.int/WHO-COVID-19-global-data.csv"
 data <- read.csv(path)
 
 data_c <- data %>% 
@@ -25,9 +25,10 @@ plot(seq(0,dim(data_c)[1]-1)[150:175], data_c$New_cases[150:175], type='l', xlab
 ## fitting
 ######################################################################
 
-nfit <- 12
+nfit <- 6
+maxpeak <- as.data.frame(data_c[150:175,]) 
 startPar <- list(logβ=-1, logD=2, logkappa=-0.2)
-fixedPar <- list(n = nfit, μ = 0.01, N = 1.4097e9, I0 = 271745, ts = 1, nbs = 1000, T = 26) # 171745
+fixedPar <- list(n = nfit, μ = 0.01, N = 1.4097e9, I0 = 171745, ts = 7, nbs = 1000, T = 7*26) # 171745
 
 sir.nll.g <- function(logβ, logD, logkappa, n, μ, N, I0, ts, T, nbs, obs) {
   
@@ -36,7 +37,7 @@ sir.nll.g <- function(logβ, logD, logkappa, n, μ, N, I0, ts, T, nbs, obs) {
 }
 
 fit0 <- mle2(sir.nll.g, 
-             data=list(obs=data_c$New_cases[150:175]),
+             data=list(obs=maxpeak$New_cases),
              start=startPar, 
              fixed=fixedPar,
              method="Nelder-Mead",
@@ -49,12 +50,12 @@ mod.prep <- as.data.frame(sinnerFlow(β=exp(logβ), D=exp(logD), kappa=exp(logka
                                      n=fixedPar$n, μ=fixedPar$μ, N=fixedPar$N, 
                                      I0=fixedPar$I0, ts=fixedPar$ts, T=fixedPar$T))
 
-df <- data.frame(Time = timeSeq(1,26), trueI = data_c$New_cases[150:175], fitI = diff(mod.prep$Cinc))
+df <- data.frame(Time = timeSeq(1,26)+150, trueI = maxpeak$New_cases, fitI = diff(mod.prep$Cinc))
 
 ggplot(df, aes(x=Time)) +
[…]

### 2024-04-19 Code/tempErrorPlot.R
+ temp error plot (with multiple input data)

@@ -1,0 +1,107 @@
+# temporary errorbar plot
+library(deSolve)
+library(ggplot2); theme_set(theme_minimal())
+library(bbmle)
+library(pseudoErlang)
+
+######################################################################
+## save time version...
+
+β <- 0.2
+D <- 10
+μ <- 0.01  
+N <- 10000
+I0 <- 10
+ts <- 1
+T <- 200
+nbs<-1000
+arp<-0.9
+kappa<- 0.3
+
+######################################################################
+
+parEst <- function(nfix, cnfit, fitPar, seeds) {
+  sigr <- sinnerFlow(β, D, kappa, nfix, μ, N, I0, ts, T)
+  startPar_g <- list(logβ=-1, logD=2, logkappa=-1)
+  startPar_l <- list(logβ=-1, logD=2)
+  
+  plots <- list()
+  for (i in names(fitPar)) {
+    data_g <- data.frame(variable = integer(),
+                         estimate = numeric(),
+                         lower = numeric(),
+                         upper = numeric())
+    data_l <- data.frame(variable = integer(),
+                         estimate = numeric(),
+                         lower = numeric(),
+                         upper = numeric())
+    for (n in cnfit) {
+      for (seed in seeds) {
+        df <- simObs(sigr, arp, nbs, seed = seed)
+        fixedPar <- list(n = n, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T)
+        fitW_g <- simplFit(startPar_g, fixedPar, df, sir.nll.g)
+        
+        log_est <- coef(fitW_g$fit)[[paste0("log", i)]]
+        log_stde <- coef(summary(fitW_g$fit))[paste0("log", i), 'Std. Error']
+        lower <- exp(log_est - 1.96*log_stde)
+        upper <- exp(log_est + 1.96*log_stde)
+        data_g <- rbind(data_g, data.frame(variable=n, estimate=exp(log_est), lower=lower, upper=upper))
+        
+        if (i != "kappa") {
+          fixedPar <- list(n = n, μ = μ, N = N, I0 = I0, ts = ts, nbs = nbs, T = T)
+          fitW_l <- simplFit(startPar_l, fixedPar, df, sir.nll)
+          
+          log_est <- coef(fitW_l$fit)[[paste0("log", i)]]
[…]

### 2024-04-19 Code/realfitting.R
temp storage

@@ -1,9 +1,10 @@
 library(deSolve)
-library(ggplot2); theme_set(theme_minimal())
+library(ggplot2); theme_set(theme_minimal(base_size = 18))
 library(bbmle)
 library(pseudoErlang)
 library(dplyr)
 
+#run temp functions before
 ######################################################################
 ## data visualization
 ######################################################################
@@ -65,17 +66,19 @@ measles_url <- "https://raw.githubusercontent.com/mac-theobio/Disease_data/maste
 data <- read.table(measles_url, header=FALSE, sep=" ")
 names(data) = c("Time", "Cases")
 
-which.min(data$Cases[330:450])
+#which.min(data$Cases[330:450])
+plot(data$Time, log(data$Cases), type="l", xlab="Time", ylab="New Cases", main="Log Scale")
 plot(data$Time, data$Cases, type="l", xlab="Time", ylab="New Cases")
 data <- as.data.frame(data)
 ggplot(data, aes(x=Time, y=Cases)) + 
   geom_line() +
   geom_point(data=data.frame(x=data$Time[c(349, 454)], y=data$Cases[c(349, 454)]), aes(x,y), color="red",size=3)
 
-which.min(subset(data, Time>1956.5 & Time<1957.5)$Cases)
-subset(data, Time>1956.5 & Time<1957.5)[11,]
+#which.min(subset(data, Time>1956.5 & Time<1957.5)$Cases)
+#subset(data, Time>1956.5 & Time<1957.5)[11,]
 onepeak <- data[349:454,] #299:402
-ggplot(onepeak, aes(x=Time, y=Cases)) + geom_line()
+ggplot(onepeak, aes(x=Time, y=Cases)) + geom_line() +
+  labs(title="One of the peaks")
 
 
 nfit <- 12
@@ -107,13 +110,14 @@ df <- data.frame(Time = timeSeq(7/365,7*106/365)+1954.696, trueI = onepeak$Cases
 ggplot(df, aes(x=Time)) +
   geom_line(aes(y=trueI, color = "trueI"), linewidth=1) +
   geom_line(aes(y=fitI, color = "fitI")) +
-  labs(x="Time, weeks", y = "Incidence")
+  labs(x="Time, weeks", y = "Incidence") +
+  labs(title="One of the peaks fitting")
 
[…]

### 2024-04-19 Code/tempErrorPlot.R
Leaving office

@@ -4,6 +4,10 @@ library(ggplot2); theme_set(theme_minimal())
 library(bbmle)
 library(pseudoErlang)
 
+library(shellpipes)
+
+loadEnvironments()
+
 ######################################################################
 ## save time version...
 

### 2024-04-21 thesis/draft.tex
modified:   draft.tex

@@
 \usepackage{subcaption}
+\usepackage{color}
 
@@
 \renewcommand{\headrulewidth}{0pt}
+
 \newcommand{\perday}{\ensuremath{\mathrm{/day}}}
+\newcommand{\Eqnref}[1]{Eq~\ref{#1}}
+\newcommand{\eqnref}[1]{Eq~\ref{#1}}
+\newcommand{\sref}[1]{Section~\ref{#1}}
+
+\newcommand{\comment}{\showcomment}
+%% \newcommand{\comment}{\nocomment}
+
+\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
+\newcommand{\nocomment}[3]{}
+
+\newcommand{\jd}[1]{\comment{cyan}{JD}{#1}}
 
@@
 Simple dynamical models based on Ordinary Differential Equations (ODE) are widely utilized to forecast the spread of infectious diseases and to analyze other biological systems.
-A limitation of this framework is its implicit assumption that residence time in any given state, such as the infectious state, follows an exponentially distributed pattern when a constant rate is employed.
+A limitation of this framework is its implicit assumption that residence time in any given state, such as the infectious state, follows an exponentially distributed pattern.
 This assumption is often not realistic.
-A common solution is the application of the Linear Chain Trick (LCT), wherein a single stage is subdivided into multiple substages, each follows an identical exponential distribution, thereby yielding a more realistic Erlang distribution for the duration.
-However, the LCT poses challenges by restricting parametric flexibility, particularly resulting in less flexible dwell time distributions, and it often necessitates comparisons between models with varying structures (i.e.
+A common solution is the application of the Linear Chain Trick (LCT), wherein a single stage is subdivided into multiple substages, each of which follows an identical exponential distribution, thereby yielding a more realistic Erlang distribution for the duration.
+However, the LCT poses challenges by restricting parametric flexibility, particularly resulting in less flexible dwell-time distributions, and it often necessitates comparisons between models with varying structures (i.e.
 different numbers of substages).
@@
 Through mathematical analysis, simulation, and model fitting, we demonstrate that our approach retains the principal benefits of the LCT while providing increased flexibility and facilitating easier data fitting.
-Unlike the recently introduced Generalized Linear Chain Trick (GLCT), which incorporates higher-dimensional auxiliary stages to model delays in transitions between main compartments, our method offers a more direct means of achieving flexibility.
+Compared to the recently introduced Generalized Linear Chain Trick (GLCT), which incorporates higher-dimensional auxiliary stages to model delays in transitions between main compartments, our method offers a more direct means of achieving flexibility.
 Although this study focuses on the application of the LCT and GCT in the epidemic SIR (susceptible-infectious-recovered) model, the potential applications of our approach span a broad array of dynamical models, including those for infectious diseases and other systems (e.g.
@@
 \section{Introduction}
-Ordinary differential equations (ODEs) are extensively utilized for mathematical modeling in the biological sciences, offering a robust framework to accurately describe the temporal dynamics of systems.
+Ordinary differential equations (ODEs) are extensively utilized for mathematical modeling in the biological sciences, offering a robust framework to describe the temporal dynamics of systems.
 Their capacity to model continuous state changes makes them essential in capturing complex biological phenomena such as population dynamics \cite{salisbury2011mathematical}\cite{sego2021generation}, epidemiological patterns of infectious diseases \cite{Anderson1991}\cite{Diekmann2000}\cite{Feng2016}, and cellular processes like proliferation in oncogenesis \cite{baker1998modelling}\cite{jarrett2018mathematical}.
@@
 A commonly used model is the SIR compartmental model \cite{Anderson1991}\cite{kermack1927contribution}, which divides the population into three states: Susceptible $(S)$, Infectious $(I)$, and Recovered $(R)$.
-In the model, a susceptible individual, upon contact with the infected population at a rate $\beta$ $(\frac{1}{time})$, transitions immediately to the infectious state.
-An infected individual then moves to the recoverd state at a recovery rate of $\gamma$ $(\frac{1}{time})$.
-However, one major limitation of this model it its implicit assumption that the duration of the infectious stage follows an exponential distribution.
-This can be observed in a simplified equation that models the dynamics of the infected population.
-It assumes the entire population was infected at the outset, and individuals can only transition out of the infectious stage at a rate of $\gamma$:
+In the model, a susceptible individual, upon contact with the infected population at a rate $\beta$ $[\frac{1}{\mathit{ time}}]$, transitions immediately to the infectious state.
+An infected individual then moves to the recovered state at a recovery rate of $\gamma$ $[\frac{1}{\mathit{time}}]$.
+A major limitation of this model is its implicit assumption that the duration of the infectious stage follows an exponential distribution.
+This can be observed in a simplified equation that models the dynamics of an infected cohort.
+The proportion of a cohort infected at the same time begins at 1, and follows:
 \begin{align}
@@
 \begin{align}
-    I(t) = I_0 e^{-\gamma t}
+    I(t) = e^{-\gamma t}.
+	 \label{explicitCohort}
 \end{align}
-where $I_0$ represent the initial infected population.
-Formula (2) implies a stochastic state transition model where the time spend in the infectious $I$ is exponentially distributed with a mean stage duration of $\frac{1}{\gamma}$.
-This dwell time distribution suggests that an individual has the highest probability of recovering from the disease immediately after infection, which is often not realistic.
+\Eqnref{explicitCohort} implies a stochastic state transition model where the time spend in the infectious $I$ is exponentially distributed with a mean stage duration of $\frac{1}{\gamma}$.
+\jd{Not really.
+Where did the stochastic part come from?
+Also, these things are implied by the assumptions, not the calculation.}
+\Eqnref{explicitCohort} implies that an individual has the highest probability of recovering from the disease immediately after infection, which is generally not realistic.
 In reality, a delay between the successful infection of an individual and the recovery from the disease is inevitable.
-The work conducted by Byrne AW et al.
-\cite{byrne2020inferred} presents the simulation of parameter distribution inferred for the duration of the infectious period, and the shape of this distribution closely resembles a gamma distribution.
-The study by Greenhalgh and Day (2017) \cite{greenhalgh2017time}, utilizing measles outbreak data, also demonstrates that the recovery rate from an infectious disease is not static but can vary based on a wide range of factors.
+The work conducted by Byrne AW et al.~\cite{byrne2020inferred} presents the simulation of parameter distribution inferred for the duration of the infectious period, and the shape of this distribution closely resembles a gamma distribution.
+The study by Greenhalgh and Day (2017) \cite{greenhalgh2017time}, utilizing measles outbreak data, also demonstrates that the recovery rate from an infectious disease is not constant but can vary based on a wide range of factors.
+\jd{We're not really interested in factors, though, only in time-dependence.
+Is this citation good for that?}
 
-It is important to note that even when the mean is constant, the shape of the stage duration distributions (i.e., the timing of stage transitions) can significantly impact the dynamics and the outcomes of predictive models in applied settings \cite{krylova2013effects}\cite{keeling2002understanding}\cite{wearing2005appropriate}\cite{nguyen2008noise}.
-Consequently, it becomes crucial to consider the type of distribution characterizing an individual in a specific state when constructing these models.
-A standard approach for developing models that incorporate different stage duration distributions is through the use of integro-differential equations (IDEs) or integral equations (IEs) \cite{hurtado2019generalizations}\cite{kermack1927contribution}\cite{hethcote1980integral}.
+Even when the mean is constant, the shape of the stage duration distributions (i.e., the timing of stage transitions) can significantly impact the dynamics and the outcomes of predictive models in applied settings \cite{krylova2013effects}\cite{keeling2002understanding}\cite{wearing2005appropriate}\cite{nguyen2008noise}.
+Consequently, it becomes crucial to consider the type of distribution characterizing an individual in a specific state when constructing these models.\jd{Try to be more consistent (and simpler) about what you call these distributions.
+Dwell-time, characterizing individual in state, stage-duration, are all saying the same thing.} A standard approach for developing models that incorporate different stage duration distributions is through the use of integro-differential equations (IDEs) or integral equations (IEs) \cite{hurtado2019generalizations}\cite{kermack1927contribution}\cite{hethcote1980integral}.
 IDEs facilitate the extension of the SIR model to include arbitrary distributions for the duration of infectiousness, as discussed in various sources \cite{feng2000endemic}\cite{hethcote1980integral}\cite{ma2006generality}.
-This approach allows for the capture of the 'memory effect', where the current rate of change in each compartment of the model (Susceptible, Infectious, Recovered) is influenced by the entire history of the epidemic, rather than relying on a constant rate assumption.
+This approach allows for the capture of the `memory effect', where the current rate of change in each compartment of the model (Susceptible, Infectious, Recovered) is influenced by the entire history of the epidemic, rather than relying on a constant rate assumption.
 
@@
 When applying LCT into the SIR model, the number of substages $n$ determines the shape of the stage duration distribution.
-As $n$ increases, the variance decreases, resulting in a distribution that is more peaked and narrow around the mean value $\frac{1}{\gamma}$.
-As $n$ approaches infinity, the distribution converges to a delta function (a deterministic value, implying that all individuals who become infectious at time $t$ recover at exactly time $t + \frac{1}{\gamma}$), and the system evolves into a delay differential equation \cite{krylova2013effects}\cite{hethcote1980integral}.
+As $n$ increases, the variance decreases, resulting in a distribution that is more peaked and narrow.
+As $n$ approaches infinity, the distribution converges to a delta function (a deterministic value, implying that all individuals who become infectious at time $t$ recover at exactly time $t + \frac{1}{\gamma}$), and the system approaches a delay differential equation \cite{krylova2013effects}\cite{hethcote1980integral}.
 
@@
 Studies have demonstrated that switching the stage duration distribution from an exponential to an Erlang distribution leads to predictions that align more closely with real data points \cite{kakizoe2015method}.
-However, despite the increased realism they bring to modeling disease processes, employing LCT in ODE models presents several limitations.
-One primary challenge is the inflexibility of the model's parameters.
-To integrate independent exponentially distributed random variables into the ODE system, $n$ must be an integer.
-As mentioned above, since the shape of the stage duration distribution is determined by $n$, it would also be constrained to discrete changes.
-Another inherent limitation of using LCT is the cumbersome and time-consuming process of determining the appropriate number of substage to divide.
-Multiple models with varying substages must be constructed and subjected to repeated data-fitting processes to identify the optimized parameter value.
+But employing LCT in ODE models presents important limitations.
+First is inflexibility: the number of substages, $n$, must be an integer.
+Another limitation is the cumbersome and time-consuming process of determining the appropriate number of substages to use.
+Multiple models with varying substages must be constructed and fitted to identify the optimized parameter value.
 Prior research has investigated approaches for estimating the number of states in the linear chain, employing techniques such as profile likelihood \cite{raue2009structural} and model reduction \cite{maiwald2016driving}.
-Yet, a comparative study of these methods indicates that the typical quality of experimental data is insufficient for reliably estimating the chain length using these approaches \cite{hauber2020estimating}.
+A comparative study of these methods indicates that the typical quality of experimental data is insufficient for reliably estimating the chain length using these approaches \cite{hauber2020estimating}.
 These limitations significantly restrict the adaptability and practical application performance of the model.
@@
 In this study, we focused on one particular direction: generalizing the Erlang distribution of stage durations currently present in the model.
-The goals were to have a more effective way for parameter estimation and to eliminate the constraints imposed by the shape parameter ($\kappa$).
+The goals were to have a more effective way for parameter estimation and to eliminate the constraints imposed on the shape parameter ($\kappa$).
 To address these limitations, we proposed the Geometric Chain Trick (GCT).
@@
 This series is represented by the formula $\gamma_i = ar^{i-1}$, governed by parameters $a$ and $r$.
-This approach enables a uniform model structure with the parameter of interest transitioning from discrete values $(n$ and $n\gamma)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability (details in Section 4.x).
+This approach enables a uniform model structure with the parameter of interest transitioning from discrete values $(n$ and $n\gamma)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability (details in \sref{Methods}).
 
@@
 
-\section{Method}
+\section{Methods}
+\label{Methods}
 Several assumptions have been made for the unforced SIR model (Susceptible, Infectious, and Recovered) that our study is based on: 
@@
 \subsection{Independent and Flexible Modification of Mean and Kappa in Stage Duration Distribution Using the GCT Model}
+\jd{What is new in this subsection? Can it be merged into Methods?}
 As discussed in previous sections, the duration distribution of the stage can be parameterized by its mean (D) and shape parameter $(\kappa^2 = CV^2)$.
@@
 
+\jd{Try to get the habit of doing referencing in latex instead of typing section numbers (see the example I tried with sref)}
 There are potential challenges that need to be addressed in the current GCT model.

### 2024-04-22 thesis/draft.tex
Abstract star

@@
 \newpage
-\section{Abstract}
+\section*{Abstract}
 Simple dynamical models based on Ordinary Differential Equations (ODE) are widely utilized to forecast the spread of infectious diseases and to analyze other biological systems.

### 2024-04-22 Code/tempErrorPlot.R
meet with xnr

@@ -99,7 +99,7 @@ parEst <- function(nfix, cnfit, fitPar, seeds) {
 nfix <- 4
 cnfit <- seq(4,12, by=2)
 fitPar = c(β=0.2, D=10, kappa=0.3)
-seeds <- c(71, 77, 121, 123)
+seeds <- 73:76
 
 parEst(nfix, cnfit, fitPar, seeds)
 

### 2024-04-23 thesis/draft.tex
some changes for thesis

@@
 \end{align}
-\Eqnref{explicitCohort} implies a stochastic state transition model where the time spend in the infectious $I$ is exponentially distributed with a mean stage duration of $\frac{1}{\gamma}$.
-\jd{Not really.
-Where did the stochastic part come from?
-Also, these things are implied by the assumptions, not the calculation.}
+In other words, the time spent in the infectious stage $I$ is exponentially distributed with a mean stage duration of $\frac{1}{\gamma}$.
 \Eqnref{explicitCohort} implies that an individual has the highest probability of recovering from the disease immediately after infection, which is generally not realistic.
@@
 The work conducted by Byrne AW et al.~\cite{byrne2020inferred} presents the simulation of parameter distribution inferred for the duration of the infectious period, and the shape of this distribution closely resembles a gamma distribution.
-The study by Greenhalgh and Day (2017) \cite{greenhalgh2017time}, utilizing measles outbreak data, also demonstrates that the recovery rate from an infectious disease is not constant but can vary based on a wide range of factors.
-\jd{We're not really interested in factors, though, only in time-dependence.
-Is this citation good for that?}
 
 Even when the mean is constant, the shape of the stage duration distributions (i.e., the timing of stage transitions) can significantly impact the dynamics and the outcomes of predictive models in applied settings \cite{krylova2013effects}\cite{keeling2002understanding}\cite{wearing2005appropriate}\cite{nguyen2008noise}.
-Consequently, it becomes crucial to consider the type of distribution characterizing an individual in a specific state when constructing these models.\jd{Try to be more consistent (and simpler) about what you call these distributions.
-Dwell-time, characterizing individual in state, stage-duration, are all saying the same thing.} A standard approach for developing models that incorporate different stage duration distributions is through the use of integro-differential equations (IDEs) or integral equations (IEs) \cite{hurtado2019generalizations}\cite{kermack1927contribution}\cite{hethcote1980integral}.
+Consequently, it becomes crucial to consider the type of distribution characterizing an individual in a specific state when constructing these models.
+A standard approach for developing models that incorporate different stage duration distributions is through the use of integro-differential equations (IDEs) or integral equations (IEs) \cite{hurtado2019generalizations}\cite{kermack1927contribution}\cite{hethcote1980integral}.
 IDEs facilitate the extension of the SIR model to include arbitrary distributions for the duration of infectiousness, as discussed in various sources \cite{feng2000endemic}\cite{hethcote1980integral}\cite{ma2006generality}.
@@
 This series is represented by the formula $\gamma_i = ar^{i-1}$, governed by parameters $a$ and $r$.
-This approach enables a uniform model structure with the parameter of interest transitioning from discrete values $(n$ and $n\gamma)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability (details in \sref{Methods}).
+This approach enables a uniform model structure with the parameter of interest transitioning from discrete values $(n$ and $n\gamma)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability (details in \sref{Models}).
 
@@
 Our approach, using a geometric series of rates, addresses this issue by allowing $\kappa$ to take any value within the range $\left[ \frac{1}{n}, 1 \right)$.
-A detailed explanation and the derivation process are provided in Section $4.1$.
+A detailed explanation and the derivation process are provided in \sref{parm result}.
 Another advantage of this method is that it enables a fixed model structure where only the parameters change, eliminating the need to determine optimal substage numbers through the construction of multiple models, a process currently required in the ODE model that employs the LCT.
@@
 
-\section{Methods}
-\label{Methods}
+\section{Models}
+\label{Models}
 Several assumptions have been made for the unforced SIR model (Susceptible, Infectious, and Recovered) that our study is based on: 
@@
 \subsection{Independent and Flexible Modification of Mean and Kappa in Stage Duration Distribution Using the GCT Model}
-\jd{What is new in this subsection? Can it be merged into Methods?}
+\label{parm result}
 As discussed in previous sections, the duration distribution of the stage can be parameterized by its mean (D) and shape parameter $(\kappa^2 = CV^2)$.
@@
 We selected a mean of 10 days.
-Applying the mean and kappa values $\left( \frac{1}{2}, \frac{1}{4}, \frac{1}{6}, \frac{1}{8} \right)$ to the system, we plotted the stage duration distributions generated from the LCT cohort model alongside those from the GCT with the corresponding a and r values.
+Applying the mean and kappa values $\left[ \frac{1}{2}, \frac{1}{4}, \frac{1}{6}, \frac{1}{8} \right]$ to the system, we plotted the stage duration distributions generated from the LCT cohort model alongside those from the GCT with the corresponding a and r values.
 As depicted in Figure \ref{distributionCompare}, the stage duration distributions exhibit slightly different shapes, despite having identical mean and shape parameters.
@@
 In this section, we discuss the parameter estimation results obtained through model fitting.
-The incidence data were generated using the GCT model with 5 substages, setting the $D$ to 10 and $\kappa$ to $\frac{2}{9}$.
+The true incidence data were generated using the GCT model with 5 substages, setting the $D$ to 10 and $\kappa$ to $\frac{2}{9}$.
+Random noises were added to the true incidence data using negative binomial to give us 4 different input data.
 The model was fitted using the GCT framework with 5 to 10 substages.
@@
     \centering
-    \includegraphics[width= \textwidth]{figures/parEst.Rout.pdf}
+    \includegraphics[width= 1.1\textwidth]{figures/parEst.Rout.pdf}
     \caption{\textbf{GCT parameter estimation for various structures.} \\ Input data for fittings were derived from the GCT transmission model featuring 5 subtages, with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$.

### 2024-04-23 thesis/draft.tex
some changes to thesis (draft)

@@
 \\McMaster University\\Hamiltion, Ontario, Canada L8S 4K1}
-\date {March 22, 2024}
+\date {April 23, 2024}
 \maketitle
@@
 We performed model fitting using the \textbf{bbmle} package in R, estimating the model parameters through Maximum Likelihood Estimation (MLE).
-Our goal was to accurately retrieve the parameter values used to generate the incidence data, assessing the fitting quality through the mean squared error (MSE).
+Our goal was to accurately retrieve the parameter values used to generate the incidence data, assessing the fitting quality through the root mean squared log error (RMSLE).
 
@@
 \subsection{Substage Configuration Minimally Impacts Model Fitting}
-\textbf{Only temporarily. I'm going to change all of this...}
 
@@
 In Figure \ref{sametypeFittingL}.A, the data, generated by the LCT model with 2 substages, were fitted using an LCT model of the identical structure, as the plot title indicates.
-The quality of the fitting was evaluated based on the mean squared error (MSE).
-With an MSE of 1.435, the fitting can be considered satisfactory.
+The quality of the fitting was evaluated based on the root mean squared logarithmic error (RMSLE), which measured the percentage errors between the actual and predicted value.
+With an RMSLE value of 1.64\%, the fitting can be considered satisfactory.
 Figure \ref{sametypeFittingL}.B showcases the parameters estimated post-fitting.
-Conversely, employing the same incidence data but altering the fitting model's structure from an LCT with 2 substages to an LCT with 6 substages, as depicted in Figure \ref{sametypeFittingL}.C, resulted in less favorable outcomes, with the MSE escalating to 25.521.
+Conversely, employing the same incidence data but altering the fitting model's structure from an LCT with 2 substages to an LCT with 6 substages, as depicted in Figure \ref{sametypeFittingL}.C, resulted in less favorable outcomes, with the RMSLE escalating to 9.30\%.
 The estimated parameters, displayed in Figure \ref{sametypeFittingL}.D, showed significant discrepancies.
@@
 
-On the other hand, employing the GCT model with 12 substages for both generating and fitting incidence data resulted in a close match, as indicated by an MSE of 0.378 (Figure \ref{sametypeFittingG}.A), and parameters post-fitting were close to the exact values (Figure \ref{sametypeFittingG}.B).
-In this case, altering the fitting model's structure from 12 substages to 6 substages, the model still exhibited a close match with the original in terms of both shape (Figure \ref{sametypeFittingG}.C) and parameters (Figure \ref{sametypeFittingG}.D), albeit with a slightly increased MSE (Figure \ref{sametypeFittingG}.C).
+On the other hand, employing the GCT model with 12 substages for both generating and fitting incidence data resulted in a close match, as indicated by an RMSLE of 2.34\% (Figure \ref{sametypeFittingG}.A), and parameters post-fitting were close to the exact values (Figure \ref{sametypeFittingG}.B).
+In this case, altering the fitting model's structure from 12 substages to 6 substages, the model still exhibited a close match with the original in terms of both shape (Figure \ref{sametypeFittingG}.C) and parameters (Figure \ref{sametypeFittingG}.D).
 This indicates that the GCT model's fitting performance is minimally impacted by changes in the model structure.
@@
 As demonstrated in Figure \ref{difftypeFitting}.A, the data generated by the LCT model with 2 substages were fitted using the GCT model with 12 substages.
-Despite the significant difference in model structure, the fitting results were reasonably favorable, with an MSE of 4.583.
+Despite the significant difference in model structure, the fitting results were reasonably favorable, with an RMSLE of 2.44\%.
 Altering the fitting model from a GCT with 2 substages to a GCT with 6 substages had negligible impact on the fitting outcomes (Figure \ref{difftypeFitting}.B).
@@
 \subsection{Parameter Estimation}
+\label{parm est}
 In this section, we discuss the parameter estimation results obtained through model fitting.
 The true incidence data were generated using the GCT model with 5 substages, setting the $D$ to 10 and $\kappa$ to $\frac{2}{9}$.
-Random noises were added to the true incidence data using negative binomial to give us 4 different input data.
-The model was fitted using the GCT framework with 5 to 10 substages.
+Random noise was added to the true incidence data using a negative binomial distribution, resulting in four different input datasets (a number we specifically set for this analysis).
+The model was fitted using both the GCT and LCT frameworks with 5 to 10 substages using the 4 sets of data.
 As depicted in Figure \ref{parEst}, the precise values for each parameter are marked by a red line, with post-fitting parameter values displayed alongside their 95\% confidence intervals.
-It can be observed that the parameter estimation is not perfect, with kappa and beta consistently overestimated, and D underestimated.
-However, these discrepancies in parameter estimation counterbalance each other, resulting in a fitting outcome characterized by a low mean squared error.
-It is possible for the parameter to lack confidence intervals, which could be attributed to structural inadequacies.
+It can be observed that the estimated parameter values derived from the LCT model, using different input data, tend to cluster together but on average vary across different substage numbers.
+These estimates could deviate substantially from the accurate parameter if an inappropriate structure is chosen.
+On the other hand, although the parameter estimation using the GCT model is not perfect—with kappa and beta being overestimated and D underestimated—the estimated values are less fluctuating across substage numbers, closer to the target values, and consistently include the target values within the 95\% confidence interval.
+The discrepancies in parameter estimation counterbalance each other, resulting in a fitting outcome characterized by a low root mean squared logarithmic error (RMSLE).
+
+Notice that it is possible for the parameter to lack confidence intervals, which could be attributed to structural inadequacies.
 As previously mentioned, the minimum possible value for kappa in the model with $n$ substages is $\frac{1}{n}$.
@@
     \includegraphics[width= 1.1\textwidth]{figures/parEst.Rout.pdf}
-    \caption{\textbf{GCT parameter estimation for various structures.} \\ Input data for fittings were derived from the GCT transmission model featuring 5 subtages, with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$.
-Fittings were conducted for 5 to 10 substages within the GCT framework.
-The targeted parameter values are marked by red lines, with the 95\% confidence intervals depicted by black lines.}
+    \caption{\textbf{GCT parameter estimation for various structures.} \\ The four input data for fittings were derived from the GCT transmission model featuring 5 subtages, with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$.
+Fittings were conducted for 5 to 10 substages within the GCT and LCT frameworks.
+The targeted parameter values are marked by red lines, with the 95\% confidence intervals depicted by black and green lines for GCT and LCT, respectively.}
     \label{parEst}
@@
 
-\jd{Try to get the habit of doing referencing in latex instead of typing section numbers (see the example I tried with sref)}
 There are potential challenges that need to be addressed in the current GCT model.
-As discussed in Section 4.5, the parameters estimated after data fitting do not precisely reflect the actual values, yet these values are often crucial for understanding disease dynamics.
+As discussed in \sref{parm est}, the parameters estimated after data fitting do not precisely reflect the actual values, yet these values are often crucial for understanding disease dynamics.
 For example, the basic reproduction number, $R_0$, which indicates how many secondary infections can be expected from a single infected individual in a fully susceptible population, is essential for assessing the potential for disease outbreak and epidemic spread.

### 2024-04-30 thesis/draft.tex
temp storage

@@
 \usepackage{color}
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric}
+\usetikzlibrary{arrows.meta,arrows}
 
@@
 \end{equation}
-Utilizing the principle of the Linear Chain Trick (LCT), we divide the single infectious stage into $n$ substages, each with a constant rate of $n\gamma$.
+Utilizing the principle of the Linear Chain Trick (LCT), we divide the single infectious stage into $n$ substages, each with a constant rate of $n\gamma$ (Figure \ref{LCTdiagram}). 
+\begin{figure}[ht]
+    \centering
+    \begin{tikzpicture}[scale=1, transform shape, >={Latex[length=2mm, width=2mm]}]
+        \node (I) at (0,1) [draw, minimum width=1cm, minimum height=1cm] {$I$};
+
+        % Nodes I1 to In
+        \node at (-4,0) [below, draw, minimum width=1cm, minimum height=1cm] (I1) {$I_1$};
+        \node at (-2,0) [below, draw, minimum width=1cm, minimum height=1cm] (I2) {$I_2$};
+        \node at (0,0) [below, draw, minimum width=1cm, minimum height=1cm] (I3) {$I_3$};
+        \node at (4,0) [below, draw, minimum width=1cm, minimum height=1cm] (In) {$I_n$};
+
+        % Dots
+        \draw[line width=0.5mm, dotted, dash pattern=on 0.5mm off 0.5mm] (1.7,-0.5) -- (2.3,-0.5);
+
+        % Arrows
+        \draw[->] (I1.east) -- (I2.west) node[midway, above] {$n\gamma$};
+        \draw[->] (I2.east) -- (I3.west) node[midway, above] {$n\gamma$};
+        \draw[->] (I3.east) -- ([xshift=1cm]I3.east) node[midway, above] {$n\gamma$};
+        \draw[->] ([xshift=-1cm]In.west) -- (In.west) node[midway, above] {$n\gamma$};
+
+        %\draw[-latex] (In.east) -- ([xshift=2cm]In.west) node[midway, above] {$n\gamma$};
+
+        % Arrows from I to I1 and In
+        \draw[-] (I.south west) -- (I1.north west);
+        \draw[-] (I.south east) -- (In.north east);
+
+    \end{tikzpicture}
+    \caption{Example diagram of the Linear Chain Trick}
+    \label{LCTdiagram}
+\end{figure}
+
 Therefore, formula (1) will be replaced by the following set of equations, which we refer to as the cohort model:
@@
 
-The Geometric Chain Trick (GCT), on the other hand, features a fixed number of substages along with a set of inter-substage transition rates characterized by a geometric series, rather than a constant rate, as illustrated below:
+The Geometric Chain Trick (GCT), on the other hand, features a fixed number of substages along with a set of inter-substage transition rates characterized by a geometric series (Figure \ref{GCTdiagram}), rather than a constant rate, as illustrated below:
 \begin{align}
@@
 \end{align}
+
+\begin{figure}[ht]
+    \centering
+    \begin{tikzpicture}[scale=1, transform shape, >={Latex[length=2mm, width=2mm]}]
+
+        % Nodes I1 to In
+        \node at (-4,0) [below, draw, minimum width=1cm, minimum height=1cm] (I1) {$I_1$};
+        \node at (-2,0) [below, draw, minimum width=1cm, minimum height=1cm] (I2) {$I_2$};
+        \node at (0,0) [below, draw, minimum width=1cm, minimum height=1cm] (I3) {$I_3$};
+        \node at (4,0) [below, draw, minimum width=1cm, minimum height=1cm] (In) {$I_n$};
+
+        % Dots
+        \draw[line width=0.5mm, dotted, dash pattern=on 0.5mm off 0.5mm] (1.7,-0.5) -- (2.3,-0.5);
+
+        % Arrows
+        \draw[->] (I1.east) -- (I2.west) node[midway, above] {$\gamma_1$};
+        \draw[->] (I2.east) -- (I3.west) node[midway, above] {$\gamma_2$};
+        \draw[->] (I3.east) -- ([xshift=1cm]I3.east) node[midway, above] {$\gamma_3$};
+        \draw[->] ([xshift=-1cm]In.west) -- (In.west) node[midway, above] {$\gamma_{n-1}$};
+        \draw[->] (In.east) -- ([xshift=1cm]In.east) node[midway, above] {$\gamma_n$};
+
+        %\draw[-latex] (In.east) -- ([xshift=2cm]In.west) node[midway, above] {$n\gamma$};
+
+    \end{tikzpicture}
+    \caption{Example diagram of the Geometric Chain Trick}
+    \label{GCTdiagram}
+\end{figure}
 
@@
 The corresponding kappa values of the stage duration distribution for them are $\frac{1}{4}$ and $\frac{1}{8}$.
+All other parameters are held constant.
+All other parameters are identical across the two models.
 Employing the GCT model with a fixed 12 substages and utilizing $a$ and $r$ values derived from the reverse-engineering system for the targeted kappa values, we successfully approximate the incidence patterns observed in the LCT model across different substage configurations, as shown by the black dashed line.
@@
 With an RMSLE value of 1.64\%, the fitting can be considered satisfactory.
-Figure \ref{sametypeFittingL}.B showcases the parameters estimated post-fitting.
+Figure \ref{sametypeFittingL}.B showcases the parameters estimated post-fitting along with their 95\% confidence intervals, with the target value represented by the red line.
 Conversely, employing the same incidence data but altering the fitting model's structure from an LCT with 2 substages to an LCT with 6 substages, as depicted in Figure \ref{sametypeFittingL}.C, resulted in less favorable outcomes, with the RMSLE escalating to 9.30\%.
@@
     \centering
-    \includegraphics[width= \textwidth]{figures/sametypeFittingL.Rout.pdf}
+    \begin{tikzpicture}
+        \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width= \textwidth]{figures/sametypeFittingL.Rout.pdf}};
+        \begin{scope}[x={(image.south east)},y={(image.north west)}]
+            % Add labels at desired positions
+            \node at (0.01,0.95) {\textbf{(A)}};
+            \node at (0.55,0.95) {\textbf{(B)}};
+            \node at (0.01,0.45) {\textbf{(C)}};
+            \node at (0.55,0.45) {\textbf{(D)}};
+        \end{scope}
+    \end{tikzpicture}
     \caption{\textbf{Model fitting and parameter estimation in LCT model.} \\ Data generated form the LCT transmission model featured 2 substages, with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$.
@@
 
-On the other hand, employing the GCT model with 12 substages for both generating and fitting incidence data resulted in a close match, as indicated by an RMSLE of 2.34\% (Figure \ref{sametypeFittingG}.A), and parameters post-fitting were close to the exact values (Figure \ref{sametypeFittingG}.B).
-In this case, altering the fitting model's structure from 12 substages to 6 substages, the model still exhibited a close match with the original in terms of both shape (Figure \ref{sametypeFittingG}.C) and parameters (Figure \ref{sametypeFittingG}.D).
+Not surprisingly, employing the GCT model with 2 substages for both generating and fitting incidence data also resulted in a close match, as indicated by an RMSLE of 2.34\% (Figure \ref{sametypeFittingG}.A), and parameters post-fitting were close to the exact values (Figure \ref{sametypeFittingG}.B).
+Unlike the case with the LCT model, when the structure of the fitting model was altered from 2 substages to 6, the model still exhibited a close match with the original in terms of both shape (Figure \ref{sametypeFittingG}.C) and parameters (Figure \ref{sametypeFittingG}.D).
 This indicates that the GCT model's fitting performance is minimally impacted by changes in the model structure.
@@
     \centering
-    \includegraphics[width= \textwidth]{figures/sametypeFittingG.Rout.pdf}
+    \begin{tikzpicture}
+        \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width= \textwidth]{figures/sametypeFittingG.Rout.pdf}};
+        \begin{scope}[x={(image.south east)},y={(image.north west)}]
+            % Add labels at desired positions
+            \node at (0.01,0.95) {\textbf{(A)}};
+            \node at (0.55,0.95) {\textbf{(B)}};
+            \node at (0.01,0.45) {\textbf{(C)}};
+            \node at (0.55,0.45) {\textbf{(D)}};
+        \end{scope}
+    \end{tikzpicture}
     \caption{\textbf{Model fitting and parameter estimation in GCT model.} \\ Data generated from the GCT model featured 12 substages, with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$ and $\kappa=\frac{2}{9}$.
@@
 \end{figure}
+
+%%%%%%%%
+% Here
+%%%%%%%%
+To more clearly illustrate the performance differences between the LCT and GCT models, fittings with various model structures were conducted to assess changes in percentage error.
+The percentage error from the fitting result, where the model generating the incidence data and the model for fitting are exactly the same, has been used as the baseline (red line).
+As shown in Figure X, when using the same input data, varying the fitting model shape causes the LCT model to experience a surge in RMSLE values, while the GCT model maintains its error rate close to the baseline.
+This result, once again, confirms that the structure of the GCT model merely affects the fitting results, indicating that changes in structure to optimize model performance are no longer necessary.
+
+To inllstriate more clear about the performance difference between LCT and GCT, fittings using different model structure have been done to see how the error would change.
+As shown in Figure x, when treating the percentage error from the the fitting result where the model to generate the incidence data and the model for fitting are exactly the same, varying the fitting model shape by using a different number of substages from the model used to generate the data causes the LCT model to experience a surge in RMSLE values, while the GCT model maintains its error rate close to the baseline.
+This result, once again, confirmed that the structure of the GCT model merely affect the model fitting result, showing that changing in structure in order to find the optimized performance of the model is no longer necessary.
 
@@
     \centering
-    \includegraphics[width= \textwidth]{figures/difftypeFitting.Rout.pdf}
+    \begin{tikzpicture}
+        \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width= \textwidth]{figures/difftypeFitting.Rout.pdf}};
+        \begin{scope}[x={(image.south east)},y={(image.north west)}]
+            % Add labels at desired positions
+            \node at (0.01,0.95) {\textbf{(A)}};
+            \node at (0.65,0.95) {\textbf{(B)}};
+            \node at (0.65,0.45) {\textbf{(C)}};
+        \end{scope}
+    \end{tikzpicture}
     \caption{\textbf{Model fitting using different model types.} \\  Input data for Plots A and B were generated using a 2-substages LCT transmission model simulation with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$.
@@
 In this section, we discuss the parameter estimation results obtained through model fitting.
-The true incidence data were generated using the GCT model with 5 substages, setting the $D$ to 10 and $\kappa$ to $\frac{2}{9}$.
+The true incidence data were generated using the GCT model with 5 substages, setting the $D$ to 10 and $\kappa$ to $\frac{3}{10}$.
 Random noise was added to the true incidence data using a negative binomial distribution, resulting in four different input datasets (a number we specifically set for this analysis).
-The model was fitted using both the GCT and LCT frameworks with 5 to 10 substages using the 4 sets of data.
+The model was fitted using both the GCT and LCT frameworks with 4 to 12 substages using the 4 sets of data.
 As depicted in Figure \ref{parEst}, the precise values for each parameter are marked by a red line, with post-fitting parameter values displayed alongside their 95\% confidence intervals.
@@
 These estimates could deviate substantially from the accurate parameter if an inappropriate structure is chosen.
-On the other hand, although the parameter estimation using the GCT model is not perfect—with kappa and beta being overestimated and D underestimated—the estimated values are less fluctuating across substage numbers, closer to the target values, and consistently include the target values within the 95\% confidence interval.
+On the other hand, although the parameter estimation using the GCT model is not perfect—with kappa and beta being overestimated and D underestimated—the estimated values are less fluctuating across substage numbers, closer to the target values, and generally include the target values within the 95\% confidence interval.
 The discrepancies in parameter estimation counterbalance each other, resulting in a fitting outcome characterized by a low root mean squared logarithmic error (RMSLE).
@@
 The model encounters limitations when aligning with a kappa value of less than $\frac{1}{n}$.
-Overall, it is important to acknowledge the imprecision of parameter estimation, which will be a focus of our subsequent efforts.
+Overall, the result demonstrates that the GCT model estimates more accurate and less fluctuating parameters.
+Moreover, it is important to acknowledge the imprecision in parameter estimation, which will be the focus of our subsequent studies.
 

### 2024-04-30 thesis/draft.tex
temp storage

@@
 
-Not surprisingly, employing the GCT model with 2 substages for both generating and fitting incidence data also resulted in a close match, as indicated by an RMSLE of 2.34\% (Figure \ref{sametypeFittingG}.A), and parameters post-fitting were close to the exact values (Figure \ref{sametypeFittingG}.B).
-Unlike the case with the LCT model, when the structure of the fitting model was altered from 2 substages to 6, the model still exhibited a close match with the original in terms of both shape (Figure \ref{sametypeFittingG}.C) and parameters (Figure \ref{sametypeFittingG}.D).
+Not surprisingly, employing the GCT model with 12 substages for both generating and fitting incidence data also resulted in a close match, as indicated by an RMSLE of 2.34\% (Figure \ref{sametypeFittingG}.A), and parameters post-fitting were close to the exact values (Figure \ref{sametypeFittingG}.B).
+Unlike the case with the LCT model, when the structure of the fitting model was altered from 12 substages to 6, the model still exhibited a close match with the original in terms of both shape (Figure \ref{sametypeFittingG}.C) and parameters (Figure \ref{sametypeFittingG}.D).
 This indicates that the GCT model's fitting performance is minimally impacted by changes in the model structure.
@@
     \end{tikzpicture}
-    \caption{\textbf{Model fitting and parameter estimation in GCT model.} \\ Data generated from the GCT model featured 12 substages, with $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$ and $\kappa=\frac{2}{9}$.
+    \caption{\textbf{Model fitting and parameter estimation in GCT model.} \\ Data generated from the GCT model featured 2 substages, with kappa= $\beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$ and $\kappa=\frac{2}{9}$.
 Remaining details are as described in Figure \ref{sametypeFittingL}.}
@@
 The percentage error from the fitting result, where the model generating the incidence data and the model for fitting are exactly the same, has been used as the baseline (red line).
-As shown in Figure X, when using the same input data, varying the fitting model shape causes the LCT model to experience a surge in RMSLE values, while the GCT model maintains its error rate close to the baseline.
+As shown in Figure \ref{rmsle}, when using the same input data, varying the fitting model shape causes the LCT model to experience a surge in RMSLE values, while the GCT model maintains its error rate close to the baseline.
 This result, once again, confirms that the structure of the GCT model merely affects the fitting results, indicating that changes in structure to optimize model performance are no longer necessary.
 
-To inllstriate more clear about the performance difference between LCT and GCT, fittings using different model structure have been done to see how the error would change.
-As shown in Figure x, when treating the percentage error from the the fitting result where the model to generate the incidence data and the model for fitting are exactly the same, varying the fitting model shape by using a different number of substages from the model used to generate the data causes the LCT model to experience a surge in RMSLE values, while the GCT model maintains its error rate close to the baseline.
-This result, once again, confirmed that the structure of the GCT model merely affect the model fitting result, showing that changing in structure in order to find the optimized performance of the model is no longer necessary.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/rmsle.Rout.pdf}
+    \caption{\textbf{Percentage error with different fitting model structures.} \\ The LCT fitting data were generated by simulating a two-substage LCT transmission model with $D=10, \beta=0.2$, $\mu=0.01$, $N=10000$, $I_0=10$.
+Similarly, GCT fitting data came from simulating the GCT model with identical properties but $\kappa=\frac{5}{9}$.
+The baseline (red) represents the error rate when the fitting and data generation models have the same structure.
+The blue circle indicates LCT errors, and the orange rhombus represents GCT errors.}
+    \label{rmsle}
+\end{figure}
 
@@
 These estimates could deviate substantially from the accurate parameter if an inappropriate structure is chosen.
-On the other hand, although the parameter estimation using the GCT model is not perfect—with kappa and beta being overestimated and D underestimated—the estimated values are less fluctuating across substage numbers, closer to the target values, and generally include the target values within the 95\% confidence interval.
+On the other hand, although the parameter estimation using the GCT model is not perfect—with kappa and beta being underestimated and D overestimated—the estimated values are less fluctuating across substage numbers, closer to the target values, and generally include the target values within the 95\% confidence interval.
 The discrepancies in parameter estimation counterbalance each other, resulting in a fitting outcome characterized by a low root mean squared logarithmic error (RMSLE).

### 2024-04-30 thesis/draft.tex
temp storage

@@
 \lhead{Ningrui Xie}
-\chead{Thesis First Draft}
+\chead{BIO 4C12}
 \rhead{ID: 400323416}
@@
 
-Therefore, formula (1) will be replaced by the following set of equations, which we refer to as the cohort model:
+Therefore, formula (1) will be replaced by the following set of equations (7-10), which we refer to as the cohort model:
 \begin{align}
@@
 
-The Geometric Chain Trick (GCT), on the other hand, features a fixed number of substages along with a set of inter-substage transition rates characterized by a geometric series (Figure \ref{GCTdiagram}), rather than a constant rate, as illustrated below:
+The Geometric Chain Trick (GCT), on the other hand, features a fixed number of substages along with a set of inter-substage transition rates characterized by a geometric series $\gamma_i = ar^{i-1}$ (Figure \ref{GCTdiagram}), rather than a constant rate, as illustrated below (12-16):
 \begin{align}
@@
 Having established the model, our initial step involved analyzing the parameters within the GCT model.
-We aim to determine whether employing a fixed model structure (i.e., a fixed number of substages) and manipulating the parameters $a$ and $r$, which define the geometric transition rate, can replicate the effects of varying the number of substages.
+We aim to determine whether employing a fixed model structure (i.e., a fixed number of substages) and manipulating the parameters $a$ and $r$, which define the geometric transition rate, can achieve the same effects as varying the number of substages.
 Additionally, we aim to evaluate whether the GCT structure offers greater flexibility in the choice of $\kappa$ $(CV^2)$.

### 2025-05-07 Code/Interval.R
Meeting with Ningrui

@@ -5,7 +5,8 @@ library(pseudoErlang)
 
 library(shellpipes)
 loadEnvironments()
-# source("./tempfunc.R")
+
+## objects(); quit()
 
 ######################################################################
 ## Gamma -> Erlang
@@ -101,6 +102,8 @@ ggplot(df, aes(x=CD)) +
   geom_line(aes(y=Interval_pe^2, color='Pseudo Erlang(sq)'), linewidth=1) +
   geom_line(aes(y=Interval_g^2, color='Gamma(sq)'), linewidth=1)
 
+
+quit()
 
 plot(x=cd, y=inversed_time, type="l",
      xlab="Cumulative Density",

### 2025-05-07 Code/tempfunc.R

@@ -289,4 +289,3 @@ qperlang_o <- function(fx, mean, kappa, box=12, log=FALSE) {
 ######################################################################
 
 saveEnvironment()
-
